---
title: "Effects of stress on substantia nigra degeneration in Parkinson's disease"
author: "M.E. Johansson"
date: "2024-07-12"
output: 
  html_document:
    toc: yes
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project description

Parkinson's disease leads to degeneration of the substantia nigra, particularly the posterior segment (pSN), which results in lower dopamine levels, basal ganglia dysfunction, and motor symptoms. There is some evidence suggesting that stress may exacerbate pSN degeneration. Here, we test this hypothesis by assessing whether longitudinal pSN degeneration is influenced by stress reactivity levels.

# Libraries

```{r LIBRARIES, echo=F, warning=F, message=F}
library(tidyverse)
library(tidymodels)
library(multilevelmod)
library(broom.mixed)
library(vip)
library(car)
library(ggeffects)
library(emmeans)
library(readxl)
library(effectsize)
library(lubridate)
library(rmcorr)
library(GGally)

options(contrasts=c("contr.sum", "contr.poly"))
sigsize=5
lsize=0.6
```

# Prepare data

## Load: clinical

Load demographic information.

```{r LOAD-CLIN, echo=F, warning=F, message=F}
# Clinical data
read_clinical_metrics <- function(filename, clinical_metrics){
  df <- read_csv(filename,
                 col_select = all_of(clinical_metrics), show_col_types = F) %>%
    filter(ParticipantType != 'PD_PIT') %>%
    mutate(ParticipantType = if_else(ParticipantType == 'PD_POM','Patient','Healthy'),
           TimepointNr = if_else(ParticipantType == 'Healthy' & TimepointNr == 1, 2, TimepointNr)) %>%
    filter(TimepointNr != 1)
  
  diagnosis <- df %>% select(pseudonym, TimepointNr, DiagParkCertain, DiagParkPersist, DiagParkReplace)
  baseline_exclusion <- diagnosis %>%
      filter(TimepointNr==0, (DiagParkCertain == 'NeitherDisease' | DiagParkCertain == 'DoubtAboutParkinsonism' | DiagParkCertain == 'Parkinsonism')) %>% 
      select(pseudonym, DiagParkCertain)
  visit2_exclusion <- diagnosis %>%
      filter(TimepointNr==2, (DiagParkPersist == 2)) %>% 
      select(pseudonym, DiagParkPersist, DiagParkReplace)
  diag_exclusions <- full_join(baseline_exclusion, visit2_exclusion, by='pseudonym') %>% 
      unique()
  df <- df %>%
      mutate(Misdiagnosis = if_else(pseudonym %in% diag_exclusions$pseudonym,1,0))
  
  df     
}
clinical_metrics <- c('pseudonym','ParticipantType','TimepointNr','Age','Gender','YearsSinceDiag','NpsEducYears',
                      'AsymmetryIndexRiLe.WeightedBradyRig2','DiagParkCertain', 'DiagParkPersist', 'DiagParkReplace')
df_clin <- read_clinical_metrics('P:/3022026.01/pep/ClinVars_10-08-2023/derivatives/merged_manipulated_2024-07-17.csv', 
                                 clinical_metrics) %>%
    mutate(MostAffSide=NA,
           MostAffSide = case_when(
               AsymmetryIndexRiLe.WeightedBradyRig2<0 ~ 'L',
               AsymmetryIndexRiLe.WeightedBradyRig2>0 ~ 'R',
               AsymmetryIndexRiLe.WeightedBradyRig2==0 ~ 'Sym',
               .default = NA
               )
           ) %>%
  relocate(MostAffSide, .after = 'Gender') %>%
    select(-c(DiagParkCertain, DiagParkPersist, DiagParkReplace))

head(df_clin)
```

## Load: stress reactivity

Load stress reactivity classes.

```{r LOAD-STRESS, echo=F, warning=F, message=F}
read_stress_class <- function(filename){
  df <- read_csv(filename, show_col_types = F) %>%
    rename(pseudonym = Subj_ID) %>%
    mutate(SR_cohort = 1)
}
df_stress_class <- read_stress_class("P:/3024006.02/Data/PPP_stressreactivity.csv")

head(df_stress_class)
```

## Load: CRP values

```{r LOAD-CRP, echo=F, warning=F, message=F}
read_stress_crpnfl <- function(filename){
  df.crpnfl <- read_csv(filename, show_col_types = F) %>%
      select(!starts_with('Date')) %>%
      rename(pseudonym = Subj_ID) %>%
      pivot_longer(cols = CRP.Visit1:NFL.Visit3,
                   names_to = c("Var","TimepointNr"),
                   names_pattern = "(.*).Visit(.*)") %>%
      mutate(TimepointNr = as.numeric(TimepointNr) - 1) %>%
      pivot_wider(id_cols = c("pseudonym","TimepointNr"),
                  names_from = "Var",
                  values_from = "value")
  
  df.date <- read_csv(filename, show_col_types = F) %>%
      select(Subj_ID, starts_with('Date')) %>%
      rename(pseudonym = Subj_ID) %>%
      pivot_longer(cols = Date.Visit1:Date.Visit3,
                   names_to = c("Var","TimepointNr"),
                   names_pattern = "(.*).Visit(.*)",
                   values_to = "Date") %>%
      mutate(TimepointNr = as.numeric(TimepointNr) - 1,
             Date = str_replace(Date,'/','-'),
             Date = mdy(Date)) %>%
    select(-Var)
  
  df <- full_join(df.crpnfl,df.date)
  
  df
}
df_stress_crpnfl <- read_stress_crpnfl("P:/3024006.02/Data/Data_CRPNFL_wide.csv") %>%
  select(-Date) %>%
  filter(TimepointNr != 1)

head(df_stress_crpnfl)
```

## Load: free water

Load substantia nigra free water.

```{r LOAD-FW, echo=F, warning=F, message=F}
# DTI data
read_dti_metrics <- function(filename){
    df <- read_delim(filename, delim = ',', trim_ws = T, show_col_types = F)
    df <- df %>%
        mutate(pseudonym = str_extract(IMG,"(?<=qsiprep/).+(?=/ses)"),
               Timepoint = str_extract(IMG,"(?<=ses-).+(?=/metrics)"),
               TimepointNr = if_else(str_detect(Timepoint,'Visit1'),0,2),
               ParticipantType = if_else(str_detect(Timepoint,'POM'),
                                         'Patient','Healthy'),
               TimepointNr = TimepointNr) %>%
        select(-c(IMG,Timepoint)) %>%
        relocate(pseudonym,ParticipantType,TimepointNr) %>%
        arrange(ParticipantType,pseudonym,TimepointNr)
    
    df
}
df_dti <- read_dti_metrics('P:/3024006.02/Analyses/MJF_FreeWater/data/n2_pasternak_fw/FW_stats_avg.csv') %>%
  left_join(., read_dti_metrics('P:/3024006.02/Analyses/MJF_FreeWater/data/n2_pasternak_fw/FW_stats_sd.csv'), by = c('pseudonym','ParticipantType','TimepointNr'))

head(df_dti)
```

## Load: cortical mean diffusivity

Load cortical mean diffusivity.

```{r LOAD-MD, echo=F, warning=F, message=F}
# FreeSurfer data
read_fs_metrics <- function(filename){
  df <- read_csv(filename, show_col_types = F) %>%
    rename(pseudonym=SubjID) %>%
    mutate(TimepointNr = str_sub(pseudonym, start=26,end=27),
           TimepointNr = if_else(TimepointNr=='t1',0,2),
           pseudonym = str_sub(pseudonym, start=1, end=24)) %>%
    relocate(pseudonym, TimepointNr)
  
  df
}
fs_exclusions <- c('sub-POMU56F70F8137CF0C55', 'sub-POMU6059DC1B31E11124')
volume_rois <- c('thal','caud','put','pal','accumb','amyg')
surface_rois <- c( "caudalanteriorcingulate","caudalmiddlefrontal",
                   "cuneus","entorhinal","fusiform","inferiorparietal",
                   "inferiortemporal","isthmuscingulate","lateraloccipital",
                   "lateralorbitofrontal","lingual","medialorbitofrontal",
                   "middletemporal","parahippocampal","paracentral",
                   "parsopercularis","parsorbitalis","parstriangularis",
                   "pericalcarine","postcentral","posteriorcingulate",
                   "precentral","precuneus","rostralanteriorcingulate",
                   "rostralmiddlefrontal","superiorfrontal","superiorparietal",
                   "superiortemporal","supramarginal","frontalpole",
                   "temporalpole","transversetemporal","insula")
df_fs_vol <- read_fs_metrics('P:/3022026.01/pep/bids/derivatives/freesurfer_v7.3.2/outputs/metrics/measures/LandRvolumes.csv') %>%
  select(pseudonym,TimepointNr,TSCGMV,TCGMV,TGMV,ICV,contains(volume_rois))
df_fs_surf <- read_fs_metrics('P:/3022026.01/pep/bids/derivatives/freesurfer_v7.3.2/outputs/metrics/measures/CorticalMeasuresENIGMA_SurfAvg.csv') %>%
  select(pseudonym,TimepointNr,contains(surface_rois))
df_fs_thick <- read_fs_metrics('P:/3022026.01/pep/bids/derivatives/freesurfer_v7.3.2/outputs/metrics/measures/CorticalMeasuresENIGMA_ThickAvg.csv') %>%
  select(pseudonym,TimepointNr,contains(surface_rois))
metric <- c('fsl_MD','TensorFWCorrected_dcmp_MD','TensorDTINoNeg_dcmp_MD','dipy-MDc','dipy-MD','FW')
metric <- metric[1]
df_fs_md <- read_csv(paste0('P:/3022026.01/pep/bids/derivatives/qsiprep/measures/1_NoPVC_projfrac02-08-01_correctb0orientation/SurfaceMeasures_',metric,'_T1w_Mean.csv'),
                     show_col_types = F) %>%
  rename(pseudonym = SubjID) %>%
  mutate(TimepointNr = if_else(str_detect(Timepoint,'Visit1'),0,2)) %>%
  relocate(pseudonym, TimepointNr) %>%
  select(-Timepoint) %>% select(pseudonym,TimepointNr,contains(surface_rois)) %>%
  na.omit()
df_fs <- df_fs_vol %>% 
  left_join(df_fs_surf, by = c('pseudonym','TimepointNr')) %>%
  left_join(df_fs_thick, by = c('pseudonym','TimepointNr')) %>% 
  left_join(df_fs_md, by = c('pseudonym', 'TimepointNr')) %>%
  filter(! pseudonym %in% fs_exclusions)
for(roi in volume_rois){
  bi_roi <- tibble(
    roi=(
      (df_fs %>% pull(paste0('L',roi)) + df_fs %>% pull(paste0('R',roi))) / 2
    )
  )
  colnames(bi_roi) <- paste0('bi_',roi)
  df_fs <- df_fs %>% bind_cols(bi_roi)
}
for(roi in surface_rois){
  for(m in c('surfavg','thickavg','dti')){
    bi_roi <- tibble(
      roi=(
        (df_fs %>% pull(paste('L',roi,m,sep='_')) + df_fs %>% pull(paste('R',roi,m,sep='_'))) / 2
      )
    )
    colnames(bi_roi) <- paste('bi',roi,m,sep='_')
    df_fs <- df_fs %>% bind_cols(bi_roi)
  }
}

head(df_fs)
```

## Combine data

Merge all data into a single set.

```{r COMBINE-DATA, echo=F, warning=F, message=F}
df <- df_clin %>%
  full_join(., df_stress_class, by = 'pseudonym'
            ) %>%
  full_join(., df_stress_crpnfl, by = c("pseudonym","TimepointNr")
            ) %>%
  full_join(., df_dti, by = c('pseudonym','ParticipantType','TimepointNr')
            ) %>%
  full_join(., df_fs, by = c('pseudonym','TimepointNr')
            )

head(df)
```

## Feature engineering

 - Factorize
 - Define complete cases
 - Add visit-specific and delta columns

```{r ENGINEER, echo=F, warning=F, message=F}
# Factorize
df <- df %>%
  mutate(TimepointNr = factor(TimepointNr, levels = c(0,2), labels = c(0,2)),
         Gender = factor(Gender, levels = c('Female','Male'), labels = c(0,1)),
         SR_class = factor(SR_class, levels = c(1,2), labels = c(0,1)),
         SR_cohort = if_else(is.na(SR_cohort),0,SR_cohort),
         SR_class2 = case_when(
           ParticipantType == 'Healthy' ~ 0,
           ParticipantType == 'Patient' & SR_class == 0 ~ 1,
           ParticipantType == 'Patient' & SR_class == 1 ~ 2
         ),
         SR_class2 = factor(SR_class2, levels = c(0,1,2), labels = c(0,1,2))) %>%
  relocate(SR_class2, .after = SR_class)

# Which subjects have complete data?
subs <- unique(df$pseudonym)
tmp <- c()
for(i in subs){
  ntp <- df %>%
    filter(pseudonym==i) %>%
    select(TimepointNr) %>%
    nrow()
  tmp <- tmp %>%
    bind_rows(tibble(pseudonym=i,ntp=ntp))
}
tmp <- tmp %>%
  mutate(complete_case = if_else(ntp > 1, 1, 0)) %>%
  select(-ntp)
df <- df %>%
  left_join(., tmp, by = 'pseudonym') %>%
  relocate(complete_case, .after = SR_cohort)

# Add time-specific values as separate columns
add_timepoints_as_columns <- function(dat, id_cols, timevar, var){
  
  # Separate timepoints and calculate delta
  dat2 <- dat %>%
    select(all_of(id_cols),
           all_of(timevar),
           all_of(var)) %>%
    pivot_wider(id_cols = id_cols,
                names_from = 'TimepointNr',
                names_prefix = 'T',
                values_from = all_of(var)) %>%
    mutate(T2sub0 = T2-T0) %>%
    rename()
  
  # Add variable name
  dat2 <- rename_with(
    dat2,
    ~ paste0(.x, '_', var, recycle0 = TRUE),
    starts_with("T")
  )
  
  # Join with old data
  dat <- 
    left_join(dat, dat2, by = c(id_cols))
  
  # Print
  dat

}
id_cols <- colnames(df)[1:2]
timevar <- 'TimepointNr'
vars <- colnames(df)[14:length(colnames(df))]
for(i in 1:length(vars)){
  df <- add_timepoints_as_columns(df, id_cols, timevar, vars[i])
}
```

## Exclusions

 - Exclude misdiagnosis
 - Retain patients with an SR class and healthy controls.
 - Exclude participants who did not pass QC procedures.

```{r EXCLUDE, echo=F, warning=F, message=F}
# Exclude misdoagnised patients
df <- df %>%
    filter(Misdiagnosis != 1)

# Retain cohort of interest
df <- df %>%
  filter(SR_cohort == 1 | ParticipantType == 'Healthy')

# Define partcipants who failed QC
read_dti_exclusions <- function(filename) {
  df <- read_excel(filename) %>%
    filter(retain == 'N') %>%
    select(pseudonym,session) %>%
    distinct() %>%
    mutate(TimepointNr = if_else(str_detect(session,'Visit1'),0,2),
           TimepointNr = factor(TimepointNr),
           Exclusion = 1) %>%
    select(-session)
}

exclusions <- read_dti_exclusions('P:/3024006.02/Analyses/MJF_FreeWater/QC/FW/Exclusions_and_Outliers_FW.xlsx')
df <- df %>%
  left_join(exclusions, by = c('pseudonym','TimepointNr')) %>%
  mutate(Exclusion = if_else(is.na(Exclusion),0,Exclusion)) %>%
  rename(Exclusion_FW = Exclusion)

exclusions <- read_dti_exclusions('P:/3024006.02/Analyses/MJF_FreeWater/QC/MD/Exclusions_and_Outliers_MD.xlsx')
df <- df %>%
  left_join(exclusions, by = c('pseudonym','TimepointNr')) %>%
  mutate(Exclusion = if_else(is.na(Exclusion),0,Exclusion)) %>%
  rename(Exclusion_MD = Exclusion)

df_fw <- df %>%
  filter(Exclusion_FW != 1)
df_md <- df %>%
  filter(Exclusion_MD != 1)
```

# pSN FW analysis

## Descriptives

### pSN by group

pSN FW is generally higher in patients than controls, but there doesn't seem to be any difference between stress reactivity groups. There are also no apparent differences in longitudinal change between any of the groups.

```{r SA-DESCRIBE, echo=F, warning=F, message=F}
# Describe
df_fw %>%
  count(SR_class2,TimepointNr)

# Plot
g <- df_fw %>%
  ggplot(aes(x=SR_class2, y = pSN_avg, color=TimepointNr)) +
  geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
  geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
  stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
               mapping = aes(group=TimepointNr), position = position_dodge(width=0.8)) +
  theme_bw() + 
  scale_x_discrete(labels=c('Healthy','Low','High')) +
  scale_color_viridis_d(option='mako', begin = .1, end = .6,
                        direction = -1) + 
  ggtitle('Posterior substantia nigra') +
  ylab('Free water (fraction)') + xlab('') + 
  guides(color = guide_legend(title='Year')) +
  ylim(0, 0.7)+
  annotate(geom='text',label='N=50',x=0.8,y=0.1) +
  annotate(geom='text',label='N=47',x=1.2,y=0.1) +
  annotate(geom='text',label='N=217',x=1.8,y=0.1) +
  annotate(geom='text',label='N=209',x=2.2,y=0.1) + 
  annotate(geom='text',label='N=119',x=2.8,y=0.1) +
  annotate(geom='text',label='N=109',x=3.2,y=0.1)
y <- 0.65
d = 0.015
segmap <- data.frame(
        x =    c(1.00, 1.00),
        xend = c(2.00, 3.00),
        y =    c(y,    y-d*2),
        yend = c(y,    y-d*2)
)
sigmap <- data.frame(
        x =     c(1.50, 2.00),
        y =     c(y,    y-d*2),
        label = c('***', '***')
)
g <- g + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     linewidth = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)
print(g)
```

### Histograms

```{r SA-HIST, echo=F, warning=F, message=F}

tmp <- df %>%
    filter(TimepointNr==0,
         T2sub0_CRP > -10,
         T2sub0_NFL < 50) %>%
    select(pseudonym,ParticipantType,TimepointNr,T0_pSN_avg,T2_pSN_avg,T2sub0_pSN_avg,T0_CRP,T2_CRP,T2sub0_CRP,T0_NFL,T2_NFL,T2sub0_NFL)
ggpairs(tmp, 
        columns = 4:12, 
        aes(alpha=0.5),
        lower = list(continuous = 'smooth')) %>%
    print()

```

## Mixed-effects modelling

Using mixed-effects modelling, we test the hypothesis that stress reactivity predisposes patients to faster pSN degeneration, as represented by a GROUP x TIME interaction.

```{r SA-MODEL, echo=F, warning=F, message=F}
# Model
sr_mod <-
  linear_reg() %>%
  set_engine('lmer') %>%
  set_mode('regression')

# Recipe
sr_rec <- 
  recipe(pSN_avg ~ SR_class2 + TimepointNr + Age + Gender + NpsEducYears + pseudonym, data = df_fw) %>%
  add_role(pseudonym, new_role = 'exp_unit') %>%
  step_interact(terms = ~ SR_class2:TimepointNr) %>%
  step_center(all_numeric_predictors()) %>%
  step_log(pSN_avg, base = 10) %>%
  step_zv(all_predictors(), -has_role('exp_unit'))

# Workflow
sr_wf <-
  workflow() %>%
  add_model(sr_mod, formula = pSN_avg ~ SR_class2 * TimepointNr + Age + Gender + NpsEducYears + (1|pseudonym)) %>%
  add_recipe(sr_rec)

# Fit
sr_fit <- 
  sr_wf %>%
  fit(data = df_fw)

# Assess
sr_ext_fit <- extract_fit_engine(sr_fit)
Anova(sr_ext_fit, type = 3) %>% 
  print()
eta_squared(sr_ext_fit) %>%
  print()
emmeans(sr_ext_fit, ~ TimepointNr|SR_class2, type = 'response') %>% 
  contrast(em, interaction = (c('revpairwise','revpairwise')), adjust = 'mvt', type = 'response', by = NULL) %>% 
  print()
emmeans(sr_ext_fit, ~ TimepointNr|SR_class2, type = 'response') %>% 
  contrast(em, interaction = (c('revpairwise','revpairwise')), adjust = 'mvt', type = 'response') %>% 
  print()
emmeans(sr_ext_fit, ~SR_class2, type = 'response') %>% contrast(interaction = c('revpairwise')) %>% 
  print()
emmeans(sr_ext_fit, ~TimepointNr, type = 'response') %>% contrast(interaction = c('revpairwise')) %>% 
  print()
ggemmeans(sr_ext_fit, c('SR_class2','TimepointNr')) %>% 
  plot()
```

Conclusion: As suggested by the plot above, patients have more pSN FW. However, there are no differences between the stress reactivity groups with respect to overall pSN FW, or longitudinal change in it.

## Three-sample t-test

Using standard linear modelling, we test the same hypothesis as above, but on the deltas of pSN FW in participants with complete data. Adjustment for baseline increases sensitivity by accounting for individual differences in longitudinal change and regression-to-the-mean effects. This would be achievable with the mixed effects approach if we had more data to fit random slopes for time, but we don't.

```{r SA-MODEL-DELTA, echo=F, warning=F, message=F}
# Subset data
tmp <- df_fw %>%
  filter(TimepointNr==0,
         complete_case==1) %>%
  select(pseudonym,SR_class2,Age,Gender,NpsEducYears,T2sub0_pSN_avg,T0_pSN_avg) %>%
  na.omit()
tmp <- bind_cols(tmp,
                 T2sub0_pSN_avg_rz = lm(T2sub0_pSN_avg~T0_pSN_avg,data=tmp)$residuals)

# Model
sr_d_mod <-
  linear_reg() %>%
  set_engine('lm') %>%
  set_mode('regression')

# Recipe
sr_d_rec <-
  recipe(T2sub0_pSN_avg_rz ~ SR_class2 + Age + Gender + NpsEducYears, data = tmp) %>%
    step_dummy(Gender) %>%
    step_center(all_numeric_predictors()) %>%
    step_zv(all_predictors())

# Workflow
sr_d_wf <-
  workflow() %>%
  add_model(sr_d_mod) %>%
  add_recipe(sr_d_rec)

# Fit
sr_d_fit <- 
  sr_d_wf %>%
  fit(data = tmp)

# Assess
sr_d_ext_fit <-
  sr_d_fit %>%
  extract_fit_engine()
Anova(sr_d_ext_fit, type = 3) %>%
  print()
emmeans(sr_d_ext_fit, ~SR_class2, type = 'response') %>% contrast(interaction = c('revpairwise')) %>% 
  print()
emmeans(sr_d_ext_fit, ~SR_class2, type = 'response') %>% contrast(interaction = c('identity')) %>% 
  print()
ggemmeans(sr_d_ext_fit, 'SR_class2') %>% 
  plot()
```

Conclusion: The results are fully consistent with the mixed-effects approach. Stress reactivity groups do not differ in longitudinal change. However, the added sensitivity of this approach reveals that patients, regardless of group, show increased pSN FW over time relative to controls. 

## Correlations

Using standard linear modelling, we test whether change in posterior SN FW is influenced by change in CRP and NFL.

### FW ~ CRP

#### Delta

```{r CORRELATE-FW-CRP, echo=F, warning=F, message=F}

# Subset data and filter outliers
tmp <- df %>%
  filter(TimepointNr==0,
         ParticipantType=='Patient',
         T2sub0_CRP > -10,
         complete_case==1) %>%
  select(pseudonym,Age,Gender,NpsEducYears,YearsSinceDiag,T2sub0_pSN_avg,T0_pSN_avg,T2_pSN_avg,T2sub0_CRP,T0_CRP,T2_CRP) %>%
  na.omit()
tmp <- bind_cols(tmp,
                 T2sub0_pSN_avg_rz = lm(T2sub0_pSN_avg~T0_pSN_avg,data=tmp)$residuals,
                 T2sub0_CRP_rz = lm(T2sub0_CRP~T0_CRP,data=tmp)$residuals)

# Model
sr_d_mod <- 
  linear_reg() %>%
  set_engine('lm') %>%
  set_mode('regression')

# Recipe
sr_d_rec <- 
  recipe(T2sub0_pSN_avg_rz ~ T2sub0_CRP_rz + Age + Gender + NpsEducYears + YearsSinceDiag, data = tmp) %>%
  step_dummy(Gender) %>%
  step_center(all_numeric_predictors()) %>%
  step_zv(all_predictors())

# Workflow
sr_d_wf <- 
  workflow() %>%
  add_model(sr_d_mod) %>%
  add_recipe(sr_d_rec)

# Fit
sr_d_fit <-
  sr_d_wf %>%
  fit(data = tmp)

# Assess
sr_d_fit_extract <-
  sr_d_fit %>%
  extract_fit_engine()
Anova(sr_d_fit_extract, type = 3) %>%
  print()
summary(sr_d_fit_extract) %>%
  print()
pred <- ggpredict(sr_d_fit_extract, terms = "T2sub0_CRP_rz")
g <- plot(pred, add.data = TRUE,color='darkred') + theme_bw() + ggtitle('') +
        ylab('Change in pSN FW') + xlab('Change in CRP') +
        geom_hline(yintercept = 0, color = 'darkgrey', lty = 2) + geom_vline(xintercept = 0, color = 'darkgrey', lty = 2)
print(g)

# Alternative: Spearman
cor.test(tmp$T2sub0_pSN_avg_rz,tmp$T2sub0_CRP_rz,method='spearman') %>%
    print()

```

#### RMC

```{r CORRELATE-FW-CRP-RMC, echo=F, warning=F, message=F}
# Subset data and filter outliers
tmp <- df %>%
    filter(ParticipantType=='Patient',
           T2sub0_CRP > -10,
           complete_case==1) %>%
    select(pseudonym,pSN_avg,CRP) %>%
    na.omit()

# Fit
rmc <- rmcorr('pseudonym','pSN_avg','CRP',dataset = tmp)

# Assess
print(rmc)
plot(rmc)
```

Conclusion: There is no association with change in CRP.

### FW ~ NFL

#### Delta

```{r CORRELATE-FW-NFL, echo=F, warning=F, message=F}

# Subset data and filter outliers
tmp <- df %>%
  filter(TimepointNr==0,
         ParticipantType=='Patient',
         T2sub0_NFL < 50,
         complete_case==1) %>%
  select(pseudonym,Age,Gender,NpsEducYears,YearsSinceDiag,T2sub0_pSN_avg,T0_pSN_avg,T2_pSN_avg,T2sub0_NFL,T0_NFL,T2_NFL) %>%
  na.omit()
tmp <- bind_cols(tmp,
                 T2sub0_pSN_avg_rz = lm(T2sub0_pSN_avg~T0_pSN_avg,data=tmp)$residuals,
                 T2sub0_NFL_rz = lm(T2sub0_NFL~T0_NFL,data=tmp)$residuals)

# Model
sr_d_mod <- 
  linear_reg() %>%
  set_engine('lm') %>%
  set_mode('regression')

# Recipe
sr_d_rec <- 
  recipe(T2sub0_pSN_avg_rz ~ T2sub0_NFL_rz + Age + Gender + NpsEducYears + YearsSinceDiag, data = tmp) %>%
  step_dummy(Gender) %>%
  step_center(all_numeric_predictors()) %>%
  step_zv(all_predictors())

# Workflow
sr_d_wf <- 
  workflow() %>%
  add_model(sr_d_mod) %>%
  add_recipe(sr_d_rec)

# Fit
sr_d_fit <-
  sr_d_wf %>%
  fit(data = tmp)

# Assess
sr_d_fit_extract <-
  sr_d_fit %>%
  extract_fit_engine()
Anova(sr_d_fit_extract, type = 3) %>%
  print()
summary(sr_d_fit_extract) %>%
  print()
pred <- ggpredict(sr_d_fit_extract, terms = "T2sub0_NFL_rz")
g <- plot(pred, add.data = TRUE,color='darkred') + theme_bw() + ggtitle('') +
        ylab('Change in pSN FW') + xlab('Change in NFL') +
        geom_hline(yintercept = 0, color = 'darkgrey', lty = 2) + geom_vline(xintercept = 0, color = 'darkgrey', lty = 2)
print(g)

# Alternative: Spearman
cor.test(tmp$T2sub0_pSN_avg_rz,tmp$T2sub0_NFL_rz,method='spearman') %>%
    print()

```

#### RMC

```{r CORRELATE-FW-NFL-RMC, echo=F, warning=F, message=F}
# Subset data and filter outliers
tmp <- df %>%
    filter(ParticipantType=='Patient',
           T2sub0_NFL < 50,
           complete_case==1) %>%
    select(pseudonym,pSN_avg,NFL) %>%
    na.omit()

# Fit
rmc <- rmcorr('pseudonym','pSN_avg','NFL',dataset = tmp)

# Assess
print(rmc)
plot(rmc)
```

Conclusion: There is no association with change in NFL.

# Bonus: Cortical MD analysis

## Descriptives

```{r DESCRIBE-MD, echo=F, warning=F, message=F, eval=F}
# Describe
df_md %>%
  count(SR_class2, TimepointNr) %>%
  print()

# Plot
vars_group <- c('SR_class2','TimepointNr')
labels <- c('inferiorparietal','superiorparietal','precuneus','supramarginal',
            'paracentral','postcentral','precentral','superiorfrontal','caudalmiddlefrontal')
measure <- c('dti')
hemi <- c('bi')
vars_dv <- paste(hemi,labels,measure,sep='_')

df.plot <- df_md %>% 
  select(pseudonym,SR_class2,TimepointNr,
         all_of(vars_dv)) %>%
  pivot_longer(cols = all_of(vars_dv),
               names_to = 'ROI',
               values_to = 'Value') %>%
  mutate(ROI=factor(ROI,
                    levels=vars_dv,
                    labels=vars_dv),
         SR_class2 = factor(SR_class2),
         TimepointNr = factor(TimepointNr)) %>%
  na.omit()

g <- ggplot(df.plot, aes(y=Value, x=SR_class2, fill=TimepointNr)) + 
  geom_boxplot() + 
  facet_grid(~ROI) +
  theme_bw() +
  ylab('Cortical MD') + xlab('SR class') +
  scale_fill_viridis_d(option='mako', begin = .7, end = .5,
                        direction = 1, labels=c('BA','2Y'),
                       guide = guide_legend(title='Time'))
print(g)
```

## Modelling

```{r MODEL-MD, echo=F, warning=F, message=F, eval=F}
# Model
sr_mod <- 
  linear_reg() %>%
  set_engine('lmer') %>%
  set_mode('regression')

sr_d_mod <-
  linear_reg() %>%
  set_engine('lm') %>%
  set_mode('regression')

# By-region modelling
labels <- c('inferiorparietal','superiorparietal','precuneus','supramarginal',
            'paracentral','postcentral','precentral',
            'superiorfrontal','caudalmiddlefrontal')
measure <- c('dti')
hemi <- c('bi')
cols <- paste(hemi,labels,measure,sep='_')

tab <- c()
for(lab in cols){
  
  ## LMER
  # Recipe
  f <- as.formula(paste0(lab, ' ~ SR_class2 + TimepointNr + Age + Gender + pseudonym'))
  sr_rec <- 
    recipe(f, data = df_md) %>%
    add_role(pseudonym, new_role = 'exp_unit') %>%
    step_interact(terms = ~ SR_class2:TimepointNr) %>%
    step_center(Age) %>%
    step_log(all_of(lab), base = 10) %>%
    step_zv(all_predictors(), -has_role('exp_unit'))
  
  # Workflow
  sr_wf <-
    workflow() %>%
    add_model(sr_mod, formula = as.formula(paste0(lab, ' ~ SR_class2 * TimepointNr + Age + Gender + (1|pseudonym)'))) %>%
    add_recipe(sr_rec)
  
  # Fit
  sr_fit <- 
    sr_wf %>%
    fit(data = df_md)
  
  # Assess
  sr_ext_fit <- extract_fit_engine(sr_fit)
  a <- Anova(sr_ext_fit, type = 3)
  etas <- eta_squared(sr_ext_fit)
  em01 <- emmeans(sr_ext_fit, ~ TimepointNr|SR_class2, type = 'response') %>% 
    contrast(em, interaction = (c('revpairwise','revpairwise')), adjust = 'mvt', type = 'response', by = NULL)
  em02 <- emmeans(sr_ext_fit, ~ TimepointNr|SR_class2, type = 'response') %>% 
    contrast(em, interaction = (c('revpairwise','revpairwise')), adjust = 'mvt', type = 'response')
  em03 <- emmeans(sr_ext_fit, ~SR_class2, type = 'response') %>% contrast(interaction = c('revpairwise'))
  em04 <- emmeans(sr_ext_fit, ~TimepointNr, type = 'response') %>% contrast(interaction = c('revpairwise'))
  
  ## DELTA
  # Subset data
  lab_d <- paste0('T2sub0_', lab)
  lab_t0 <- paste0('T0_', lab)
  tmp <- df_md %>%
    filter(TimepointNr==0) %>%
    select(pseudonym,SR_class2,Age,Gender,all_of(c(lab_d, lab_t0))) %>%
    na.omit()
  
  # Recipe
  f <- as.formula(paste0(lab_d, ' ~ SR_class2 + Age + Gender + ', lab_t0))
  sr_d_rec <-
    recipe(f, data = tmp) %>%
    step_center(Age,all_of(c(lab_d,lab_t0))) %>%
    step_zv(all_predictors())
  
  # Workflow
  sr_d_wf <-
    workflow() %>%
    add_model(sr_d_mod) %>%
    add_recipe(sr_d_rec)
  
  # Fit
  sr_d_fit <- 
    sr_d_wf %>%
    fit(data = tmp)
  
  # Assess
  sr_d_ext_fit <-
    sr_d_fit %>%
    extract_fit_engine()
  a_d <- Anova(sr_d_ext_fit, type = 3)
  etas_d <- eta_squared(a_d)
  em01_d <- emmeans(sr_d_ext_fit, ~SR_class2, type = 'response') %>% contrast(interaction = c('revpairwise'))
  em02_d <- emmeans(sr_d_ext_fit, ~SR_class2, type = 'response') %>% contrast(interaction = c('identity'))
  
  ## Summarize
  tab <- bind_rows(tab,
                   tibble(label=lab,
                          p_Group_x_Time=a[rownames(a)=='SR_class2:TimepointNr',3],
                          p_Time=a[rownames(a)=='TimepointNr',3],
                          p_Group=a[rownames(a)=='SR_class2',3],
                          eta_squared=etas$Eta2_partial[etas$Parameter=='SR_class2:TimepointNr'],
                          p_Delta_Group=a_d[rownames(a_d)=='SR_class2',4],
                          eta_squared_d=etas_d$Eta2_partial[etas_d$Parameter=='SR_class2'])
  )
}
print(tab)
```


