---
title: 'Structural changes related to clinical progression'
author: "M.E. Johansson"
date: "10/30/2023"
output: 
  html_document: 
    toc: yes
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project description

This project explores whether PD-related changes in brain activity are related to underlying changes in brain structure. Here, we focus on tissue integrity in the substantia nigra and cortex, as measured using diffusion-weighted imaging. The substantia nigra is investigated using free water imaging. The cortex is investigated using mean diffusivity imaging.

# Main

## Data preparation

Loading the following data:

 - Free water (FW) estimates from the substantia nigra (SN)
 - Surface-projected estimates of cortical mean diffusivity (cMD)
 - Compensatory brain activity, to correlate with cMD
 - Clinical data

```{r, echo=F,warning=F,message=F}
# options(contrasts=c("contr.treatment", "contr.poly"))
options(contrasts=c("contr.sum", "contr.poly"))
sigsize=5
lsize=0.6
legend_position <- c(0.8,1.1)

library(ppcor)
library(tidyverse)
library(tidymodels)
library(multilevelmod)
library(readxl)
library(sjPlot)
library(cowplot)
library(lme4)
library(lmerTest)
library(broom.mixed)
library(vtable)
library(rmcorr)
library(msm)
library(emmeans)
library(ggeffects)
library(GGally)
library(MatchIt)
library(rstatix)
library(car)
library(effectsize)
library(ggpubr)
library(MatchIt)
library(extrafont)
loadfonts(device='all',quiet = TRUE)

# DTI data
read_dti_metrics <- function(filename){
    df <- read_delim(filename, delim = ',', trim_ws = T)
    df <- df %>%
        mutate(pseudonym = str_extract(IMG,"(?<=qsiprep/).+(?=/ses)"),
               Timepoint = str_extract(IMG,"(?<=ses-).+(?=/metrics)"),
               TimepointNr = if_else(str_detect(Timepoint,'Visit1'),0,2),
               ParticipantType = if_else(str_detect(Timepoint,'POM'),
                                         'Patient','Healthy'),
               TimepointNr = TimepointNr) %>%
        select(-c(IMG,Timepoint)) %>%
        relocate(pseudonym,ParticipantType,TimepointNr) %>%
        arrange(ParticipantType,pseudonym,TimepointNr)
    
    df
}
df_dti <- read_dti_metrics('/project/3024006.02/Analyses/MJF_FreeWater/data/n2_pasternak_fw/FW_stats_avg.csv') %>%
  left_join(., read_dti_metrics('/project/3024006.02/Analyses/MJF_FreeWater/data/n2_pasternak_fw/FW_stats_sd.csv'), by = c('pseudonym','ParticipantType','TimepointNr'))

# # fMRI data (basal ganglia activity from BRAIN 2024)
# read_fmri <- function(fname){
#   dat <- read_csv(fname) %>%
#     select(-scans) %>%
#     group_by(pseudonym,Group) %>%
#     summarise(Beta = mean(Vals)) %>%
#     ungroup() %>%
#     mutate(ParticipantType = if_else(Group=='PD_POM','Patient','Healthy'),
#            TimepointNr = 0) %>%
#     select(-Group)
#   
# }
# df_fmri <- read_fmri('/project/3024006.02/Analyses/BRAIN_2023/fMRI/Group_comparisons/HcOn_x_ExtInt2Int3Catch_NoOutliers/marsbar/HC_PD_(IntExt)_-27_-15_6.csv') %>%
#   rename(BG_activity = Beta) %>%
#   left_join(., read_fmri('/project/3024006.02/Analyses/BRAIN_2023/fMRI/Group_comparisons/HcOn_x_ExtInt2Int3Catch_NoOutliers/marsbar/HC_PD_(IntExt)_17_-47_-20.csv')) %>%
#   rename(CB_activity = Beta) %>%
#   left_join(., read_fmri('/project/3024006.02/Analyses/BRAIN_2023/fMRI/Group_comparisons/HcOn_x_ExtInt2Int3Catch_NoOutliers/marsbar/HC_PD_(IntExt)_-31_-27_62.csv')) %>%
#   rename(M1_activity = Beta) %>%
#   relocate(BG_activity, .after = TimepointNr)

# fMRI data (longitudinal changes in compensation correlated with bradykinesia progression)
read_fmri2 <- function(fname){
  dat <- read_csv(fname, col_names = F) %>%
    mutate(X1=basename(X1),
           pseudonym = str_sub(X1, 8,31),
           ParticipantType = str_sub(X1, 1,6),
           ParticipantType = if_else(ParticipantType=='PD_POM','Patient','Healthy')) %>%
    relocate(pseudonym,ParticipantType) %>%
    select(-c(X1))
  dat
}
df_compensation_AddCov2 <- read_fmri2('/project/3024006.02/Analyses/motor_task/Group/Longitudinal/FSL/stats/con_0007/vals/rand_delta_clincorr_all_vxlEV_AddCov2_tfce_corrp_tstat2_stats_stats_avg_agg.txt') %>%
  rename(T2sub0_R_6a_id1=X2,T2sub0_L_6a_id2=X3,T2sub0_L_6a_id3=X4,T2sub0_R_6a_id4=X5,T2sub0_R_6d_id5=X6,T2sub0_R_FEF_id6=X7,T2sub0_L_6ma_id7=X8) %>%
  mutate(TimepointNr=0,contrast='con_0007', .after = ParticipantType) %>%
  left_join(., read_fmri2('/project/3024006.02/Analyses/motor_task/Group/Longitudinal/FSL/stats/con_0007/vals/rand_delta_clincorr_all_vxlEV_AddCov2_tfce_corrp_tstat2_stats_BASELINE_stats_avg_agg.txt')) %>%
  rename(T0_R_6a_id1=X2,T0_L_6a_id2=X3,T0_L_6a_id3=X4,T0_R_6a_id4=X5,T0_R_6d_id5=X6,T0_R_FEF_id6=X7,T0_L_6ma_id7=X8) %>%
  mutate(T2sub0_bi_premotor_comp = (T2sub0_R_6a_id1+T2sub0_L_6a_id2+T2sub0_L_6a_id3+T2sub0_R_6a_id4+T2sub0_R_6d_id5+T2sub0_R_FEF_id6+T2sub0_L_6ma_id7) / 7,
         T0_bi_premotor_comp = (T0_R_6a_id1+T0_L_6a_id2+T0_L_6a_id3+T0_R_6a_id4+T0_R_6d_id5+T0_R_FEF_id6+T0_L_6ma_id7) / 7)

df_compensation_AddCov0 <- read_fmri2('/project/3024006.02/Analyses/motor_task/Group/Longitudinal/FSL/stats/con_0007/vals/rand_delta_clincorr_all_vxlEV_tfce_corrp_tstat2_stats_stats_avg_agg.txt') %>%
  rename(T2sub0_R_6a_id1=X2,T2sub0_L_LIPv_id2=X3,T2sub0_L_6ma_id3=X4,T2sub0_R_6d_id4=X5,T2sub0_R_FEF_id5=X6,T2sub0_R_6a_id6=X7) %>%
  mutate(TimepointNr=0,contrast='con_0007', .after = ParticipantType) %>%
  left_join(., read_fmri2('/project/3024006.02/Analyses/motor_task/Group/Longitudinal/FSL/stats/con_0007/vals/rand_delta_clincorr_all_vxlEV_tfce_corrp_tstat2_stats_BASELINE_stats_avg_agg.txt')) %>%
  rename(T0_R_6a_id1=X2,T0_L_LIPv_id2=X3,T0_L_6ma_id3=X4,T0_R_6d_id4=X5,T0_R_FEF_id5=X6,T0_R_6a_id6=X7) %>%
  mutate(T2sub0_bi_premotor_comp = (T2sub0_R_6a_id1+T2sub0_L_6ma_id3+T2sub0_R_6d_id4+T2sub0_R_FEF_id5+T2sub0_R_6a_id6) / 5,
         T0_bi_premotor_comp = (T0_R_6a_id1+T0_L_6ma_id3+T0_R_6d_id4+T0_R_FEF_id5+T0_R_6a_id6) / 5,
         T2sub0_L_parietal_comp = T2sub0_L_LIPv_id2,
         T0_L_parietal_comp = T0_L_LIPv_id2,
         T2sub0_bi_parietopremotor_comp = (T2sub0_R_6a_id1+T2sub0_L_6ma_id3+T2sub0_R_6d_id4+T2sub0_R_FEF_id5+T2sub0_R_6a_id6+T2sub0_L_LIPv_id2) / 6,
         T0_bi_parietopremotor_comp = (T0_R_6a_id1+T0_L_6ma_id3+T0_R_6d_id4+T0_R_FEF_id5+T0_R_6a_id6+T0_L_LIPv_id2) / 6)

# FreeSurfer data
read_fs_metrics <- function(filename){
  df <- read_csv(filename) %>%
    rename(pseudonym=SubjID) %>%
    mutate(TimepointNr = str_sub(pseudonym, start=26,end=27),
           TimepointNr = if_else(TimepointNr=='t1',0,2),
           pseudonym = str_sub(pseudonym, start=1, end=24)) %>%
    relocate(pseudonym, TimepointNr)
  
  df
}
fs_exclusions <- c('sub-POMU56F70F8137CF0C55', 'sub-POMU6059DC1B31E11124')
volume_rois <- c('thal','caud','put','pal','accumb','amyg')
# surface_rois <- c('inferiorparietal','superiorparietal','precuneus','supramarginal',
          # 'paracentral','postcentral','precentral',
          # 'superiorfrontal','caudalmiddlefrontal','rostralanteriorcingulate')
surface_rois <- c( "caudalanteriorcingulate","caudalmiddlefrontal",
                   "cuneus","entorhinal","fusiform","inferiorparietal",
                   "inferiortemporal","isthmuscingulate","lateraloccipital",
                   "lateralorbitofrontal","lingual","medialorbitofrontal",
                   "middletemporal","parahippocampal","paracentral",
                   "parsopercularis","parsorbitalis","parstriangularis",
                   "pericalcarine","postcentral","posteriorcingulate",
                   "precentral","precuneus","rostralanteriorcingulate",
                   "rostralmiddlefrontal","superiorfrontal","superiorparietal",
                   "superiortemporal","supramarginal","frontalpole",
                   "temporalpole","transversetemporal","insula")
df_fs_vol <- read_fs_metrics('/project/3022026.01/pep/bids/derivatives/freesurfer_v7.3.2/outputs/metrics/measures/LandRvolumes.csv') %>%
  select(pseudonym,TimepointNr,TSCGMV,TCGMV,TGMV,ICV,contains(volume_rois))
df_fs_surf <- read_fs_metrics('/project/3022026.01/pep/bids/derivatives/freesurfer_v7.3.2/outputs/metrics/measures/CorticalMeasuresENIGMA_SurfAvg.csv') %>%
  select(pseudonym,TimepointNr,contains(surface_rois))
df_fs_thick <- read_fs_metrics('/project/3022026.01/pep/bids/derivatives/freesurfer_v7.3.2/outputs/metrics/measures/CorticalMeasuresENIGMA_ThickAvg.csv') %>%
  select(pseudonym,TimepointNr,contains(surface_rois))
metric <- c('fsl_MD','TensorFWCorrected_dcmp_MD','TensorDTINoNeg_dcmp_MD','dipy-MDc','dipy-MD','FW')
metric <- metric[1]
df_fs_md <- read_csv(paste0('/project/3022026.01/pep/bids/derivatives/qsiprep/measures/1_NoPVC_projfrac02-08-01_correctb0orientation//SurfaceMeasures_',metric,'_T1w_Mean.csv')) %>%
  rename(pseudonym = SubjID) %>%
  mutate(TimepointNr = if_else(str_detect(Timepoint,'Visit1'),0,2)) %>%
  relocate(pseudonym, TimepointNr) %>%
  select(-Timepoint) %>% select(pseudonym,TimepointNr,contains(surface_rois)) %>%
  na.omit()
df_fs <- df_fs_vol %>% 
  left_join(df_fs_surf, by = c('pseudonym','TimepointNr')) %>%
  left_join(df_fs_thick, by = c('pseudonym','TimepointNr')) %>% 
  left_join(df_fs_md, by = c('pseudonym', 'TimepointNr')) %>%
  filter(! pseudonym %in% fs_exclusions)
for(roi in volume_rois){
  bi_roi <- tibble(
    roi=(
      (df_fs %>% pull(paste0('L',roi)) + df_fs %>% pull(paste0('R',roi))) / 2
    )
  )
  colnames(bi_roi) <- paste0('bi_',roi)
  df_fs <- df_fs %>% bind_cols(bi_roi)
}
for(roi in surface_rois){
  for(m in c('surfavg','thickavg','dti')){
    bi_roi <- tibble(
      roi=(
        (df_fs %>% pull(paste('L',roi,m,sep='_')) + df_fs %>% pull(paste('R',roi,m,sep='_'))) / 2
      )
    )
    colnames(bi_roi) <- paste('bi',roi,m,sep='_')
    df_fs <- df_fs %>% bind_cols(bi_roi)
  }
}

# Clinical data
read_clinical_metrics <- function(filename, clinical_metrics){
  df <- read_csv(filename,
                 col_select = all_of(clinical_metrics)) %>%
    filter(ParticipantType != 'PD_PIT') %>%
    mutate(ParticipantType = if_else(ParticipantType == 'PD_POM','Patient','Healthy'),
           Subtype_DiagEx3_DisDurSplit = if_else(ParticipantType=='Healthy','0_Healthy',Subtype_DiagEx3_DisDurSplit),
           Subtype_DiagEx3_DisDurSplit.MCI_z = if_else(ParticipantType=='Healthy','0_Healthy',Subtype_DiagEx3_DisDurSplit.MCI_z),
           TimepointNr = if_else(ParticipantType == 'Healthy' & TimepointNr == 1, 2, TimepointNr)) %>%
    filter(TimepointNr != 1)
  subtypes <- df %>%
    filter(TimepointNr == 0) %>%
    select(pseudonym,
           Subtype_DiagEx3_DisDurSplit,
           Subtype_DiagEx3_DisDurSplit.MCI_z) %>%
    rename(Subtype_DiagEx3_DisDurSplit_ba = Subtype_DiagEx3_DisDurSplit,
           Subtype_DiagEx3_DisDurSplit.MCI_z_ba = Subtype_DiagEx3_DisDurSplit.MCI_z)
  df <- df %>%
    left_join(subtypes, by = 'pseudonym')
  df$Subtype_DiagEx3_DisDurSplit[is.na(df$Subtype_DiagEx3_DisDurSplit)] <- '4_Undefined'
  df$Subtype_DiagEx3_DisDurSplit.MCI_z[is.na(df$Subtype_DiagEx3_DisDurSplit.MCI_z)] <- '4_Undefined'
  df$Subtype_DiagEx3_DisDurSplit_ba[is.na(df$Subtype_DiagEx3_DisDurSplit_ba)] <- '4_Undefined'
  df$Subtype_DiagEx3_DisDurSplit.MCI_z_ba[is.na(df$Subtype_DiagEx3_DisDurSplit.MCI_z_ba)] <- '4_Undefined'
  
  df     
}
clinical_metrics <- c('pseudonym','ParticipantType','TimepointNr','YearsToFollowUp','MriNeuroPsychTask',
                      'Subtype_DiagEx3_DisDurSplit','Subtype_DiagEx3_DisDurSplit.MCI_z','Age','Gender','YearsSinceDiag','NpsEducYears',
                      'Up3OfTotal','Up3OfBradySum','Up3OfRigiditySum', 'Up3OfRestTremAmpSum', 'Up3OfCompositeTremorSum',
                      'Up3OnTotal','Up3OnBradySum','Up3OnRigiditySum', 'Up3OnRestTremAmpSum', 'Up3OnCompositeTremorSum',
                      'z_MoCA__total', 'z_CognitiveComposite2','MotorComposite','RBDSQSum','SCOPA_AUTSum','LEDD',
                      'AsymmetryIndexRiLe.WeightedBradyRig','AsymmetryIndexRiLe.WeightedBradyRig2')
df_clin <- read_clinical_metrics('/project/3022026.01/pep/ClinVars_10-08-2023/derivatives/merged_manipulated_2023-10-18.csv', 
                                 clinical_metrics) %>%
    mutate(MostAffSide=NA,
           MostAffSide = case_when(
               AsymmetryIndexRiLe.WeightedBradyRig2<0 ~ 'L',
               AsymmetryIndexRiLe.WeightedBradyRig2>0 ~ 'R',
               AsymmetryIndexRiLe.WeightedBradyRig2==0 ~ 'Sym',
               .default = NA
               )
           ) %>%
  relocate(Subtype_DiagEx3_DisDurSplit, .after = 'Gender') %>%
  relocate(Subtype_DiagEx3_DisDurSplit_ba, .after = 'Gender') %>%
  relocate(Subtype_DiagEx3_DisDurSplit.MCI_z, .after = 'Subtype_DiagEx3_DisDurSplit') %>%
  relocate(Subtype_DiagEx3_DisDurSplit.MCI_z_ba, .after = 'Subtype_DiagEx3_DisDurSplit') %>%
  relocate(MostAffSide, .after = 'Gender')

# Combine
df <- df_clin %>%
  full_join(., df_dti, by = c('pseudonym','ParticipantType','TimepointNr')
            ) %>%
  full_join(., df_fs, by = c('pseudonym','TimepointNr')
            ) #%>%
  #left_join(., df_fmri, by = c('pseudonym','ParticipantType','TimepointNr'))
```

## Feature engineering

We generate a number of new variables:

 - Compute signal-to-noise and contrast-to-noise ratios for SN FW.
 - Compute asymmetry indices of SN FW.
 - Determine most affected side from motor symptom asymmetry.
 - Compute MD averages across multiple regions.
 - Normalize regional MD by overall MD.
 - Calculate deltas.
 - Factorize categorical variables.
 - Determine completeness of data per participant.
 

```{r, echo=F,warning=F,message=F}

# Contrast-to-noise and signal-to-noise ratios for pSN FW
df <- df %>%
  mutate(pSN_cr = (pSN_avg - aSN_avg) / aSN_sd,
         L_pSN_cr = (L_pSN_avg - L_aSN_avg) / L_aSN_sd,
         R_pSN_cr = (R_pSN_avg - R_aSN_avg) / R_aSN_sd,
         pSN_sr = (pSN_avg / aSN_avg) * 100,
         L_pSN_sr = (L_pSN_avg / L_aSN_avg) * 100,
         R_pSN_sr = (R_pSN_avg / R_aSN_avg) * 100)

# Asymmetry indices for pSN FW
df <- df %>%
  mutate(AsymIdxAbs_pSN_avg = abs(R_pSN_avg - L_pSN_avg) / (R_pSN_avg + L_pSN_avg),
         AsymIdxNoAbs_pSN_avg = (R_pSN_avg - L_pSN_avg) / (R_pSN_avg + L_pSN_avg),
         AsymIdxAbs_pSN_cr = abs(R_pSN_cr - L_pSN_cr) / (R_pSN_cr + L_pSN_cr),
         AsymIdxNoAbs_pSN_cr = (R_pSN_cr - L_pSN_cr) / (R_pSN_cr + L_pSN_cr),
         AsymIdxAbs_pSN_sr = abs(R_pSN_sr - L_pSN_sr) / (R_pSN_cr + L_pSN_sr),
         AsymIdxNoAbs_pSN_sr = (R_pSN_sr - L_pSN_sr) / (R_pSN_cr + L_pSN_sr))

# Determine imaging metrics for most and least affected sides
# Remember that left is right and right is left
df <- df %>%
  mutate(
    MAF_pSN_avg = case_when(
      MostAffSide == 'R' ~ L_pSN_avg,
      MostAffSide == 'L' ~ R_pSN_avg,
      .default = NA),
    LAF_pSN_avg = case_when(
      MostAffSide == 'R' ~ R_pSN_avg,
      MostAffSide == 'L' ~ L_pSN_avg,
      .default = NA),
    MAF_pSN_cr = case_when(
      MostAffSide == 'R' ~ L_pSN_cr,
      MostAffSide == 'L' ~ R_pSN_cr,
      .default = NA),
    LAF_pSN_cr = case_when(
      MostAffSide == 'R' ~ R_pSN_cr,
      MostAffSide == 'L' ~ L_pSN_cr,
      .default = NA),
    MAF_pSN_sr = case_when(
      MostAffSide == 'R' ~ L_pSN_sr,
      MostAffSide == 'L' ~ R_pSN_sr,
      .default = NA),
    LAF_pSN_sr = case_when(
      MostAffSide == 'R' ~ R_pSN_sr,
      MostAffSide == 'L' ~ L_pSN_sr,
      .default = NA)
  )

# Averages across FS parcellations for cortical MD
df <- df %>%
  mutate(bi_premotor_dti = (bi_caudalmiddlefrontal_dti + bi_superiorfrontal_dti) / 2,
         bi_parietal_dti = (bi_superiorfrontal_dti+bi_inferiorparietal_dti+bi_precuneus_dti+bi_supramarginal_dti) / 4,
         bi_sensorimotor_dti = (bi_precentral_dti + bi_postcentral_dti + bi_paracentral_dti) / 3,
         bi_allregions_dti = (bi_premotor_dti + bi_parietal_dti + bi_sensorimotor_dti) / 3)
df <- df %>%
  mutate(bi_premotor_thickavg = (bi_caudalmiddlefrontal_thickavg + bi_superiorfrontal_thickavg) / 2,
         bi_parietal_thickavg = (bi_superiorfrontal_thickavg+bi_inferiorparietal_thickavg+bi_precuneus_thickavg+bi_supramarginal_thickavg) / 4,
         bi_sensorimotor_thickavg = (bi_precentral_thickavg + bi_postcentral_thickavg + bi_paracentral_thickavg) / 3,
         bi_allregions_thickavg = (bi_premotor_thickavg + bi_parietal_thickavg + bi_sensorimotor_thickavg) / 3)
df <- df %>%
  mutate(bi_premotor_surfavg = (bi_caudalmiddlefrontal_surfavg + bi_superiorfrontal_surfavg) / 2,
         bi_parietal_surfavg = (bi_superiorfrontal_surfavg+bi_inferiorparietal_surfavg+bi_precuneus_surfavg+bi_supramarginal_surfavg) / 4,
         bi_sensorimotor_surfavg = (bi_precentral_surfavg + bi_postcentral_surfavg + bi_paracentral_surfavg) / 3,
         bi_allregions_surfavg = (bi_premotor_surfavg + bi_parietal_surfavg + bi_sensorimotor_surfavg) / 3)
# Normalizing regional MD by overall MD
labels <- c('inferiorparietal','superiorparietal','precuneus','supramarginal',
            'paracentral','postcentral','precentral',
            'superiorfrontal','caudalmiddlefrontal')
for(lab in labels){
  
  var <- paste('bi',lab,'dti',sep='_')
  newvar <- paste('bi',lab,'norm','dti',sep='_')
  tmp1 <- df %>%
    select(all_of(var))
  tmp2 <- df %>%
    select(bi_allregions_dti)
  tmp3 <- tmp1/tmp2
  colnames(tmp3) <- newvar
  
  df <- df %>% bind_cols(tmp3)
  
  
}

# Add time-specific values as separate columns
add_timepoints_as_columns <- function(dat, id_cols, timevar, var){
  
  # Separate timepoints and calculate delta
  dat2 <- dat %>%
    select(all_of(id_cols),
           all_of(timevar),
           all_of(var)) %>%
    pivot_wider(id_cols = id_cols,
                names_from = 'TimepointNr',
                names_prefix = 'T',
                values_from = all_of(var)) %>%
    mutate(T2sub0 = T2-T0) %>%
    rename()
  
  # Add variable name
  dat2 <- rename_with(
    dat2,
    ~ paste0(.x, '_', var, recycle0 = TRUE),
    starts_with("T")
  )
  
  # Join with old data
  dat <- 
    left_join(dat, dat2, by = c(id_cols))
  
  # Print
  dat

}

id_cols <- colnames(df)[1:2]
timevar <- 'TimepointNr'
vars <- colnames(df)[15:length(colnames(df))]

for(i in 1:length(vars)){
  df <- add_timepoints_as_columns(df, id_cols, timevar, vars[i])
}

# Factorize
df <- df %>%
  mutate(ParticipantType = factor(ParticipantType),
         Subtype_DiagEx3_DisDurSplit = factor(Subtype_DiagEx3_DisDurSplit),
         Subtype_DiagEx3_DisDurSplit.MCI_z = factor(Subtype_DiagEx3_DisDurSplit.MCI_z),
         Subtype_DiagEx3_DisDurSplit_ba = factor(Subtype_DiagEx3_DisDurSplit_ba),
         Subtype_DiagEx3_DisDurSplit.MCI_z_ba = factor(Subtype_DiagEx3_DisDurSplit.MCI_z_ba),
         Gender = factor(Gender),
         TimepointNr = factor(TimepointNr))

# Which subjects have complete data?
subs <- unique(df$pseudonym)
tmp <- c()
for(i in subs){
  ntp <- df %>%
    filter(pseudonym==i) %>%
    select(TimepointNr) %>%
    nrow()
  tmp <- tmp %>%
    bind_rows(tibble(pseudonym=i,ntp=ntp))
}
tmp <- tmp %>%
  mutate(complete_case = if_else(ntp > 1, 1, 0)) %>%
  select(-ntp)
df <- df %>%
  left_join(., tmp, by = 'pseudonym')
  
```

## Exclusions

 - Select only participants who performed the motor task.
 - Exclude additional participants who did not pass quality control procedures.

```{r, echo=F,message=F}

# Only process motor task
df <- df %>%
  filter(MriNeuroPsychTask=='Motor')

# Remove 
read_dti_exclusions <- function(filename) {
  df <- read_excel(filename) %>%
    filter(retain == 'N') %>%
    select(pseudonym,session) %>%
    distinct() %>%
    mutate(TimepointNr = if_else(str_detect(session,'Visit1'),0,2),
           TimepointNr = factor(TimepointNr),
           Exclusion = 1) %>%
    select(-session)
}

exclusions <- read_dti_exclusions('/project/3024006.02/Analyses/MJF_FreeWater/QC/FW/Exclusions_and_Outliers_FW.xlsx')
df <- df %>%
  left_join(exclusions, by = c('pseudonym','TimepointNr')) %>%
  mutate(Exclusion = if_else(is.na(Exclusion),0,Exclusion)) %>%
  rename(Exclusion_FW = Exclusion)
df_fw <- df %>%
  filter(Exclusion_FW!=1)

exclusions <- read_dti_exclusions('/project/3024006.02/Analyses/MJF_FreeWater/QC/MD/Exclusions_and_Outliers_MD.xlsx')
df <- df %>%
  left_join(exclusions, by = c('pseudonym','TimepointNr')) %>%
  mutate(Exclusion = if_else(is.na(Exclusion),0,Exclusion)) %>%
  rename(Exclusion_MD = Exclusion)
df_md <- df %>%
  filter(Exclusion_MD!=1)

```

## Descriptive statistics

```{r, echo=F,warning=F,message=F}
summary_stats_table <- function(dat, vars_group, vars_dv){
  
  # Select data
  dat <- 
    dat %>%
    select(all_of(vars_group), all_of(vars_dv))
  
  # Group data iteratively
  for(i in 1:length(vars_group)){
    dat <- 
      dat %>%
      group_by(.data[[vars_group[i]]], .add = T)
  }
  
  # Initialize data frame
  stats <-
    tibble(
      n = numeric(),
      avg = numeric(),
      sd = numeric(),
      se = numeric(),
      ci.lwr = numeric(),
      ci.upr = numeric(),
      dv = character()
    )
  # Add descriptives for every dv
  for(i in 1:length(vars_dv)){
    row <- 
      dat %>%
      summarise(n=n(),
                avg = mean(.data[[vars_dv[i]]],na.rm=T),
                sd = sd(.data[[vars_dv[i]]],na.rm=T),
                se = sd / sqrt(n), 
                ci.lwr = avg - 1.96*se,
                ci.upr = avg + 1.96*se) %>%
      ungroup() %>%
      mutate(dv = vars_dv[i])
    
    stats <- 
      stats %>%
      bind_rows(row)

  }
  
  # Print
  stats
  
}
```

### Free water

```{r, echo=F,warning=F,message=F}

vars_group <- c('ParticipantType','TimepointNr')
vars_dv <- c('pSN_avg', 'aSN_avg')

# Values per timepoint with error, separated by group
stats_tab <- 
  summary_stats_table(df_fw, vars_group, vars_dv)
g <- ggplot(stats_tab, aes(x = avg, y = dv, color = ParticipantType)) +
  geom_point() + 
  geom_errorbarh(aes(xmin = ci.lwr, xmax = ci.upr)) + 
  facet_wrap(~ParticipantType+TimepointNr,ncol=2,nrow=2) +
  theme_bw() + 
  theme(legend.position = 'none')
print(g)

# Main plot: Deltas per region with error
vars_dv_delta <- paste0('T2sub0_', vars_dv)
stats_delta_tab <- 
  summary_stats_table(df_fw %>% filter(TimepointNr==0), vars_group, vars_dv_delta) %>%
  mutate(dv = str_replace(dv,'T2sub0_',''))

g <- ggplot(stats_delta_tab, aes(x = avg, y = dv, color = ParticipantType)) + 
  geom_vline(xintercept = 0, lty = 2) +
  geom_point() + 
  geom_errorbarh(aes(xmin = ci.lwr, xmax = ci.upr)) + 
  facet_grid(~ParticipantType) +
  theme_bw() +
  theme(legend.position = 'none') + 
  ylab('Region') + 
  xlab('Two-year change (delta; Follow-up - Baseline)') + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  ggtitle('Longitudinal change in free water')
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/fw_errorbar_deltas.png', dpi=300, device='png', base_width = 8)

stats_delta_tab.subtype <- 
  summary_stats_table(df_fw %>% filter(TimepointNr==0), c('Subtype_DiagEx3_DisDurSplit.MCI_z_ba','TimepointNr'), vars_dv_delta) %>%
  mutate(dv = str_replace(dv,'T2sub0_','')) %>%
  filter(Subtype_DiagEx3_DisDurSplit.MCI_z_ba != '4_Undefined')

g <- ggplot(stats_delta_tab.subtype, aes(x = avg, y = dv, color = Subtype_DiagEx3_DisDurSplit.MCI_z_ba)) + 
  geom_vline(xintercept = 0, lty = 2) +
  geom_point() + 
  geom_errorbarh(aes(xmin = ci.lwr, xmax = ci.upr)) + 
  facet_grid(~Subtype_DiagEx3_DisDurSplit.MCI_z_ba) +
  theme_bw() +
  theme(legend.position = 'none') + 
  ylab('Region') + 
  xlab('Two-year change (delta; Follow-up - Baseline)') + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  ggtitle('Longitudinal change in free water')
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/fw_errorbar_deltas_subtypes.png', dpi=300, device='png', base_width = 8)

# Main plot: Boxplots to visualize group differences
  # PD
df.plot <- df_fw %>% 
  select(pseudonym,ParticipantType,TimepointNr,
         all_of(vars_dv)) %>%
  pivot_longer(cols = all_of(vars_dv),
               names_to = 'ROI',
               values_to = 'Value') %>%
  mutate(ROI=factor(ROI,
                    levels=vars_dv,
                    labels=vars_dv),
         ParticipantType = factor(ParticipantType),
         TimepointNr = factor(TimepointNr)) %>%
  na.omit()

g <- ggplot(df.plot, aes(y=Value, x=ParticipantType, fill=TimepointNr)) + 
  geom_boxplot() + 
  facet_grid(~ROI) +
  theme_bw() +
  ylab('Free water (fraction)') + xlab('') +
  scale_fill_viridis_d(option='mako', begin = .7, end = .5,
                        direction = 1, labels=c('BA','2Y'),
                       guide = guide_legend(title='Time'))
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/fw_box_HcVsPd.png', dpi=300, device='png', base_width = 7)

  # Subtypes
df.plot.sub <- df_fw %>% 
  select(pseudonym,Subtype_DiagEx3_DisDurSplit.MCI_z_ba,TimepointNr,
         all_of(vars_dv)) %>%
  pivot_longer(cols = all_of(vars_dv),
               names_to = 'ROI',
               values_to = 'Value') %>%
  mutate(ROI=factor(ROI,
                    levels=vars_dv,
                    labels=vars_dv),
         Subtype_DiagEx3_DisDurSplit.MCI_z_ba = factor(Subtype_DiagEx3_DisDurSplit.MCI_z_ba),
         TimepointNr = factor(TimepointNr)) %>%
  na.omit() %>%
  filter(Subtype_DiagEx3_DisDurSplit.MCI_z_ba != '4_Undefined')

g <- ggplot(df.plot.sub, aes(y=Value, x=Subtype_DiagEx3_DisDurSplit.MCI_z_ba, fill=TimepointNr)) + 
  geom_boxplot() + 
  facet_grid(~ROI) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) +
  ylab('Free water (fraction)') + xlab('') +
  scale_fill_viridis_d(option='mako', begin = .7, end = .5,
                        direction = 1, labels=c('BA','2Y'),
                       guide = guide_legend(title='Time'))
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/fw_box_Subtypes.png', dpi=300, device='png', base_width = 7)

# Count of participants
df.plot %>%
  filter(ROI == 'pSN_avg') %>%
  count(ParticipantType,TimepointNr) %>%
  print()

# Main plot: Individual trajectories
delta_vals <- df_fw %>% 
  select(pseudonym,ParticipantType,TimepointNr,
         all_of(vars_dv_delta)) %>%
  pivot_longer(cols = all_of(vars_dv_delta),
               names_to = 'ROI',
               values_to = 'Delta') %>%
  mutate(ROI=factor(ROI,
                    levels=vars_dv_delta,
                    labels=vars_dv),
         ParticipantType = factor(ParticipantType),
         TimepointNr = factor(TimepointNr))
df.plot <- left_join(df.plot, delta_vals, by = c('pseudonym','ParticipantType','TimepointNr','ROI'))

g <- df.plot %>%
  filter(ROI == 'pSN_avg') %>%
  ggplot(., aes(y=Value, x = as.numeric(TimepointNr))) + 
  geom_point(alpha=.3) + 
  geom_line(aes(group = pseudonym, color = Delta), alpha=.3) + 
  geom_smooth(method = 'lm', color = 'black') +
  facet_wrap(~ROI+ParticipantType, ncol = 2, nrow = 1) +
  scale_color_viridis_c(option = 'inferno', begin=0.0,end=0.8)
print(g)

# Main plot: Boxplot of deltas
g <- df.plot %>%
  filter(TimepointNr==0) %>%
  ggplot(aes(y=Delta, x=ParticipantType, fill=ROI)) + 
  geom_boxplot() + 
  facet_grid(~ROI) +
  theme_bw() +
  ylab('Free water delta (fraction)') + xlab('') +
  scale_fill_viridis_d(option='mako', begin = .9, end = .1,
                        direction = 1,
                       guide = guide_legend(title='ROI'))
print(g)

# Checking for outliers
df.plot %>% filter(ROI=='pSN_avg', ParticipantType=='Patient') %>% arrange(-Value) %>% print(n=20)
df.plot %>% filter(ROI=='pSN_avg', ParticipantType=='Patient') %>% arrange(Delta) %>% print(n=20)
df.plot %>% filter(ROI=='pSN_avg', ParticipantType=='Patient') %>% arrange(-Delta) %>% print(n=20)

df.plot %>% filter(ROI=='pSN_avg', ParticipantType=='Healthy') %>% arrange(-Value) %>% print(n=20)
df.plot %>% filter(ROI=='pSN_avg', ParticipantType=='Healthy') %>% arrange(Delta) %>% print(n=20)
df.plot %>% filter(ROI=='pSN_avg', ParticipantType=='Healthy') %>% arrange(-Delta) %>% print(n=20)
```

### FreeSurfer metrics

```{r, echo=F,warning=F,message=F}

vars_group <- c('ParticipantType','TimepointNr')
labels <- c('inferiorparietal','superiorparietal','precuneus','supramarginal',
            'paracentral','postcentral','precentral','superiorfrontal','caudalmiddlefrontal')
measure <- c('dti')
hemi <- c('bi')
vars_dv <- paste(hemi,labels,measure,sep='_')

# Values per timepoint with error, separated by group
stats_tab <- 
  summary_stats_table(df_md, vars_group, vars_dv)
g <- ggplot(stats_tab, aes(x = avg, y = dv, color = ParticipantType)) +
  geom_point() + 
  geom_errorbarh(aes(xmin = ci.lwr, xmax = ci.upr)) + 
  facet_wrap(~ParticipantType+TimepointNr,ncol=2,nrow=2) +
  theme_bw() + 
  theme(legend.position = 'none')
print(g)

# Main plot: Deltas per region with error
vars_dv_delta <- paste0('T2sub0_', vars_dv)
stats_delta_tab <- 
  summary_stats_table(df_md %>% filter(TimepointNr==0), vars_group, vars_dv_delta) %>%
  mutate(dv = str_replace(dv,'T2sub0_',''))

g <- ggplot(stats_delta_tab, aes(x = avg, y = dv, color = ParticipantType)) + 
  geom_vline(xintercept = 0, lty = 2) +
  geom_point() + 
  geom_errorbarh(aes(xmin = ci.lwr, xmax = ci.upr)) + 
  facet_grid(~ParticipantType) +
  theme_bw() +
  theme(legend.position = 'none') + 
  ylab('Region') + 
  xlab('Two-year change (delta; Follow-up - Baseline)') + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  ggtitle('Longitudinal change in cortical MD')
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/cortMD_errorbar_deltas.png', dpi=300, device='png', base_width = 8)

stats_delta_tab.subtype <- 
  summary_stats_table(df_md %>% filter(TimepointNr==0), c('Subtype_DiagEx3_DisDurSplit.MCI_z_ba','TimepointNr'), vars_dv_delta) %>%
  mutate(dv = str_replace(dv,'T2sub0_','')) %>%
  filter(Subtype_DiagEx3_DisDurSplit.MCI_z_ba != '4_Undefined')

g <- ggplot(stats_delta_tab.subtype, aes(x = avg, y = dv, color = Subtype_DiagEx3_DisDurSplit.MCI_z_ba)) + 
  geom_vline(xintercept = 0, lty = 2) +
  geom_point() + 
  geom_errorbarh(aes(xmin = ci.lwr, xmax = ci.upr)) + 
  facet_grid(~Subtype_DiagEx3_DisDurSplit.MCI_z_ba) +
  theme_bw() +
  theme(legend.position = 'none') + 
  ylab('Region') + 
  xlab('Two-year change (delta; Follow-up - Baseline)') + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  ggtitle('Longitudinal change in cortical MD')
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/cortMD_errorbar_deltas_subtypes.png', dpi=300, device='png', base_width = 8)

# Main plot: Boxplots to visualize group differences
  # PD
df.plot <- df_md %>% 
  select(pseudonym,ParticipantType,TimepointNr,
         all_of(vars_dv)) %>%
  pivot_longer(cols = all_of(vars_dv),
               names_to = 'ROI',
               values_to = 'Value') %>%
  mutate(ROI=factor(ROI,
                    levels=vars_dv,
                    labels=vars_dv),
         ParticipantType = factor(ParticipantType),
         TimepointNr = factor(TimepointNr)) %>%
  na.omit()

g <- ggplot(df.plot, aes(y=Value, x=ParticipantType, fill=TimepointNr)) + 
  geom_boxplot() + 
  facet_grid(~ROI) +
  theme_bw() +
  ylab('Cortical MD') + xlab('') +
  scale_fill_viridis_d(option='mako', begin = .7, end = .5,
                        direction = 1, labels=c('BA','2Y'),
                       guide = guide_legend(title='Time'))
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/cortMD_box_HcVsPd.png', dpi=300, device='png', base_width = 7)

  # Subtypes
df.plot.sub <- df_md %>% 
  select(pseudonym,Subtype_DiagEx3_DisDurSplit.MCI_z_ba,TimepointNr,
         all_of(vars_dv)) %>%
  pivot_longer(cols = all_of(vars_dv),
               names_to = 'ROI',
               values_to = 'Value') %>%
  mutate(ROI=factor(ROI,
                    levels=vars_dv,
                    labels=vars_dv),
         Subtype_DiagEx3_DisDurSplit.MCI_z_ba = factor(Subtype_DiagEx3_DisDurSplit.MCI_z_ba),
         TimepointNr = factor(TimepointNr)) %>%
  na.omit() %>%
  filter(Subtype_DiagEx3_DisDurSplit.MCI_z_ba != '4_Undefined')

g <- ggplot(df.plot.sub, aes(y=Value, x=Subtype_DiagEx3_DisDurSplit.MCI_z_ba, fill=TimepointNr)) + 
  geom_boxplot() + 
  facet_grid(~ROI) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1)) +
  ylab('Mean diffusivity') + xlab('') +
  scale_fill_viridis_d(option='mako', begin = .7, end = .5,
                        direction = 1, labels=c('BA','2Y'),
                       guide = guide_legend(title='Time'))
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/cortMD_box_Subtypes.png', dpi=300, device='png', base_width = 7)

# Count of participants
df.plot %>%
  filter(ROI == 'bi_inferiorparietal_dti') %>%
  count(ParticipantType,TimepointNr) %>%
  print()

# Main plot: Individual trajectories
delta_vals <- df_md %>% 
  select(pseudonym,ParticipantType,TimepointNr,
         all_of(vars_dv_delta)) %>%
  pivot_longer(cols = all_of(vars_dv_delta),
               names_to = 'ROI',
               values_to = 'Delta') %>%
  mutate(ROI=factor(ROI,
                    levels=vars_dv_delta,
                    labels=vars_dv),
         ParticipantType = factor(ParticipantType),
         TimepointNr = factor(TimepointNr))
df.plot <- left_join(df.plot, delta_vals, by = c('pseudonym','ParticipantType','TimepointNr','ROI'))

g <- df.plot %>%
  filter(ROI == 'bi_caudalmiddlefrontal_dti') %>%
  ggplot(., aes(y=Value, x = as.numeric(TimepointNr))) + 
  geom_point(alpha=.3) + 
  geom_line(aes(group = pseudonym, color = Delta), alpha=.3) + 
  geom_smooth(method = 'lm', color = 'black') +
  facet_wrap(~ROI+ParticipantType, ncol = 2, nrow = 1) +
  scale_color_viridis_c(option = 'inferno', begin=0.0,end=0.8)
print(g)

# Main plot: Boxplot of deltas
g <- df.plot %>%
  filter(TimepointNr==0) %>%
  ggplot(aes(y=Delta, x=ParticipantType, fill=ROI)) + 
  geom_boxplot() + 
  facet_grid(~ROI) +
  theme_bw() +
  ylab('Mean diffusivity (delta)') + xlab('') +
  scale_fill_viridis_d(option='mako', begin = .9, end = .1,
                        direction = 1,
                       guide = guide_legend(title='ROI'))
print(g)

# Checking for outliers
df.plot %>% filter(ROI=='bi_inferiorparietal_dti', ParticipantType=='Patient') %>% arrange(Value) %>% print(n=20)
df.plot %>% filter(ROI=='bi_inferiorparietal_dti', ParticipantType=='Patient') %>% arrange(-Value) %>% print(n=20)
df.plot %>% filter(ROI=='bi_inferiorparietal_dti', ParticipantType=='Patient') %>% arrange(Delta) %>% print(n=20)
df.plot %>% filter(ROI=='bi_inferiorparietal_dti', ParticipantType=='Patient') %>% arrange(-Delta) %>% print(n=20)

df.plot %>% filter(ROI=='bi_inferiorparietal_dti', ParticipantType=='Healthy') %>% arrange(Value) %>% print(n=20)
df.plot %>% filter(ROI=='bi_inferiorparietal_dti', ParticipantType=='Healthy') %>% arrange(-Value) %>% print(n=20)
df.plot %>% filter(ROI=='bi_inferiorparietal_dti', ParticipantType=='Healthy') %>% arrange(Delta) %>% print(n=20)
df.plot %>% filter(ROI=='bi_inferiorparietal_dti', ParticipantType=='Healthy') %>% arrange(-Delta) %>% print(n=20)

```

## Baseline comparisons (FW)

Baseline comparisons will be omitted since the same information is contained in the longitudinal analyses below.

### Healthy vs Patient: Avg

```{r, echo=F,warning=F,message=F,eval=F}
outcome <- c('pSN_avg', 'aSN_avg')#, 'AsymIdxNoAbs_pSN_avg', 'AsymIdxNoAbs_pSN_cr', 'AsymIdxNoAbs_pSN_sr')
compare_baseline <- function(data, outcome){
  print(paste0('>>>>> ', outcome))
  f <- paste0(outcome, ' ~ ParticipantType + Age + Gender + NpsEducYears')
  m <- data %>%
    filter(TimepointNr==0) %>%
    lm(f, data = .)
  print(anova(m))
  print(summary(m))
  ggemmeans(m, 'ParticipantType',back.transform=T) %>% plot(connect.lines=T) %>% print()
}
for(i in outcome){
  compare_baseline(df_fw, i)
}
```

### Patients: MAF vs. LAF

```{r, echo=F,warning=F,message=F,eval=F}
mafdat <- df_fw %>%
    filter(TimepointNr == 0,
           ParticipantType == 'Patient') %>%
    select(pseudonym,Age,Gender,NpsEducYears,YearsSinceDiag,MAF_pSN_avg,LAF_pSN_avg) %>%
    pivot_longer(cols = c('MAF_pSN_avg','LAF_pSN_avg'),
                 names_to = 'Lateralization',
                 values_to = 'FW') %>%
  mutate(Lateralization = factor(Lateralization,
                                 levels=c('LAF_pSN_avg','MAF_pSN_avg'),
                                 labels=c('Least','Most'))) %>%
  na.omit()
mafdat %>% st(., group='Lateralization')
g <- mafdat %>% ggplot(aes(x=Lateralization,y=FW)) + 
  geom_boxplot(fill='grey') + 
  theme_bw() + 
  ylab('Free water (fraction)') + xlab('Lateralization')
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/box_LeastVsMost.png', dpi=300, device='png', base_width = 4)

m <- lmer(FW ~ Lateralization + Age + Gender + NpsEducYears + YearsSinceDiag + (1|pseudonym), data = mafdat)
anova(m) %>% print()
m %>% summary() %>% print()
ggemmeans(m, 'Lateralization',back.transform=T) %>% plot(connect.lines=T) %>% print()
```

### Subtypes: Avg

```{r, echo=F,warning=F,message=F,eval=F}
outcome <- c('pSN_avg', 'aSN_avg')#, 'AsymIdxNoAbs_pSN_avg', 'AsymIdxNoAbs_pSN_cr', 'AsymIdxNoAbs_pSN_sr')
df.subtype <- df_fw %>%
  rename(Subtype = Subtype_DiagEx3_DisDurSplit.MCI) %>%
  filter(TimepointNr==0,
         Subtype != '0_Healthy',
         Subtype != '4_Undefined') %>%
  select(pseudonym, Subtype, Age, Gender, all_of(outcome)) %>%
  pivot_longer(cols = all_of(outcome),
               names_to = 'Variable',
               values_to = 'Val')

df.subtype %>%
  ggplot(aes(y=Val, x=Subtype, fill=Variable, Color=Variable)) + 
  geom_boxplot() + 
  facet_wrap(~Variable, scales = 'free',ncol=2,nrow=2) + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  ylab('Free water (fraction)') + xlab('Subtype')

compare_baseline <- function(data, outcome){
  print(paste0('>>>>> ', outcome))
  f <- paste0('Val ~ Subtype + Age + Gender')
  m <- data %>%
    filter(Variable == outcome) %>%
    lm(f, data = .)
  print(anova(m))
  print(summary(m))
  
  em <- emmeans(m, revpairwise ~ Subtype, type='response')
  em$contrasts %>% print()
  ggemmeans(m, terms=c('Subtype'),
            back.transform = TRUE) %>% 
    plot(connect.lines=TRUE) %>% print()
}
for(i in outcome){
  compare_baseline(df.subtype, i)
}

```

### Subtypes: MAF vs LAF

```{r, echo=F,warning=F,message=F,eval=F}

df.subtype <- df_fw %>%
  rename(Subtype = Subtype_DiagEx3_DisDurSplit.MCI) %>%
  filter(TimepointNr==0,
         Subtype != '0_Healthy',
         Subtype != '4_Undefined') %>%
  select(pseudonym, Subtype, Age, Gender, starts_with('MAF'), starts_with('LAF')) %>%
  pivot_longer(cols = c(starts_with('MAF'),starts_with('LAF')),
               names_sep = '_',
               names_to = c('Side','Region','Metric'),
               values_to = 'Val') %>%
  mutate(Subtype = factor(Subtype),
         Side = factor(Side)) %>%
  filter(Metric == 'avg')

df.subtype %>%
  ggplot(aes(y=Val, x=Subtype, fill=Side, Color=Side)) + 
  geom_boxplot() + 
  facet_wrap(Region ~ Subtype + Metric, scales = 'free',
             nrow = 3, ncol = 2)

outcome <- c('pSN')
compare_baseline <- function(data, outcome){
  print(paste0('>>>>> ', outcome))
  f <- paste0('Val ~ Subtype*Side + Age + Gender')
  m <- data %>%
    filter(Region == outcome) %>%
    lm(f, data = .)
  print(anova(m))
  print(summary(m))
  
  em <- emmeans(m, ~ Subtype|Side, type='response')
  contrast(em, interaction = c('revpairwise','revpairwise'), adjust='none', type='response', by = NULL) %>% 
    kable(., 'pipe', digits = 3) %>% print()
  em <- emmeans(m, ~ Subtype|Side, type='response')
  contrast(em, interaction = c('revpairwise','revpairwise'), adjust='none', type='response') %>% 
    kable(., 'pipe', digits = 3) %>% print()
  em <- emmeans(m, revpairwise ~ Subtype, type='response')
  em$contrasts %>% print()
  em <- emmeans(m, revpairwise ~ Side, type='response')
  em$contrasts %>% print()
  ggemmeans(m, terms=c('Side','Subtype'),
            back.transform = TRUE) %>% 
    plot(connect.lines=TRUE) %>% print()
}
for(i in outcome){
  compare_baseline(df.subtype, i)
}

```

## *Longitudinal change (FW)

This section analyzes longitudinal changes in SN FW and how it differs between patients and controls. Changes in SN FW are also compared between least and most affected sides.

### *Healthy vs Patient: Avg

```{r, echo=F,warning=F,message=F}
outcome <- c('pSN_avg', 'aSN_avg')#, 'AsymIdxNoAbs_pSN_avg', 'AsymIdxNoAbs_pSN_cr', 'AsymIdxNoAbs_pSN_sr')
df.long <- df_fw %>%
  select(pseudonym, ParticipantType, TimepointNr, Age, Gender, NpsEducYears, all_of(outcome)) %>%
  pivot_longer(cols = all_of(outcome),
               names_to = 'Variable',
               values_to = 'Val')

compare_baseline <- function(data, outcome){
  print(paste0('>>>>> ', outcome))
  f <- paste0('log(Val) ~ ParticipantType*TimepointNr + Age + Gender + NpsEducYears + (1|pseudonym)')
  tmp <- data %>%
    filter(Variable==outcome) %>%
    na.omit()
  n <- tmp %>% count(ParticipantType,TimepointNr)
  print(n)
  m <- lmer(f, data = tmp)
  tab_model(m,show.se = T, digits=5)
  a <- Anova(m,type=3)
  print(a)
  # eta_squared(m) %>% print()
  
  em <- emmeans(m, ~ TimepointNr|ParticipantType, type='response')
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response', by=NULL) %>% 
    kable(., 'pipe', digits = 3, padding=0) %>% print()
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response') %>% 
    kable(., 'pipe', digits = 3, padding=0) %>% print()
  # ggemmeans(m, terms=c('TimepointNr','ParticipantType'),
  #           back.transform = TRUE) %>% 
  #   plot(connect.lines=TRUE) %>% plot()
}
for(i in outcome){
  compare_baseline(df.long, i)
}

g <- df.long %>%
        filter(Variable=='pSN_avg') %>%
  ggplot(aes(x=ParticipantType, y = Val, color=TimepointNr)) +
  geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
  geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
  stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
               mapping = aes(group=TimepointNr), position = position_dodge(width=0.8)) +
  theme_bw() + 
  scale_x_discrete(labels=c('Control','PD')) +
  scale_color_viridis_d(option='mako', begin = .1, end = .6,
                        direction = -1) + 
  ggtitle('Posterior substantia nigra') +
  ylab('Free water (fraction)') + xlab('') + 
  guides(color = guide_legend(title='Year')) +
  ylim(0.1, 0.5) + 
  annotate(geom='text',label='N=50',x=0.8,y=0.1) +
  annotate(geom='text',label='N=45',x=1.2,y=0.1) +
  annotate(geom='text',label='N=347',x=1.8,y=0.1) +
  annotate(geom='text',label='N=264',x=2.2,y=0.1)
y <- 0.47
d = 0.015
segmap <- data.frame(
        x =    c(1.00, 0.80,  1.80, 1.00, 2.00, 0.80, 1.20, 1.80, 2.20, 1.80),
        xend = c(2.00, 1.20,  2.20, 1.00, 2.00, 0.80, 1.20, 1.80, 2.20, 2.20),
        y =    c(y,    y-d,   y-d,  y,    y,    y-d*1,y-d*1,y-d*1,y-d*1, y-d*3),
        yend = c(y,    y-d,   y-d,  y-d*1,y-d*1,y-d*2,y-d*2,y-d*2,y-d*2, y-d*3)
)
sigmap <- data.frame(
        x =     c(1.50, 2.00),
        y =     c(y+0.015,    y-d*3),
        label = c('+', '**')
)
g <- g + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     size = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)
print(g)
save_plot(g, filename = '/home/sysneu/marjoh/Visualization/R_LongitudinalComparisonFmri/SnFW/FW_pSN_HCvsPD.png', dpi=300, device='png', base_width = 5)

# lmer_mod <- 
#   linear_reg() %>%
#   set_engine('lmer') %>%
#   set_mode('regression')
# 
# lmer_rec <- 
#   recipe(pSN_avg ~ ParticipantType + TimepointNr + Age + Gender + pseudonym, data = df) %>%
#   add_role(pseudonym, new_role = 'exp_unit') %>%
#   step_log(pSN_avg, base = 10) %>%
#   step_center(Age) %>%
#   step_interact(terms = ~ ParticipantType:TimepointNr) %>%
#   step_zv(all_predictors(), -has_role('exp_unit'))
# 
# lmer_wflow <- 
#   workflow() %>%
#   add_model(lmer_mod, formula = pSN_avg ~ ParticipantType*TimepointNr + Age + Gender + (1|pseudonym)) %>%
#   add_recipe(lmer_rec)
# 
# lmer_fit <- 
#   lmer_wflow %>%
#   fit(data = df)
# 
# lmer_res <- 
#   tidy(lmer_fit, conf.int = T)
```

### Healthy controls: Avg

```{r, echo=F,warning=F,message=F,eval=F}
m <- df_fw %>%
    filter(ParticipantType=='Healthy') %>%
    lmer(pSN_avg ~ TimepointNr + Age + Gender + (1|pseudonym), data = .)
anova(m) %>% print()
ggemmeans(m, 'TimepointNr') %>% plot(connect.lines=T) %>% print()

df_fw[,c(2:3,12:33)] %>%
    filter(ParticipantType=='Healthy') %>%
    select(-c(ParticipantType)) %>%
    st(., group='TimepointNr',
       group.long = F,
       group.test = T)

```

### Patients: Avg

```{r, echo=F,warning=F,message=F,eval=F}
m <- df_fw %>%
    filter(ParticipantType=='Patient') %>%
    lmer(pSN_avg ~ TimepointNr + Age + Gender + NpsEducYears + YearsSinceDiag + (1|pseudonym), data = .)
anova(m) %>% print()
ggemmeans(m, 'TimepointNr') %>% plot(connect.lines=T) %>% print()

df.pairwise <- df_fw %>%
  filter(ParticipantType=='Patient') %>%
  select(pseudonym,TimepointNr,pSN_avg) %>%
  pivot_wider(id_cols = 'pseudonym',
              names_from = 'TimepointNr',
              names_prefix = 'T_',
              values_from = 'pSN_avg') %>%
  na.omit() %>%
  pivot_longer(cols = c('T_0','T_2'),
               names_to = 'TimepointNr',
               values_to = 'pSN_avg') %>%
  mutate(TimepointNr = as.factor(TimepointNr))
df.pairwise %>%
  group_by(TimepointNr) %>%
  summarise(avg = mean(pSN_avg),
            sd = sd(pSN_avg))
pt <- df.pairwise %>%
  pairwise_t_test(pSN_avg ~ TimepointNr, paired = T)
print(pt)
pwt <- df.pairwise %>%
  pairwise_wilcox_test(pSN_avg ~ TimepointNr, paired = T)
print(pwt)

df.pairwise %>%
  lmer(pSN_avg ~ TimepointNr + (1|pseudonym), data = .) %>%
  anova()

df_fw[,c(2:3,13:34)] %>%
    filter(ParticipantType=='Patient') %>%
    select(-c(ParticipantType)) %>%
    st(., group='TimepointNr',
       group.long = F,
       group.test = T)
```

### *Patients: MAF vs LAF

```{r, echo=F,warning=F,message=F}

mafdat <- df_fw %>%
    filter(ParticipantType == 'Patient') %>%
    select(pseudonym,TimepointNr,Age,Gender,NpsEducYears,YearsSinceDiag,MAF_pSN_avg,LAF_pSN_avg) %>%
    pivot_longer(cols = c('MAF_pSN_avg','LAF_pSN_avg'),
                 names_to = 'Lateralization',
                 values_to = 'FW') %>%
  mutate(Lateralization = factor(Lateralization,
                                 levels=c('LAF_pSN_avg','MAF_pSN_avg'),
                                 labels=c('Least','Most'))) %>%
  na.omit()

m <- mafdat %>%
    lmer(FW ~ TimepointNr*Lateralization + Age + Gender + NpsEducYears + YearsSinceDiag + (1+Lateralization|pseudonym), data = .)
Anova(m,type=3) %>% print()
ggemmeans(m, c('TimepointNr','Lateralization')) %>% plot(connect.lines=T) %>% print()

g <- mafdat %>%
  ggplot(aes(x=Lateralization, y = FW, color=TimepointNr)) +
  geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
  geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
  stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
               mapping = aes(group=TimepointNr), position = position_dodge(width=0.8)) +
  theme_bw() + 
  scale_x_discrete(labels=c('Least','Most')) +
  scale_color_viridis_d(option='mako', begin = .1, end = .6,
                        direction = -1) + 
  ggtitle('') +
  ylab('Free water (fraction)') + xlab('Most affected side') + ggtitle('') +
  guides(color = guide_legend(title='Year')) +
  ylim(0.05, 0.5)
y <- 0.5
d = 0.02
segmap <- data.frame(
        x =    c(1.00, 0.80,  1.80),
        xend = c(2.00, 1.20,  2.20),
        y =    c(y,    y-d,   y-d),
        yend = c(y,    y-d,   y-d)
)
sigmap <- data.frame(
        x =     c(1.50, 1.00,  2.00),
        y =     c(y,    y-d*1, y-d*1),
        label = c('***', '*','*')
)
g <- g + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     size = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/fw_final_MAFvsLAF.png', dpi=300, device='png', base_width = 5)

# At each separate visit
for(i in c(0,2)){
        print(paste0('>>> Analzying timepoint ', i))
        m <- mafdat %>%
                filter(TimepointNr==i) %>%
                lm(log(FW) ~ Lateralization + Age + Gender + NpsEducYears + YearsSinceDiag, data = .)
        Anova(m, type=3) %>% print()
        tab_model(m,show.se=T,digits = 5)
        eta_squared(a) %>% print()
}

```

### Subtypes: Avg

```{r, echo=F,warning=T,message=F,eval=F}
outcome <- c('pSN_avg')#, 'AsymIdxNoAbs_pSN_avg', 'AsymIdxNoAbs_pSN_cr', 'AsymIdxNoAbs_pSN_sr')
df.subtype <- df_fw %>%
  rename(Subtype = Subtype_DiagEx3_DisDurSplit.MCI_z_ba) %>%
  filter(Subtype != '4_Undefined',
         Subtype != '0_Healthy') %>%
  select(pseudonym, Subtype, TimepointNr, Age, Gender, YearsSinceDiag, all_of(outcome)) %>%
  pivot_longer(cols = all_of(outcome),
               names_to = 'Variable',
               values_to = 'Val')

df.subtype %>%
  ggplot(aes(y=Val, x=TimepointNr, fill=Subtype, Color=Subtype)) + 
  geom_boxplot() + 
  facet_wrap(~Variable, scales = 'free')

compare_baseline <- function(data, outcome){
  print(paste0('>>>>> ', outcome))
  f <- paste0('Val ~ Subtype*TimepointNr + Age + Gender + (1|pseudonym)')
  m <- data %>%
    filter(Variable == outcome) %>%
    lmer(f, data = .)
  print(anova(m))
  print(summary(m))
  
  em <- emmeans(m, ~ TimepointNr|Subtype, type='response')
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response', by=NULL) %>% 
    kable(., 'pipe', digits = 3, padding=0) %>% print()
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response') %>% 
    kable(., 'pipe', digits = 3, padding=0) %>% print()
  ggemmeans(m, terms=c('TimepointNr','Subtype'),
            back.transform = TRUE) %>% 
    plot(connect.lines=TRUE) %>% plot()
}
for(i in outcome){
  compare_baseline(df.subtype, i)
}

```

### Subtypes: MAF vs LAF

```{r, echo=F,warning=F,message=F,eval=F}

df.subtype <- df_fw %>%
  rename(Subtype = Subtype_DiagEx3_DisDurSplit.MCI_z_ba) %>%
  filter(Subtype != '0_Healthy',
         Subtype != '4_Undefined') %>%
  select(pseudonym, Subtype, TimepointNr, Age, Gender, starts_with('MAF'), starts_with('LAF')) %>%
  pivot_longer(cols = c(starts_with('MAF'),starts_with('LAF')),
               names_sep = '_',
               names_to = c('Side','Region','Metric'),
               values_to = 'Val') %>%
  mutate(Subtype = factor(Subtype),
         Side = factor(Side)) %>%
  filter(Metric == 'avg')

df.subtype %>%
  ggplot(aes(y=Val, x=Subtype, fill=TimepointNr, Color=TimepointNr)) + 
  geom_boxplot() + 
  facet_grid(Region ~ Subtype+Side, scales = 'free')

outcome <- c('pSN')
compare_baseline <- function(data, outcome){
  print(paste0('>>>>> ', outcome))
  f <- paste0('Val ~ Subtype*TimepointNr*Side + Age + Gender + (1|pseudonym)')
  m <- data %>%
    filter(Region == outcome) %>%
    lmer(f, data = .)
  print(anova(m))
  print(summary(m))
  
  em <- emmeans(m, ~ Side*TimepointNr|Subtype, type='response')
  contrast(em, interaction = c('revpairwise','revpairwise','revpairwise'),
           adjust='mvt', type='response', by=NULL) %>% 
    kable(., 'pipe', digits = 3, padding=0) %>% print()
  contrast(em, interaction = c('revpairwise','revpairwise','revpairwise'),
           adjust='mvt', type='response') %>% 
    kable(., 'pipe', digits = 3, padding=0) %>% print()
  contrast(em, interaction = c('revpairwise','identity','identity'),
           adjust='mvt', type='response') %>% 
    kable(., 'pipe', digits = 3, padding=0) %>% print()
  
  em <- emmeans(m, ~ Side|Subtype, type = 'response')
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response', by = NULL) %>% print()
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response') %>% print()
  
  em <- emmeans(m, ~ TimepointNr|Subtype, type = 'response')
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response', by = NULL) %>% print()
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response') %>% print()
  
  em <- emmeans(m, ~ TimepointNr|Side, type = 'response')
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response', by = NULL) %>% print()
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response') %>% print()
  
  ggemmeans(m, terms=c('TimepointNr','Side','Subtype'),
            back.transform = TRUE) %>% 
    plot(connect.lines=TRUE) %>% plot()
}
for(i in outcome){
  compare_baseline(df.subtype, i)
}

```

## *Clinical correlations (FW)

### *Severity: baseline and follow-up

```{r, echo=F,warning=F,message=F}
corrdat <- df_fw %>%
    filter(ParticipantType=='Patient')

# Bradykinesia severity
# Baseline
print('>>> Baseline: On-state bradykinesia vs pSN FW')
cor.test(df_fw$Up3OnBradySum[df_fw$TimepointNr==0],df_fw$pSN_avg[df_fw$TimepointNr==0], method = 'spearman') %>% print()
g <- df_fw %>%
    filter(ParticipantType=='Patient',
           TimepointNr==0) %>%
    ggplot(aes(x=Up3OnBradySum,y=pSN_avg)) + 
    geom_point() + 
    geom_smooth(method='lm')
print(g)
m <- corrdat %>%
  filter(TimepointNr==0) %>%
  lm(log(pSN_avg) ~ Up3OnBradySum + Age + Gender + NpsEducYears + YearsSinceDiag, data = .)
Anova(m,type=3) %>% print()
summary(m) %>% print()
# Follow-up
print('>>> Follow-up: On-state bradykinesia vs pSN FW')
cor.test(df_fw$Up3OnBradySum[df_fw$TimepointNr==2],df_fw$pSN_avg[df_fw$TimepointNr==2], method = 'spearman') %>% print()
g <- df_fw %>%
    filter(ParticipantType=='Patient',
           TimepointNr==2) %>%
    ggplot(aes(x=Up3OnBradySum,y=pSN_avg)) + 
    geom_point() + 
    geom_smooth(method='lm')
print(g)
m <- corrdat %>%
  filter(TimepointNr==2) %>%
  lm(log(pSN_avg) ~ Up3OnBradySum + Age + Gender + NpsEducYears + YearsSinceDiag, data = .)
Anova(m,type=3) %>% print()
summary(m) %>% print()

# Motor asymmetry
# Baseline
print('>>> Baseline: Asymmetry vs pSN FW')
cor.test(df_fw$AsymmetryIndexRiLe.WeightedBradyRig2[df_fw$TimepointNr==0],df_fw$AsymIdxNoAbs_pSN_avg[df_fw$TimepointNr==0], method = 'spearman') %>% print()
m <- corrdat %>%
  filter(TimepointNr==0) %>%
  lm(AsymIdxNoAbs_pSN_avg ~ AsymmetryIndexRiLe.WeightedBradyRig2 + Age + Gender + NpsEducYears + YearsSinceDiag, data = .)
tab_model(m, digits=5, show.se=T)
g <- ggemmeans(m, terms = 'AsymmetryIndexRiLe.WeightedBradyRig2') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + 
  ggtitle('Baseline') + xlab('Motor asymetry index') + ylab('Free water asymmetry index') +
  annotate(geom = 'text', x=-1, y=-0.5, hjust=0, label = expression(italic('N')*'= 344, '*italic('')*' = -0.071, SE = 0.022, '*italic('P')*' = 0.001')) +
  ylim(-0.5,0.5)
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/fw_final_asymmetry_ba.png', dpi=300, device='png', base_width = 5)
Anova(m,type=3) %>% print()
summary(m) %>% print()
# Follow-up
print('>>> Follow-up: Asymmetry vs pSN FW')
cor.test(df_fw$AsymmetryIndexRiLe.WeightedBradyRig2[df_fw$TimepointNr==2],df_fw$AsymIdxNoAbs_pSN_avg[df_fw$TimepointNr==2], method = 'spearman') %>% print()
m <- corrdat %>%
  filter(TimepointNr==2) %>%
  lm(AsymIdxNoAbs_pSN_avg ~ AsymmetryIndexRiLe.WeightedBradyRig2 + Age + Gender + NpsEducYears + YearsSinceDiag, data = .)
tab_model(m, digits=5, show.se=T)
g <- ggemmeans(m, terms = 'AsymmetryIndexRiLe.WeightedBradyRig2') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + 
  ggtitle('Follow-up') + xlab('Motor asymetry index') + ylab('Free water asymmetry index') +
  annotate(geom = 'text', x=-1, y=-0.5, hjust=0, label = expression(italic('N')*'= 251, '*italic('')*' = -0.100, SE = 0.032, '*italic('P')*' = 0.002')) +
  ylim(-0.5,0.5)
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/fw_final_asymmetry_fu.png', dpi=300, device='png', base_width = 5)
Anova(m,type=3) %>% print()
summary(m) %>% print()

```

### Severity: longitudinal (RM-CORR)

```{r, echo=F,warning=F,message=F,eval=F}
# Bradykinesia severity 
rmc <- rmcorr('pseudonym',Up3OfBradySum,pSN_avg,dataset = corrdat)
print(rmc)
plot(rmc)

# Motor asymmetry
rmc <- rmcorr('pseudonym',AsymmetryIndexRiLe.WeightedBradyRig2,AsymIdxNoAbs_pSN_avg,dataset = corrdat)
print(rmc)
plot(rmc)

# Tremor
# rmc <- rmcorr('pseudonym',Up3OfRestTremAmpSum,RRA_avg,dataset = corrdat)
# print(rmc)
# plot(rmc)
```

### *Severity: longitudinal (delta)

```{r, echo=F,warning=F,message=F}

deltadat <- df_fw %>%
  filter(ParticipantType=='Patient',
         TimepointNr==0)

# Bradykinesia severity
m <- lm(T2sub0_Up3OnBradySum ~ T2sub0_pSN_avg + T0_Up3OnBradySum + T0_pSN_avg + Age + Gender + NpsEducYears + YearsSinceDiag, data = deltadat)
Anova(m,type=3)
summary(m)
ggemmeans(m,c('T2sub0_pSN_avg')) %>% plot() %>% print()
g <- deltadat %>%
    ggplot(aes(x=T2sub0_Up3OnBradySum,y=T2sub0_pSN_avg)) + 
    geom_point() + 
    geom_smooth(method='lm',color='black')
plot(g)
cor.test(deltadat$T2sub0_Up3OnBradySum[deltadat$TimepointNr==0],deltadat$T2sub0_pSN_avg[deltadat$TimepointNr==0], method = 'spearman') %>% print()


# Motor asymmetry
m <- lm(T2sub0_AsymIdxNoAbs_pSN_avg ~ T2sub0_AsymmetryIndexRiLe.WeightedBradyRig2 + T0_AsymmetryIndexRiLe.WeightedBradyRig2 + T0_AsymIdxNoAbs_pSN_avg + Age + Gender + NpsEducYears + YearsSinceDiag, data = deltadat)
Anova(m,type=3)
summary(m)
ggemmeans(m,'T2sub0_AsymmetryIndexRiLe.WeightedBradyRig2') %>% plot() %>% print()
g <- deltadat %>%
    ggplot(aes(x=T2sub0_AsymmetryIndexRiLe.WeightedBradyRig2,y=T2sub0_AsymIdxNoAbs_pSN_avg)) + 
    geom_point() + 
    geom_smooth(method='lm',color='black')
plot(g)
cor.test(deltadat$T2sub0_AsymmetryIndexRiLe.WeightedBradyRig2[deltadat$TimepointNr==0],deltadat$T2sub0_pSN_avg[deltadat$TimepointNr==0], method = 'spearman') %>% print()
tab_model(m, digits=5, show.se=T)
g <- ggemmeans(m, terms = 'T2sub0_AsymmetryIndexRiLe.WeightedBradyRig2') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + 
  ggtitle('Between-session change') + xlab('Change in motor asymetry index') + ylab('Change in free water asymmetry index') +
  annotate(geom = 'text', x=-1, y=-0.5, hjust=0, label = expression(italic('N')*'= 290, '*italic('')*' = -0.010, SE = 0.038, '*italic('P')*' = 0.797')) +
  ylim(-0.5,0.5)
print(g)
save_plot(g, filename = '/project/3024006.02/Analyses/MJF_FreeWater/vis/fw_final_asymmetry_delta.png', dpi=300, device='png', base_width = 5)
```

### Severity: longitudinal (disaggregation)

```{r, echo=F,warning=F,message=F,eval=F}

library(bmlm)
complete_subjects <- df_fw %>%
  filter(ParticipantType=='Patient') %>%
  select(pseudonym,TimepointNr,Age,Gender,Up3OfBradySum,pSN_avg) %>%
  pivot_wider(id_cols = c('pseudonym','Age','Gender'),
              names_from = 'TimepointNr',
              names_prefix = 'T',
              values_from = c('Up3OfBradySum','pSN_avg')) %>%
  na.omit() %>%
  pull(pseudonym)

disaggdat <- df_fw %>%
  filter(ParticipantType=='Patient',
         pseudonym %in% complete_subjects) %>%
  isolate(.,
          by = 'pseudonym',
          value = c('Up3OfBradySum','pSN_avg'),
          which = 'both') %>%
  select(pseudonym,TimepointNr,Age,Gender,starts_with('Up3OfBradySum'),starts_with('pSN_avg'))

disaggdat %>%
  ggplot(aes(x=pSN_avg_cw,y=Up3OfBradySum_cw)) +
  geom_point() + 
  geom_line(aes(group=pseudonym)) + 
  geom_smooth(method='lm')

m <- lmer(Up3OfBradySum ~ pSN_avg_cw + pSN_avg_cb + Age + Gender + (1|pseudonym), data = disaggdat,
          control = lmerControl(optimizer = 'bobyqa', optCtrl=list(maxfun=2e5)), REML = F)
m_null <- lmer(Up3OfBradySum ~ pSN_avg_cb + Age + Gender + (1|pseudonym), data = disaggdat,
               control = lmerControl(optimizer = 'bobyqa', optCtrl=list(maxfun=2e5)), REML = F)
anova(m_null,m)
ggemmeans(m,'pSN_avg_cw') %>% plot() %>% print()

```

### *LEDD: baseline and follow-up

```{r, echo=F,warning=F,message=F}

# Baseline
print('>>> Baseline: LEDD vs pSN FW')
cor.test(df_fw$LEDD[df_fw$TimepointNr==0],df_fw$pSN_avg[df_fw$TimepointNr==0], method = 'spearman') %>% print()
g <- df_fw %>%
    filter(ParticipantType=='Patient',
           TimepointNr==0) %>%
    ggplot(aes(x=LEDD,y=pSN_avg)) + 
    geom_point() + 
    geom_smooth(method='lm')
print(g)
m <- corrdat %>%
  filter(TimepointNr==0) %>%
  lm(pSN_avg ~ LEDD + Age + Gender + NpsEducYears + YearsSinceDiag, data = .)
Anova(m,type=3) %>% print()
summary(m) %>% print()
# Follow-up
print('>>> Follow-up: LEDD vs pSN FW')
cor.test(df_fw$LEDD[df_fw$TimepointNr==2],df_fw$pSN_avg[df_fw$TimepointNr==2], method = 'spearman') %>% print()
g <- df_fw %>%
    filter(ParticipantType=='Patient',
           TimepointNr==2) %>%
    ggplot(aes(x=LEDD,y=pSN_avg)) + 
    geom_point() + 
    geom_smooth(method='lm')
print(g)
m <- corrdat %>%
  filter(TimepointNr==2) %>%
  lm(pSN_avg ~ Up3OfBradySum + Age + Gender + NpsEducYears + YearsSinceDiag, data = .)
Anova(m,type=3) %>% print()
summary(m) %>% print()

```

### LEDD: longitudinal (RM-CORR)

```{r, echo=F,warning=F,message=F,eval=F}

corrdat <- df_fw %>%
    filter(ParticipantType=='Patient')
rmc <- rmcorr('pseudonym',LEDD,pSN_avg,dataset = corrdat)
print(rmc)
plot(rmc)

```

### *LEDD: longitudinal (delta)

```{r, echo=F,warning=F,message=F}

deltadat <- df_fw %>%
  filter(ParticipantType=='Patient',
         TimepointNr==0)
m <- lm(T2sub0_LEDD ~ T2sub0_pSN_avg + T0_LEDD + T0_pSN_avg + Age + Gender + NpsEducYears + YearsSinceDiag, data = deltadat)
Anova(m,type=3)
summary(m)
ggemmeans(m,c('T2sub0_pSN_avg')) %>% plot() %>% print()
g <- deltadat %>%
    ggplot(aes(x=T2sub0_LEDD,y=T2sub0_pSN_avg)) + 
    geom_point() + 
    geom_smooth(method='lm',color='black')
plot(g)
cor.test(deltadat$T2sub0_LEDD[deltadat$TimepointNr==0],deltadat$T2sub0_pSN_avg[deltadat$TimepointNr==0], method = 'spearman') %>% print()

```

### LEDD: longitudinal (disaggregation)

```{r, echo=F,warning=F,message=F,eval=F}

library(bmlm)
complete_subjects <- df_fw %>%
  filter(ParticipantType=='Patient') %>%
  select(pseudonym,TimepointNr,Age,Gender,LEDD,pSN_avg) %>%
  pivot_wider(id_cols = c('pseudonym','Age','Gender'),
              names_from = 'TimepointNr',
              names_prefix = 'T',
              values_from = c('LEDD','pSN_avg')) %>%
  na.omit() %>%
  pull(pseudonym)

disaggdat <- df_fw %>%
  filter(ParticipantType=='Patient',
         pseudonym %in% complete_subjects) %>%
  isolate(.,
          by = 'pseudonym',
          value = c('LEDD','pSN_avg'),
          which = 'both') %>%
  select(pseudonym,TimepointNr,Age,Gender,starts_with('LEDD'),starts_with('pSN_avg'))

disaggdat %>%
  ggplot(aes(x=pSN_avg_cw,y=LEDD_cw)) +
  geom_point() + 
  geom_line(aes(group=pseudonym)) + 
  geom_smooth(method='lm')

m <- lmer(LEDD ~ pSN_avg_cw + pSN_avg_cb + Age + Gender + (1|pseudonym), data = disaggdat,
          control = lmerControl(optimizer = 'bobyqa', optCtrl=list(maxfun=2e5)), REML = F)
m_null <- lmer(LEDD ~ pSN_avg_cb + Age + Gender + (1|pseudonym), data = disaggdat,
               control = lmerControl(optimizer = 'bobyqa', optCtrl=list(maxfun=2e5)), REML = F)
anova(m_null,m)
ggemmeans(m,'pSN_avg_cw') %>% plot() %>% print()

```

## Basal ganglia ~ pSN

```{r, echo=F,warning=F,message=F, eval=F}

df.bg <- df_fw %>%
  filter(TimepointNr==0,
         ParticipantType=='Patient') %>%
  select(pseudonym,Age,Gender,MAF_pSN_avg,BG_activity,CB_activity,M1_activity,LEDD) %>%
  na.omit()

gp <- ggpairs(df.bg, columns = 4:8, aes(alpha=0.5),
        lower = list(continuous = 'smooth'))
print(gp)

m <- lm(BG_activity ~ MAF_pSN_avg + Age + Gender, data = df.bg)
anova(m) %>% print()
summary(m) %>% print()
ggemmeans(m, 'MAF_pSN_avg') %>% plot() %>% print()

```

## *Write FW values for FSL covariate analysis

This section should be run manually.

```{r, echo=F,warning=F,message=F,eval=F}

cons <- c('con_0007', 'con_0010')

for(i in cons){
  
  pth <- paste0('/project/3024006.02/Analyses/motor_task/Group/Longitudinal/FSL/data/', i, '/imgs__delta_clincorr.txt')
  pths <- read_csv(pth, col_names = c('paths')) %>%
    mutate(pseudonym = str_sub(paths, start = 72, end = 95))
  
  fsl_covs <-
    df_fw %>%
    filter(ParticipantType=='Patient') %>%
    select(pseudonym, TimepointNr, pSN_avg) %>%
    pivot_wider(id_cols = 'pseudonym',
                names_from = 'TimepointNr',
                names_prefix = 'T',
                values_from = 'pSN_avg') %>%
    mutate(pSN_Delta = T2-T0,
           across(where(is.numeric), \(x) round(x, 5))) %>%
    rename(pSN_T0 = T0) %>%
    select(-T2) %>%
    relocate(pseudonym, pSN_Delta, pSN_T0)
  
  fsl_covs <- pths %>%
    left_join(fsl_covs, by = 'pseudonym')
  
  outpth <- paste0('/project/3024006.02/Analyses/motor_task/Group/Longitudinal/FSL/data/', i, '/covs__delta_pSN_FW.csv')
  write_csv(fsl_covs, outpth)
  
  fsl_covs <-
    df_md %>%
    filter(ParticipantType=='Patient') %>%
    select(pseudonym, TimepointNr, bi_caudalmiddlefrontal_dti) %>%
    pivot_wider(id_cols = 'pseudonym',
                names_from = 'TimepointNr',
                names_prefix = 'T',
                values_from = 'bi_caudalmiddlefrontal_dti') %>%
    mutate(cmf_Delta = T2-T0,
           across(where(is.numeric), \(x) round(x, 5))) %>%
    rename(cmf_T0 = T0) %>%
    select(-T2) %>%
    relocate(pseudonym, cmf_Delta, cmf_T0)
  
  fsl_covs <- pths %>%
    left_join(fsl_covs, by = 'pseudonym')
  
  outpth <- paste0('/project/3024006.02/Analyses/motor_task/Group/Longitudinal/FSL/data/', i, '/covs__delta_caudalmiddlefrontal_MD.csv')
  write_csv(fsl_covs, outpth)
  
}

```

## *FreeSurfer metrics

```{r, echo=F,warning=F,message=F}

# Weights to match patients and controls
set.seed(312)
matching0 <- df_md %>%
  filter(TimepointNr==0) %>%
  select(pseudonym,ParticipantType,Age,Gender,NpsEducYears) %>%
  mutate(ParticipantType = if_else(ParticipantType=='Healthy',0,1)) %>%
  na.omit() %>%
  matchit(ParticipantType ~ Age + Gender + NpsEducYears,
          data = .,
          method = NULL,
          distance = 'glm',
          link = 'probit',
          estimand = 'ATT')
summary(matching0)
matching1 <- df_md %>%
  filter(TimepointNr==0) %>%
  select(pseudonym,ParticipantType,Age,Gender,NpsEducYears) %>%
  mutate(ParticipantType = if_else(ParticipantType=='Healthy',0,1)) %>%
  na.omit() %>%
  matchit(ParticipantType ~ Age + Gender + NpsEducYears,
          data = .,
          method = 'full',
          distance = 'glm',
          link = 'probit',
          estimand = 'ATT',
          tol = 1e-4)
summary(matching1, un = F)
plot(summary(matching1))
plot(matching1, type = 'jitter', interactive = F)
plot(matching1, type = 'density', interactive = F,
     which.xs = ~Age + Gender + NpsEducYears)
w <- match.data(matching1) %>%
  select(pseudonym, ParticipantType, weights) %>%
  mutate(ParticipantType = factor(if_else(ParticipantType==0,'Healthy','Patient')))
df_md <- df_md %>%
  left_join(., w, by = c('pseudonym','ParticipantType'))

# Write data for plotting purposes
write_csv(df_md,'/project/3024006.02/wd/ggseg_all_data.csv')

# Group x Time comparisons, 1 observation per timepoint

labels <- c('inferiorparietal','superiorparietal','precuneus','supramarginal',
            'paracentral','postcentral','precentral',
            'superiorfrontal','caudalmiddlefrontal')
# 'premotor','sensorimotor','parietal'
measure <- c('dti')
hemi <- c('bi')
cols <- paste(hemi,labels,measure,sep='_')

plot_dat <- df_md %>%
  select(pseudonym,ParticipantType,TimepointNr,
         all_of(cols)) %>%
  pivot_longer(cols = all_of(cols),
               names_to = 'region',
               values_to = 'metric')
plot_dat %>%
  ggplot(aes(x=factor(region),y=metric,fill=TimepointNr)) + 
  geom_boxplot() + 
  facet_wrap(~ParticipantType) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

tab <- c()
tab2 <- c()
for(lab in labels){
  
        print(paste0('>>> Region: ', lab))
        
  # Variables to use
  dv <- paste0(hemi,'_',lab, '_', measure)
  dv_delta <- paste0('T2sub0_',dv)
  dv_t0 <- paste0('T0_',dv)
  dv_t2 <- paste0('T2_',dv)
  iv_thick_ba <- paste0('T0_',str_replace(dv,'dti','thickavg'))
  iv_thick_fu <- paste0('T2_',str_replace(dv,'dti','thickavg'))
  iv_thick_delta <- paste0('T2sub0_',str_replace(dv,'dti','thickavg'))
  # iv_allregions_ba <- paste0('T0_',hemi,'_','allregions', '_', measure)
  # iv_allregions_fu <- paste0('T2_',hemi,'_','allregions', '_', measure)
  # iv_allregions_delta <- paste0('T2sub0_',hemi,'_','allregions', '_', measure)
  
  # Group comparisons
  f <- paste0('log(', dv, ') ~ ParticipantType*TimepointNr + Age + Gender + NpsEducYears + ', iv_thick_delta,'+',iv_thick_ba, ' + (1|pseudonym)')
  m <- lmer(f, df_md, weights = NULL)
  tab_model(m,show.se = T, digits=5)
  a <- Anova(m,type=3)
  print(a)
  eta_squared(m) %>% print()
  em <- emmeans(m, ~ TimepointNr|ParticipantType, type='response')
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response', by=NULL) %>% 
    kable(., 'pipe', digits = 3, padding=0) %>% print()
  contrast(em, interaction = c('revpairwise','revpairwise'),
           adjust='mvt', type='response') %>% 
    kable(., 'pipe', digits = 3, padding=0) %>% print()
  
  # g1 <- df_md %>%
  #   ggplot(aes(y=.data[[dv]],x=ParticipantType,fill=TimepointNr)) +
  #   geom_boxplot() + theme_bw()
  # g2 <- ggemmeans(m, terms = c('ParticipantType','TimepointNr')) %>% plot()
  
  # Correlations
  pcor_dat <- df_md %>%
    filter(ParticipantType=='Patient',
           TimepointNr==0,
           complete_case == 1) %>%
    rename(T0_Up3BradySum=T0_Up3OfBradySum, T2sub0_Up3BradySum=T2sub0_Up3OfBradySum) %>%
    select(pseudonym,all_of(dv_delta),all_of(dv_t0),
           T0_Up3BradySum,T2sub0_Up3BradySum,
           T0_z_MoCA__total,T2sub0_z_MoCA__total,
           T0_LEDD,T2sub0_LEDD,
           all_of(c(iv_thick_ba,iv_thick_delta)),
           Age,NpsEducYears,Gender,YearsSinceDiag) %>%
    na.omit()
  
  # Symptom-specific correlations
  m_delta <- lm(paste0(dv_delta, '~', dv_t0, 
                       '+T2sub0_Up3BradySum+T0_Up3BradySum+T2sub0_z_MoCA__total+T0_z_MoCA__total+T2sub0_LEDD+T0_LEDD+Age+Gender+NpsEducYears+YearsSinceDiag+',
                       iv_thick_delta,'+',iv_thick_ba), data = pcor_dat)
  s_delta <- summary(m_delta)
  
  # g3 <- pcor_dat %>%
  #   ggplot(aes(y=.data[[dv_delta]],x=T2sub0_Up3BradySum)) +
  #   geom_point() +
  #   geom_smooth(method = 'lm')
  # g4 <- ggemmeans(m_delta, terms='T2sub0_Up3BradySum') %>%
  #   plot(add.data = T)
  # g5 <- ggemmeans(m_delta, terms='T2sub0_z_MoCA__total') %>%
  #   plot(add.data = T)
  # g6 <- ggemmeans(m_delta, terms='T2sub0_LEDD') %>%
  #   plot(add.data = T)
  
  # Non-specific correlations
  m_delta_brady <- lm(paste0(dv_delta, '~', dv_t0, 
                       '+T2sub0_Up3BradySum+T0_Up3BradySum+Age+Gender+NpsEducYears+YearsSinceDiag+',
                       iv_thick_delta,'+',iv_thick_ba), data = pcor_dat)
  s_delta_brady <- summary(m_delta_brady)
  m_delta_moca <- lm(paste0(dv_delta, '~', dv_t0, 
                       '+T2sub0_z_MoCA__total+T0_z_MoCA__total+Age+Gender+NpsEducYears+YearsSinceDiag+',
                       iv_thick_delta,'+',iv_thick_ba), data = pcor_dat)
  s_delta_moca <- summary(m_delta_moca)
  m_delta_ledd <- lm(paste0(dv_delta, '~', dv_t0, 
                       '+T2sub0_LEDD+T0_LEDD+Age+Gender+NpsEducYears+YearsSinceDiag+',
                       iv_thick_delta,'+',iv_thick_ba), data = pcor_dat)
  s_delta_ledd <- summary(m_delta_ledd)
  
  # Time-specific correlations
    # T0
  tcor_dat_t0 <- df_md %>%
    filter(ParticipantType=='Patient',
           TimepointNr==0) %>%
    rename(T0_Up3BradySum=T0_Up3OfBradySum) %>%
    select(pseudonym,all_of(dv_t0),
           T0_Up3BradySum,
           T0_z_MoCA__total,
           all_of(c(iv_thick_ba)),
           Age,NpsEducYears,Gender,YearsSinceDiag) %>%
    na.omit()
  s_ba_all <- lm(paste0(dv_t0, '~T0_Up3BradySum+T0_z_MoCA__total+Age+Gender+NpsEducYears+YearsSinceDiag'), data = tcor_dat_t0) %>% summary()
  s_ba_brady <- lm(paste0(dv_t0, '~T0_Up3BradySum+Age+Gender+NpsEducYears+YearsSinceDiag'), data = tcor_dat_t0) %>% summary()
  s_ba_moca <- lm(paste0(dv_t0, '~T0_z_MoCA__total+Age+Gender+NpsEducYears+YearsSinceDiag'), data = tcor_dat_t0) %>% summary()
    # T2
  tcor_dat_t2 <- df_md %>%
    filter(ParticipantType=='Patient',
           TimepointNr==2) %>%
    rename(T2_Up3BradySum=T2_Up3OfBradySum) %>%
    select(pseudonym,all_of(dv_t2),
           T2_Up3BradySum,
           T2_z_MoCA__total,
           all_of(c(iv_thick_fu)),
           Age,NpsEducYears,Gender,YearsSinceDiag) %>%
    na.omit()
  s_fu_all <- lm(paste0(dv_t2, '~T2_Up3BradySum+T2_z_MoCA__total+Age+Gender+NpsEducYears+YearsSinceDiag'), data = tcor_dat_t2) %>% summary()
  s_fu_brady <- lm(paste0(dv_t2, '~T2_Up3BradySum+Age+Gender+NpsEducYears+YearsSinceDiag'), data = tcor_dat_t2) %>% summary()
  s_fu_moca <- lm(paste0(dv_t2, '~T2_z_MoCA__total+Age+Gender+NpsEducYears+YearsSinceDiag'), data = tcor_dat_t2) %>% summary()

  # g <- ggarrange(g1,g2,g3,ncol = 2, nrow = 2)
  # print(g)

  # g <- ggarrange(g4,g5,g6,ncol = 2, nrow = 2)
  # print(g)
  
  # RM correlations
  rcor_dat <- df_md %>%
    filter(ParticipantType=='Patient',
           complete_case==1) %>%
    rename(Up3BradySum=Up3OfBradySum)
  rmc_brady <- rmcorr('pseudonym',dv,'Up3BradySum',dataset = rcor_dat)
  rmc_moca <- rmcorr('pseudonym',dv,'z_MoCA__total',dataset = rcor_dat)
  rmc_ledd <- rmcorr('pseudonym',dv,'LEDD',dataset = rcor_dat)
  
  # Cohen's d
  d <- df_md %>%
    filter(TimepointNr==0) %>%
    select(all_of(dv_delta),ParticipantType) %>%
    na.omit() %>%
    cohens_d(dv_delta,'ParticipantType',data=.,pooled_sd = F)
  # Eta squared
  eta <- eta_squared(m)
  
  tab <- bind_rows(tab,
                   tibble(label=dv,
                          p_Group_x_Time=a[rownames(a)=='ParticipantType:TimepointNr',3],
                          p_Time=a[rownames(a)=='TimepointNr',3],
                          p_Group=a[rownames(a)=='ParticipantType',3],
                          cohens_d=d$Cohens_d,
                          eta_squared=eta$Eta2_partial[eta$Parameter=='ParticipantType:TimepointNr'],
                          p_brady_delta=s_delta$coefficients[rownames(s_delta$coefficients)=='T2sub0_Up3BradySum',4],
                          p_bradyOnly_delta=s_delta_brady$coefficients[rownames(s_delta_brady$coefficients)=='T2sub0_Up3BradySum',4],
                          p_brady_rmc=rmc_brady$p,
                          p_moca_delta=s_delta$coefficients[rownames(s_delta$coefficients)=='T2sub0_z_MoCA__total',4],
                          p_mocaOnly_delta=s_delta_moca$coefficients[rownames(s_delta_moca$coefficients)=='T2sub0_z_MoCA__total',4],
                          p_moca_rmc=rmc_moca$p,
                          p_ledd_delta=s_delta$coefficients[rownames(s_delta$coefficients)=='T2sub0_LEDD',4],
                          p_leddOnly_delta=s_delta_ledd$coefficients[rownames(s_delta_ledd$coefficients)=='T2sub0_LEDD',4],
                          p_ledd_rmc=rmc_ledd$p)
  )
  tab2 <- bind_rows(tab2,
                   tibble(label=dv,
                          p_all_brady_ba=s_ba_all$coefficients[rownames(s_ba_all$coefficients)=='T0_Up3BradySum',4],
                          p_ba_brady=s_ba_brady$coefficients[rownames(s_ba_brady$coefficients)=='T0_Up3BradySum',4],
                          p_all_brady_fu=s_fu_all$coefficients[rownames(s_fu_all$coefficients)=='T2_Up3BradySum',4],
                          p_fu_brady=s_fu_brady$coefficients[rownames(s_fu_brady$coefficients)=='T2_Up3BradySum',4],
                          p_all_moca_ba=s_ba_all$coefficients[rownames(s_ba_all$coefficients)=='T0_z_MoCA__total',4],
                          p_ba_moca=s_ba_moca$coefficients[rownames(s_ba_moca$coefficients)=='T0_z_MoCA__total',4],
                          p_all_moca_fu=s_fu_all$coefficients[rownames(s_fu_all$coefficients)=='T2_z_MoCA__total',4],
                          p_mocaOnly_fu=s_fu_moca$coefficients[rownames(s_fu_moca$coefficients)=='T2_z_MoCA__total',4])
  )
}

print(tab)
print(paste0('>>> FDR-corrected GROUP x TIME')) 
p.adjust(tab$p_Group_x_Time,method='fdr') %>% round(digits=5)
write_csv(tab,'/project/3024006.02/wd/ggseg_stats_CLINvsMD.csv')
print(tab2)
print(paste0('>>> FDR-corrected BRADYKINESIA')) 
p.adjust(tab2$p_all_brady_ba,method='fdr') %>% round(digits=5)

# Visualization

n <- df_md %>%
  select(ParticipantType,TimepointNr,bi_caudalmiddlefrontal_dti) %>%
  na.omit() %>%
  count(ParticipantType,TimepointNr)

## Caudal middle frontal
g <- df_md %>%
  ggplot(aes(x=ParticipantType, y = bi_caudalmiddlefrontal_dti, color=TimepointNr)) +
  geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
  geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
  stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
               mapping = aes(group=TimepointNr), position = position_dodge(width=0.8)) +
  theme_bw() + 
  scale_x_discrete(labels=c('Control','PD')) +
  scale_color_viridis_d(option='mako', begin = .1, end = .6,
                        direction = -1) + 
  ggtitle('') +
  ylab('Mean diffusivity') + xlab('') + ggtitle('Caudal middle frontal gyrus') +
  guides(color = guide_legend(title='Year')) +
  ylim(0.68, 1) + 
  annotate(geom='text',label='N=60',x=0.8,y=0.68) +
  annotate(geom='text',label='N=54',x=1.2,y=0.68) +
  annotate(geom='text',label='N=355',x=1.8,y=0.68) +
  annotate(geom='text',label='N=308',x=2.2,y=0.68)
y <- 1
d = 0.015
segmap <- data.frame(
        x =    c(1.00, 0.80,  1.80, 1.00, 2.00, 0.80, 1.20, 1.80, 2.20, 1.80, 0.80),
        xend = c(2.00, 1.20,  2.20, 1.00, 2.00, 0.80, 1.20, 1.80, 2.20, 2.20, 1.20),
        y =    c(y,    y-d,   y-d,  y,    y,    y-d*1,y-d*1,y-d*1,y-d*1, y-d*3, y-d*3),
        yend = c(y,    y-d,   y-d,  y-d*1,y-d*1,y-d*2,y-d*2,y-d*2,y-d*2, y-d*3, y-d*3)
)
sigmap <- data.frame(
        x =     c(1.50, 1.00,        2.00),
        y =     c(y,    y-d*3, y-d*3),
        label = c('***','*', '***')
)
g <- g + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     size = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)
print(g)
save_plot(g, filename = '/home/sysneu/marjoh/Visualization/R_LongitudinalComparisonFmri/CorticalMD/cMD_caudalmiddlefrontal_HCvsPD.png', dpi=300, device='png', base_width = 5)

## Inferior parietal cortex
g <- df_md %>%
  ggplot(aes(x=ParticipantType, y = bi_inferiorparietal_dti, color=TimepointNr)) +
  geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
  geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
  stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
               mapping = aes(group=TimepointNr), position = position_dodge(width=0.8)) +
  theme_bw() + 
  scale_x_discrete(labels=c('Control','PD')) +
  scale_color_viridis_d(option='mako', begin = .1, end = .6,
                        direction = -1) + 
  ggtitle('') +
  ylab('Mean diffusivity') + xlab('') + ggtitle('Inferior parietal cortex') +
  guides(color = guide_legend(title='Year')) +
  ylim(0.68, 1) + 
  annotate(geom='text',label='N=60',x=0.8,y=0.68) +
  annotate(geom='text',label='N=54',x=1.2,y=0.68) +
  annotate(geom='text',label='N=355',x=1.8,y=0.68) +
  annotate(geom='text',label='N=308',x=2.2,y=0.68)
y <- 1
d = 0.015
segmap <- data.frame(
        x =    c(1.00, 0.80,  1.80, 1.00, 2.00, 0.80, 1.20, 1.80, 2.20, 1.80, 0.80),
        xend = c(2.00, 1.20,  2.20, 1.00, 2.00, 0.80, 1.20, 1.80, 2.20, 2.20, 1.20),
        y =    c(y,    y-d,   y-d,  y,    y,    y-d*1,y-d*1,y-d*1,y-d*1, y-d*3, y-d*3),
        yend = c(y,    y-d,   y-d,  y-d*1,y-d*1,y-d*2,y-d*2,y-d*2,y-d*2, y-d*3, y-d*3)
)
sigmap <- data.frame(
        x =     c(1.50, 1.00,        2.00),
        y =     c(y,    y-d*3+0.01, y-d*3),
        label = c('***','+', '***')
)
g <- g + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     size = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)
print(g)
save_plot(g, filename = '/home/sysneu/marjoh/Visualization/R_LongitudinalComparisonFmri/CorticalMD/cMD_inferiorparietal_HCvsPD.png', dpi=300, device='png', base_width = 5)

```

## *Compensation ~ cortical microstructure

```{r, echo=F,warning=F,message=F}

labels <- c('inferiorparietal','superiorparietal','precuneus','supramarginal',
            'paracentral','postcentral','precentral',
            'superiorfrontal','caudalmiddlefrontal')
# 'premotor','sensorimotor','parietal','allregions'
measure <- c('dti')
hemi <- c('bi')
prefix <- c('T2sub0')
cols_T2sub0 <- paste(prefix,hemi,labels,measure,sep='_')
prefix <- c('T0')
cols_T0 <- paste(prefix,hemi,labels,measure,sep='_')

tmp <- df_md %>% 
  filter(ParticipantType=='Patient',
         TimepointNr==0) %>%
  select(pseudonym, ParticipantType, TimepointNr, Age, Gender, NpsEducYears, YearsSinceDiag,
         T2sub0_Up3OnBradySum,T0_Up3OnBradySum,T2sub0_Up3OfBradySum,T0_Up3OfBradySum,
         all_of(cols_T2sub0),
         all_of(cols_T0),
         all_of(str_replace(cols_T0,'dti','thickavg')),
         all_of(str_replace(cols_T2sub0,'dti','thickavg')),
         T2sub0_bi_allregions_dti,T0_bi_allregions_dti) %>%
  select(-c(ParticipantType,TimepointNr))

tmp <- df_compensation_AddCov0 %>%
  left_join(., tmp, by = c('pseudonym')) %>%
  na.omit()

# Explore correlation structure
# tmp %>%
#   select(T2sub0_bi_premotor_comp,
#          T2sub0_Up3OnBradySum,
#          all_of(cols_T2sub0)) %>%
#   ggpairs()

tab <- c()
for(lab in 1:length(labels)){
  iv_T2sub0 <- cols_T2sub0[lab]
  iv_T0 <- cols_T0[lab]
  iv_thick_ba <- paste0(str_replace(iv_T0,'dti','thickavg'))
  iv_thick_delta <- paste0(str_replace(iv_T2sub0,'dti','thickavg'))
  iv_allregions_ba <- paste0('T0_',hemi,'_','allregions', '_', measure)
  iv_allregions_delta <- paste0('T2sub0_',hemi,'_','allregions', '_', measure)
  
  # Parietal comp
  f <- paste0('T2sub0_L_parietal_comp ~ ', iv_T2sub0, ' + Age + Gender + NpsEducYears + YearsSinceDiag + T0_L_parietal_comp +',iv_T0)
  m <- lm(f, data = tmp)
  a1 <- Anova(m,type=3)
  
  g1 <- tmp %>%
    ggplot(aes(x=.data[[iv_T2sub0]],y=T2sub0_L_parietal_comp)) +
    geom_point() + geom_smooth(method='lm',color='blue') + theme_bw() +
    ggtitle('Raw values of T2sub0_L_parietal_comp')
  pred <- ggpredict(m, terms=iv_T2sub0)
  g2 <- plot(pred, add.data = TRUE) + theme_bw() + 
        geom_hline(yintercept = 0, color = 'darkgrey', lty = 2) + geom_vline(xintercept = 0, color = 'darkgrey', lty = 2)
  g <- ggarrange(g1,g2,nrow = 2,ncol = 1)
  # print(g)
  
  # Premotor comp
  f <- paste0('T2sub0_bi_premotor_comp ~ ', iv_T2sub0, ' + Age + Gender + NpsEducYears + YearsSinceDiag + T0_bi_premotor_comp +',iv_T0)
  m <- lm(f, data = tmp)
  a2 <- Anova(m,type=3)
  
  # Parietopremotor comp
  f <- paste0('T2sub0_bi_parietopremotor_comp ~ ', iv_T2sub0, ' + Age + Gender + NpsEducYears + YearsSinceDiag + T0_bi_parietopremotor_comp +',iv_T0)
  m <- lm(f, data = tmp)
  a3 <- Anova(m,type=3)
  
  # rcor_dat <- df_md %>%
  #   filter(ParticipantType=='Patient',
  #          Exclusion_MD!=1,
  #          complete_case==1)
  # rmc_brady <- rmcorr('pseudonym',dv,'Up3OnBradySum',dataset = rcor_dat)
  # rmc_moca <- rmcorr('pseudonym',dv,'z_MoCA__total',dataset = rcor_dat)
  # rmc_ledd <- rmcorr('pseudonym',dv,'LEDD',dataset = rcor_dat)
  
  tab <- bind_rows(tab,
                   tibble(region=iv_T2sub0,
                          p_T2sub0_L_parietal_comp=a1[rownames(a1)==iv_T2sub0,4],
                          p_T2sub0_bi_premotor_comp=a2[rownames(a2)==iv_T2sub0,4],
                          p_T2sub0_bi_parietopremotor_comp=a3[rownames(a3)==iv_T2sub0,4])
  )
}
print(tab)
write_csv(tab,'/project/3024006.02/wd/ggseg_stats_COMPvsMD.csv')
write_csv(tmp,'/project/3024006.02/wd/ggseg_data_COMPvsMD.csv')

```

## Work that needs to be done



