---
title: "Longitudinal comparisons of motor task performance"
author: "M.E. Johansson"
date: "6/7/2022"
output: 
  html_document: 
    toc: yes
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare R environment

```{r libraries, echo=FALSE, warning=TRUE, message=F}
# Load libraries
options(contrasts=c('contr.sum','contr.poly'))
library(MASS); library(rstatix)
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(lme4); library(lmerTest); library(emmeans); library(effectsize); library(finalfit); library(effects); library(bmlm)
emm_options(lmerTest.df = 'satterthwaite')  # Calculates DFs in the same way as lmerTest
emm_options(lmerTest.limit = 50000)
emm_options(disable.pbkrtest=TRUE)
emm_options(opt.digits=F)
library(car)
library(extrafont)
library(ggpubr);library(ggforce);library(gghalves);library(ggdist);library(ggeffects); library(GGally); library(sjPlot); library(cowplot); library(viridis); library(lattice); library(see); library(kableExtra); library(gridExtra)
library(mice); library(miceadds)
library(rmcorr); library(ppcor); library(PResiduals); library(cocor); library(olsrr)
loadfonts(device='win',quiet = TRUE)

# Define paths and visualization parameters
glmnet_dir <- 'P:/3024006.02/Data/GLMNET_NatComm/'
sigsize=5
lsize=0.6
legend_position <- c(0.8,1.1)

# Define functions
LmPlots <- function(model){
        
        count <- 1
        while(count<7)
        {
                plot(model, which = count)
                count <- count + 1
        }
        
        modplot <- predictorEffects(model)
        # plot(modplot)
        
}
```

# Load data

## Clinical

```{r clinvars, echo=FALSE, warning=TRUE, message=TRUE}
# Load clinical data
fClin <- 'P:/3022026.01/pep/ClinVars_10-08-2023/derivatives/merged_manipulated_2024-07-17.csv'
dfClin <- read_csv(fClin, show_col_types = FALSE)

# Initial variable selection and clean up
        # Define relevant variables
clinvars <- c('pseudonym', 'TimepointNr', 'YearsToFollowUp', 'ParticipantType',
              'Subtype_DiagEx3_DisDurSplit','MriNeuroPsychTask',
              'Age', 'Gender', 'NpsEducYears', 'PrefHand', 'YearsSinceDiag',
              'ParkinMedUser', 'LEDD', 'Up3OfHoeYah',
              'Up3OfTotal', 'Up3OfAppendicularSum', 'Up3OfBradySum',
              'Up3OfRigiditySum', 'Up3OnTotal','Up3OnAppendicularSum', 'Up3OnBradySum','Up3OnRigiditySum',
              'Up3OnRestTremAmpSum','Up3OnActionTremorSum','Up3OnCompositeTremorSum','z_MoCA__total', 'z_CognitiveComposite2',
              'AsymmetryIndexRiLe.WeightedBradyRig2',
              'BDI2Sum','DiagParkCertain', 'DiagParkPersist','DiagParkReplace',
              'STAITraitSum', 'STAIStateSum', 'QUIPicdSum','MoCASum')
        # Fix TimepointNr inconsistency between patients and controls
        # Remove PD_PIT sessions and include only motor task subjects
        # Factorize and convert units
dfClin <- dfClin %>%
        mutate(TimepointNr=if_else(ParticipantType=='HC_PIT' & TimepointNr==1, 2,
                                   TimepointNr)) %>%
        filter(ParticipantType=='HC_PIT' | ParticipantType=='PD_POM') %>%
        select(all_of(clinvars)) %>%
        rename(Subtype = Subtype_DiagEx3_DisDurSplit) %>%
        mutate(TimepointNr = factor(TimepointNr, levels = c(0,1,2), labels = c('0','1','2')),
               ParticipantType = factor(ParticipantType),
               Gender = factor(Gender),
               Subtype = if_else(ParticipantType=='HC_PIT', '0_Healthy', Subtype),
               Subtype = if_else(is.na(Subtype),'4_Undefined',Subtype),
               Subtype = factor(Subtype),
               Up3OfHoeYah=factor(Up3OfHoeYah),
               MostAffSide = case_when(AsymmetryIndexRiLe.WeightedBradyRig2 < 0 ~ 'Left',
                                       AsymmetryIndexRiLe.WeightedBradyRig2 > 0 ~ 'Right',
                                       AsymmetryIndexRiLe.WeightedBradyRig2 == 0 ~ 'None',
                                       .default = NA))

# Set relevant covariates to their baseline values
baseline_covs <- dfClin %>%
        filter(TimepointNr==0) %>%
        select(pseudonym,ParticipantType,Age,Gender,NpsEducYears,PrefHand,MostAffSide,Subtype)
dfClin <- dfClin %>% 
        select(-c(Age,Gender,NpsEducYears,PrefHand,Subtype,MostAffSide)) %>%
        left_join(.,baseline_covs,by=c('pseudonym','ParticipantType'))

# Print N
dfClin %>% select(ParticipantType,TimepointNr) %>% table() %>% print()

# Select ON or OFF medication scores
CLINVAR='Up3OnBradySum'         # This variable is used for the multiple imputation
dfClin <- dfClin %>%
        mutate(Up3Total=Up3OnTotal, 
               Up3AppendicularSum=Up3OnAppendicularSum, 
               Up3BradySum=Up3OnBradySum,
               Up3RigiditySum=Up3OnRigiditySum)

glimpse(dfClin)

# This participant has a very peculiar clinical trajectory. At baseline,
# patient presents with extremely high Up3 values in both OFF and ON. Over two
# years, the symptoms decline gradually while medication stays very low. This
# decline is not incremental, it is massive. 20 points drop in two years.
dfClin <- dfClin %>%
        filter(pseudonym != 'sub-POMU0E19B895DF700AB0')
```

## Task performance

```{r behav-fmri, echo=FALSE, warning=TRUE, message=F}
# Load task performance data
fTask <- 'P:/3022026.01/pep/bids/derivatives/manipulated_merged_motor_task_mri_2023-09-15.csv'
dfTask <- read_csv(fTask, show_col_types = FALSE)

# Initial variable selection and clean up
dfTask <- dfTask %>%
        filter(Group=='HC_PIT' | Group=='PD_POM') %>%
    select(pseudonym, Group, Timepoint, trial_type, response_time, trial_number,
           event_type, correct_response, block, RespondingHand, Group, 
           Button.Press.SwitchRatio, Button.Press.CoV) %>%
    mutate(TimepointNr = if_else(str_detect(Timepoint, 'Visit1'), '0', '2'),
           TimepointNr = factor(TimepointNr),
           trial_type=factor(trial_type),
           block=factor(block),
           RespondingHand=factor(RespondingHand),
           error = if_else(correct_response=='Incorrect',1,0)) %>%
    relocate(pseudonym, TimepointNr) %>%
    select(-Timepoint) %>%
        rename(ParticipantType=Group)

# Count missed responses
misses <- dfTask %>%
        filter(is.na(response_time), event_type=='cue', correct_response=='Miss') %>%
        group_by(pseudonym, ParticipantType, TimepointNr, trial_type, block) %>%
        summarise(n_misses=n()) %>%
        ungroup()
dfTask <- dfTask %>%
        left_join(., misses) %>%
        mutate(n_misses = if_else(is.na(n_misses),0,n_misses))

# Divide data into response time and error rate
dfTask.rt <- dfTask %>%
    filter(event_type == 'response') %>%
    filter(trial_type != 'Catch') %>%
    filter(response_time > 0.3) %>%
    filter(correct_response=='Hit') %>%
    select(pseudonym, ParticipantType, TimepointNr, trial_type, block, RespondingHand,
           response_time, Button.Press.CoV, Button.Press.SwitchRatio, n_misses) %>%
    mutate(trial_type=factor(trial_type))
dfTask.err <- dfTask %>%
    filter(event_type == 'response') %>%
    filter(trial_type != 'Catch') %>%
    filter(response_time > 0.3) %>%
    select(pseudonym, ParticipantType, TimepointNr, trial_type, block, RespondingHand,
           error, Button.Press.CoV, Button.Press.SwitchRatio, n_misses) %>%
    mutate(trial_type=factor(trial_type))
# Aggregate
dfTask.rt <- dfTask.rt %>%
    group_by(pseudonym, ParticipantType, TimepointNr, trial_type, block) %>%
    summarise(RespondingHand = first(RespondingHand),
              Button.Press.CoV = first(Button.Press.CoV),
              Button.Press.SwitchRatio=first(Button.Press.SwitchRatio),
              n_misses = first(n_misses),
              response_time = median(response_time,na.rm=TRUE)) %>%
    ungroup()
        # n_trials   = No. of correct and incorrect trials. Misses are excluded
dfTask.err <- dfTask.err %>%
    group_by(pseudonym, ParticipantType, TimepointNr, trial_type, block) %>%
    summarise(RespondingHand = first(RespondingHand),
              Button.Press.CoV = first(Button.Press.CoV),
              Button.Press.SwitchRatio=first(Button.Press.SwitchRatio),
              n_misses = first(n_misses),
              n_trials=n(),
              n_errors=sum(error),
              n_corrects=n_trials - n_errors,
              error_rate = (n_errors/n_trials),
              correct_rate = (n_corrects/n_trials)) %>%
    ungroup()

# Merge response time and error rate data again, add grouping variable from
# clinical data
dfTask <- full_join(dfTask.rt,dfTask.err) %>%
        mutate(total_misses_by_time = sum(n_misses), .by=c(pseudonym,ParticipantType,TimepointNr))

glimpse(dfTask)
```

## Practice performance

```{r behav-prac, echo=FALSE, warning=TRUE, message=F}
# Practice data
# Load
fTaskPrac <- 'P:/3022026.01/pep/bids/derivatives/manipulated_merged_motor_task_practice_2023-09-15.csv'
dfTaskPrac <- read_csv(fTaskPrac, show_col_types = FALSE)
dfTaskPrac <- dfTaskPrac %>%
        filter(Group=='HC_PIT' | Group=='PD_POM') %>%
    select(pseudonym, Group, Timepoint, trial_type, response_time, trial_number,
           event_type, correct_response, RespondingHand) %>%
    mutate(TimepointNr = if_else(str_detect(Timepoint, 'Visit1'), '0', '2'),
           TimepointNr = factor(TimepointNr),
           RespondingHand=factor(RespondingHand),
           error = if_else(correct_response=='Incorrect',1,0)) %>%
    relocate(pseudonym, TimepointNr) %>%
    select(-Timepoint) %>%
        rename(ParticipantType=Group)
# Clean
dfTaskPrac.rt <- dfTaskPrac %>%
    filter(event_type == 'response') %>%
    filter(trial_type != 'Catch') %>%
    filter(response_time > 0.3) %>%
    filter(correct_response=='Hit') %>%
    select(pseudonym, ParticipantType, TimepointNr, trial_type, RespondingHand, response_time) %>%
    mutate(trial_type=factor(trial_type))
dfTaskPrac.err <- dfTaskPrac %>%
    filter(event_type == 'response') %>%
    filter(trial_type != 'Catch') %>%
    filter(response_time > 0.3) %>%
    select(pseudonym, ParticipantType, TimepointNr, trial_type, RespondingHand, error) %>%
    mutate(trial_type=factor(trial_type))
# Aggregate
dfTaskPrac.rt <- dfTaskPrac.rt %>%
    group_by(pseudonym, ParticipantType, TimepointNr, trial_type) %>%
    summarise(RespondingHand = first(RespondingHand),
              response_time = median(response_time,na.rm=TRUE)) %>%
    ungroup()
dfTaskPrac.err <- dfTaskPrac.err %>%
    group_by(pseudonym, ParticipantType, TimepointNr, trial_type) %>%
    summarise(RespondingHand = first(RespondingHand),
              n_trials=n(),
              n_errors=sum(error),
              n_corrects=n_trials - n_errors,
              error_rate = (n_errors/n_trials),
              correct_rate = (n_corrects/n_trials)) %>%
    ungroup()
dfTaskPrac <- full_join(dfTaskPrac.rt,dfTaskPrac.err)

glimpse(dfTaskPrac)
```

# Exclusions
 
## Diagnosis

 - Misdiagnosis: Some patients had their diagnosis re-evaluated by a movement disorder
 specialist at Radboudumc, either at baseline or at two-year follow-up.

```{r exclude-diag, echo=FALSE, warning=TRUE, message=TRUE}

# Define misdiagnoses
diagnosis <- dfClin %>% select(pseudonym, TimepointNr, DiagParkCertain, DiagParkPersist, DiagParkReplace)
baseline_exclusion <- diagnosis %>%
    filter(TimepointNr=='0', (DiagParkCertain == 'NeitherDisease' | DiagParkCertain == 'DoubtAboutParkinsonism' | DiagParkCertain == 'Parkinsonism')) %>% 
    select(pseudonym, DiagParkCertain)
visit2_exclusion <- diagnosis %>%
    filter(TimepointNr=='2', (DiagParkPersist == 2)) %>% 
    select(pseudonym, DiagParkPersist, DiagParkReplace)
diag_exclusions <- full_join(baseline_exclusion, visit2_exclusion, by='pseudonym') %>% 
    unique()
dfClin <- dfClin %>%
    mutate(Misdiagnosis = if_else(pseudonym %in% diag_exclusions$pseudonym,1,0))

# Report patients with confirmed non-PD diagnosis
cat('>>> Excluding patients with confirmed non-PD diagnosis at baseline and patients who received a non-PD diagnosis at 2-year follow-up\n')
diagnosis %>% filter(TimepointNr=='0') %>% select(DiagParkCertain) %>% table() %>% kable(., 'pipe', digits = 5, padding=0) %>% print()
diagnosis %>% filter(TimepointNr=='2') %>% select(DiagParkPersist) %>% table() %>% kable(., 'pipe', digits = 5, padding=0) %>% print()
diagnosis %>% filter(TimepointNr=='2') %>% select(DiagParkReplace) %>% table() %>% kable(., 'pipe', digits = 5, padding=0) %>% print()
cat('>>> Excluded based on misdiagnosis:\n')
diag_exclusions %>% kable(., 'pipe', digits = 5, padding=0)
write_csv(diag_exclusions, 'P:/3024006.02/Data/exclusions_diagnosis.csv')

# Checks on patients who had a doubtful diagnosis at baseline
cat('>>> How does the 2-year follow-up diagnosis change for patients with uncertain PD diagnoses at baseline?\n')
cat('>>> DoubtAboutPD\n')
doubt_at_ba <- diagnosis %>%
    filter(DiagParkCertain == 'DoubtAboutPD') %>%
    select(pseudonym)
diagnosis %>% filter(TimepointNr=='2', pseudonym %in% doubt_at_ba$pseudonym) %>% select(DiagParkPersist) %>% table() %>% kable(., 'pipe', digits = 5, padding=0) %>% print()
cat('>>> Parkinsonism\n')
parkinsonism_at_ba <- diagnosis %>%
    filter(DiagParkCertain == 'Parkinsonism') %>%
    select(pseudonym)
diagnosis %>% filter(TimepointNr=='2', pseudonym %in% parkinsonism_at_ba$pseudonym) %>% select(DiagParkPersist) %>% table() %>% kable(., 'pipe', digits = 5, padding=0) %>% print()
cat('>>> NeitherDisease\n')
neither_at_ba <- diagnosis %>%
    filter(DiagParkCertain == 'NeitherDisease') %>%
    select(pseudonym)
diagnosis %>% filter(TimepointNr=='2', pseudonym %in% neither_at_ba$pseudonym) %>% select(DiagParkPersist) %>% table() %>% kable(., 'pipe', digits = 5, padding=0) %>% print()
cat('>>> DoubtAboutParkinsonism\n')
doubt2_at_ba <- diagnosis %>%
    filter(DiagParkCertain == 'DoubtAboutParkinsonism') %>%
    select(pseudonym)
diagnosis %>% filter(TimepointNr=='2', pseudonym %in% doubt2_at_ba$pseudonym) %>% select(DiagParkPersist) %>% table() %>% kable(., 'pipe', digits = 5, padding=0) %>% print()

# Exclude subjects with non-PD diagnosis
n1 <- dfTask %>%
    group_by(pseudonym,ParticipantType,TimepointNr) %>%
    summarise(across(everything(), ~first(.x))) %>%
    group_by(ParticipantType,TimepointNr) %>%
    summarise(nTask=n())
n2 <- dfTaskPrac %>%
    group_by(pseudonym,ParticipantType,TimepointNr) %>%
    summarise(across(everything(), ~first(.x))) %>%
    group_by(ParticipantType,TimepointNr) %>%
    summarise(nTaskPrac=n())
n3 <- dfClin %>%
    group_by(ParticipantType,TimepointNr) %>%
    summarise(nClin=n())
n_before <- full_join(n1, n2, by=c('ParticipantType','TimepointNr')) %>%
    full_join(., n3, by=c('ParticipantType','TimepointNr'))

dfTask <- dfTask %>%
    filter(!(pseudonym %in% diag_exclusions$pseudonym))
dfTaskPrac <- dfTaskPrac %>%
    filter(!(pseudonym %in% diag_exclusions$pseudonym))
dfClin <- dfClin %>%
       filter(!(pseudonym %in% diag_exclusions$pseudonym)) 

n1 <- dfTask %>%
    group_by(pseudonym,ParticipantType,TimepointNr) %>%
    summarise(across(everything(), ~first(.x))) %>%
    group_by(ParticipantType,TimepointNr) %>%
    summarise(nTask=n())
n2 <- dfTaskPrac %>%
    group_by(pseudonym,ParticipantType,TimepointNr) %>%
    summarise(across(everything(), ~first(.x))) %>%
    group_by(ParticipantType,TimepointNr) %>%
    summarise(nTaskPrac=n())
n3 <- dfClin %>%
    group_by(ParticipantType,TimepointNr) %>%
    summarise(nClin=n())
n_after <- full_join(n1, n2, by=c('ParticipantType','TimepointNr')) %>%
    full_join(., n3, by=c('ParticipantType','TimepointNr'))

cat('>>> Before diagnosis exclusions:\n', sep='')
print(n_before)
cat('>>> After diagnosis exclusions:\n', sep='')
print(n_after)

```

## Accuracy

 - Low accuracy (< 25% correct on one-choice trials, averaged across blocks):
 Excluding participants due to poor accuracy could be done in multiple ways. Here,
 we opted to define a criterion for exclusion based on the idea that random presses of
 buttons should result in 25% correct responses on single-choice trials.

```{r exclude-perf, echo=FALSE, warning=TRUE, message=TRUE}

cat('>>> Excluding sessions where the across-block error rate in 1choice trials is <=25%\n')
# Define poor performance
poor_performance <- dfTask %>%
    filter(trial_type=='NChoice1') %>%
    group_by(ParticipantType, pseudonym, TimepointNr, trial_type, block) %>%
    summarise(across(everything(), ~first(.x))) %>% 
    ungroup() %>%
    select(pseudonym, ParticipantType, TimepointNr, trial_type, block, correct_rate) %>%
    pivot_wider(id_cols = c('pseudonym','ParticipantType','TimepointNr', 'trial_type'),
                values_from = correct_rate,
                names_from = block,
                names_prefix = 'block_') %>%
    mutate(success_rate.avg = ((block_1+block_2+block_3)/3),
           poor_performance = if_else(success_rate.avg<=0.25,1,0)) %>% 
    filter(poor_performance==1) %>%
    select(pseudonym, ParticipantType, TimepointNr, block_1, block_2, block_3, success_rate.avg)
cat('>>> Excluded sessions based on error rate:',nrow(poor_performance),'\n')
poor_performance %>% kable(., 'pipe', digits = 5, padding=0)
poor_performance %>% select(ParticipantType,TimepointNr) %>% table() %>% print()
# Write excluded participants to file
write_csv(poor_performance, 'P:/3024006.02/Data/exclusions_accuarcy.csv')

# Count number before excluding
n_before <- dfTask %>%
    group_by(pseudonym,ParticipantType,TimepointNr) %>%
    summarise(across(everything(), ~first(.x))) %>%
    group_by(ParticipantType,TimepointNr) %>%
    summarise(nTask=n())

# Exclude
for(n in 1:nrow(poor_performance)){
    row <- poor_performance[n,]
    idx <- which(dfTask$pseudonym==row$pseudonym & dfTask$TimepointNr==row$TimepointNr)
    dfTask <- dfTask[-idx,]
}

# Count number after exlcuding
n_after <- dfTask %>%
    group_by(pseudonym,ParticipantType,TimepointNr) %>%
    summarise(across(everything(), ~first(.x))) %>%
    group_by(ParticipantType,TimepointNr) %>%
    summarise(nTask=n())

cat('>>> Before accuracy exclusions:\n', sep='')
print(n_before)
cat('>>> After accuracy exclusions:\n', sep='')
print(n_after)

```

## Misses

 - Excessive missing (< 60 misses across the whole session): 60 is an arbitrarily
 chosen value. We consider 60 missed trials to reflect major data loss, leading to
 unreliable estimates of response times, which justifies exclusion.

```{r exclude-missresp, echo=FALSE, warning=TRUE, message=TRUE}

thr=60
cat('>>> Excluding sessions where there were too many misses (n>', thr, '):\n', sep='')

# Count number before exclusion
n_before <- dfTask %>%
        group_by(pseudonym,ParticipantType,TimepointNr) %>%
        summarise(ParticipantType=first(ParticipantType),
                  TimepointNr=first(TimepointNr),
                  total_misses_by_time=first(total_misses_by_time)) %>%
        ungroup() %>%
        group_by(ParticipantType,TimepointNr) %>%
        summarise(N=n(),
                  Min=min(total_misses_by_time),
                  Max=max(total_misses_by_time),
                  Mean=mean(total_misses_by_time),
                  SD=sd(total_misses_by_time)) %>%
        ungroup()

# Define poor performance
poor_performance <- dfTask %>%
        group_by(pseudonym,ParticipantType,TimepointNr) %>%
        summarise(ParticipantType=first(ParticipantType),
                  TimepointNr=first(TimepointNr),
                  total_misses_by_time=first(total_misses_by_time),
                  error_rate = mean(error_rate),
                  RT = mean(response_time)) %>%
        mutate(poor_performance = if_else(total_misses_by_time > thr,1,0)) %>%
        filter(poor_performance == 1) %>%
        ungroup()
cat('>>> Excluded sessions based on number of misses:',nrow(poor_performance),'\n')
poor_performance %>% kable(., 'pipe', digits = 5, padding=0)
poor_performance %>% select(ParticipantType,TimepointNr) %>% table() %>% print()
# Write excluded participants to file
write_csv(poor_performance, 'P:/3024006.02/Data/exclusions_misses.csv')

# Count number after exclusion
n_after <- dfTask %>%
        filter(total_misses_by_time < thr) %>%
        group_by(pseudonym,ParticipantType,TimepointNr) %>%
        summarise(ParticipantType=first(ParticipantType),
                  TimepointNr=first(TimepointNr),
                  total_misses_by_time=first(total_misses_by_time)) %>%
        ungroup() %>%
        group_by(ParticipantType,TimepointNr) %>%
        summarise(N=n(),
                  Min=min(total_misses_by_time),
                  Max=max(total_misses_by_time),
                  Mean=mean(total_misses_by_time),
                  SD=sd(total_misses_by_time)) %>%
        ungroup()

cat('>>> Before exclusions due to too many misses:\n', sep='')
print(n_before)
cat('>>> After exclusions due to too many misses:\n', sep='')
print(n_after)

# Carry out exclusion
dfTask <- dfTask %>%
        filter(total_misses_by_time < thr)

```

# Data organization

## Assemble data and impute missing values

```{r data-assembly, echo=FALSE, warning=TRUE, message=TRUE}
# Extract confounders
vars.conf <- c('pseudonym', 'ParticipantType', 'Age', 'Gender', 'YearsSinceDiag', 'NpsEducYears', 'PrefHand', 'MostAffSide')
dfClin.conf <- dfClin %>%
    filter(TimepointNr==0) %>%
    select(all_of(vars.conf))

# Extract clinical progression
vars.prog <- c('pseudonym', 'TimepointNr', 'Age', 'Gender', 'YearsSinceDiag', 'NpsEducYears',
               'ParticipantType', 'Up3OfBradySum', 'Up3OnBradySum', 'Up3BradySum','z_MoCA__total','LEDD')
dfClin.prog <- dfClin %>%
    filter(ParticipantType=='PD_POM',
           TimepointNr == 0 | TimepointNr == 2) %>%
    select(all_of(vars.prog))

dfClin.prog %>%
        mutate(UP3OFF.na = is.na(Up3OfBradySum),
               UP3ON.na = is.na(Up3OnBradySum)) %>%
        group_by(ParticipantType,TimepointNr) %>%
        count(UP3OFF.na) %>% print()

dfClin.prog %>%
        mutate(UP3OFF.na = is.na(Up3OfBradySum),
               UP3ON.na = is.na(Up3OnBradySum)) %>%
        group_by(ParticipantType,TimepointNr) %>%
        count(UP3ON.na) %>% print()

## Multivariate imputation by chained equations (MICE)
## Note that this stage is fragile with respect to changes in
## the clinical variable that is investigated. "CLINVAR" was set up
## as a workaround to enable the imputation to account for medication effects
dfClin.prog.imp <- dfClin.prog %>%
        select(pseudonym, Up3OfBradySum, Up3OnBradySum, z_MoCA__total, LEDD, TimepointNr, Age, Gender, YearsSinceDiag, NpsEducYears) %>%
        pivot_longer(cols=c('Up3OfBradySum','Up3OnBradySum'),
                     names_to = 'Medication',
                     values_to = 'Up3BradySum') %>%
        mutate(Medication=factor(Medication))
md.pattern(dfClin.prog.imp)
isna <- apply(dfClin.prog.imp, 2, is.na) %>% colSums()
cat(paste('\n ', '>>> Percentage missing values\n', sep=''))
missing_perc <- round(isna/nrow(dfClin.prog.imp), digits = 5)*100
print(missing_perc)
imputed <- dfClin.prog.imp %>% 
        mice(m=round(5*missing_perc[names(missing_perc)=='Up3BradySum']), 
             maxit = 10, 
             method='pmm', 
             seed=157, 
             print=FALSE)
summary(imputed)
print(stripplot(imputed, formula('Up3BradySum ~ TimepointNr',sep='')),
      pch = 19)
print(stripplot(imputed, formula('z_MoCA__total ~ TimepointNr',sep='')),
      pch = 19)
print(stripplot(imputed, formula('LEDD ~ TimepointNr',sep='')),
      pch = 19)
dfClin.prog.imp <- imputed %>%
        complete() %>%
        tibble() %>%
        filter(Medication==CLINVAR) %>%
        select(pseudonym, TimepointNr, Up3BradySum, z_MoCA__total, LEDD) %>%
        rename(Up3BradySum.imp = Up3BradySum,
               z_MoCA__total.imp = z_MoCA__total,
               LEDD.imp = LEDD)

# Merge imputed variables with original data
dfClin.prog <- dfClin.prog %>%
        left_join(., dfClin.prog.imp, by=c('pseudonym','TimepointNr'))

## Pivot wider to allow calculation of change scores
dfClin.prog <- dfClin.prog %>%
    pivot_wider(id_cols = c('pseudonym','ParticipantType'),
                names_from = TimepointNr,
                names_prefix = 'T',
                values_from = c(Up3OfBradySum,Up3OnBradySum,Up3BradySum,Up3BradySum.imp,z_MoCA__total,z_MoCA__total.imp,LEDD,LEDD.imp))

## Calculate raw change score
dfClin.prog <- dfClin.prog %>%
    mutate(Up3OfBradySum_RawChange = Up3OfBradySum_T2-Up3OfBradySum_T0,
           Up3OnBradySum_RawChange = Up3OnBradySum_T2-Up3OnBradySum_T0,
           Up3BradySum_RawChange = Up3BradySum_T2-Up3BradySum_T0,
           Up3BradySum.imp_RawChange = Up3BradySum.imp_T2-Up3BradySum.imp_T0,
           z_MoCA__total_RawChange = z_MoCA__total.imp_T2 - z_MoCA__total_T0,
           z_MoCA__total.imp_RawChange = z_MoCA__total.imp_T2 - z_MoCA__total.imp_T0,
           LEDD_RawChange = LEDD_T2 - LEDD_T0,
           LEDD.imp_RawChange = LEDD.imp_T2 - LEDD.imp_T0) %>%
    select(pseudonym, ParticipantType, 
           Up3OfBradySum_T0, Up3OfBradySum_T2, Up3OfBradySum_RawChange,
           Up3OnBradySum_T0, Up3OnBradySum_T2, Up3OnBradySum_RawChange,
           Up3BradySum_T0, Up3BradySum_T2, Up3BradySum_RawChange,
           Up3BradySum.imp_T0, Up3BradySum.imp_T2, Up3BradySum.imp_RawChange,
           z_MoCA__total_T0, z_MoCA__total_T2, z_MoCA__total_RawChange,
           z_MoCA__total.imp_T0, z_MoCA__total.imp_T2, z_MoCA__total.imp_RawChange,
           LEDD_T0, LEDD_T2, LEDD_RawChange, LEDD.imp_T0, LEDD.imp_T2, LEDD.imp_RawChange)

# Extract time to follow-up
vars.time <- c('pseudonym', 'TimepointNr', 'YearsToFollowUp')
dfClin.time <- dfClin %>%
    select(all_of(vars.time))

# Merge clinical and task data
dfFMRI <- left_join(dfTask, dfClin.conf, by = c('pseudonym', 'ParticipantType')) %>%
    left_join(., dfClin.time, by = c('pseudonym','TimepointNr')) %>%
    left_join(., dfClin.prog, by = c('pseudonym','ParticipantType'))

# Do the same for practice data
dfPrac <- left_join(dfTaskPrac, dfClin.conf, by = c('pseudonym', 'ParticipantType')) %>%
    left_join(., dfClin.time, by = c('pseudonym', 'TimepointNr')) %>%
    left_join(., dfClin.prog, by = c('pseudonym','ParticipantType'))

```

## Mutation

```{r data-mutation, echo=FALSE, warning=TRUE, message=TRUE}

# Restore clinical scores to their time-varying originals
dfFMRI <- dfFMRI %>%
    mutate(Up3OfBradySum = Up3OfBradySum_T0,
           Up3OfBradySum = if_else(TimepointNr=='2',Up3OfBradySum_T0+Up3OfBradySum_RawChange,Up3OfBradySum),
           Up3OnBradySum = Up3OnBradySum_T0,
           Up3OnBradySum = if_else(TimepointNr=='2',Up3OnBradySum_T0+Up3OnBradySum_RawChange,Up3OnBradySum),
           Up3BradySum = Up3BradySum.imp_T0,
           Up3BradySum = if_else(TimepointNr=='2',Up3BradySum.imp_T0+Up3BradySum.imp_RawChange,Up3BradySum),
           z_MoCA__total = z_MoCA__total.imp_T0,
           z_MoCA__total = if_else(TimepointNr=='2',z_MoCA__total.imp_T0+z_MoCA__total.imp_RawChange,z_MoCA__total),
           LEDD = LEDD.imp_T0,
           LEDD = if_else(TimepointNr=='2',LEDD.imp_T0+LEDD.imp_RawChange,LEDD))
dfPrac <- dfPrac %>%
    mutate(Up3OfBradySum = Up3OfBradySum_T0,
           Up3OfBradySum = if_else(TimepointNr=='2',Up3OfBradySum_T0+Up3OfBradySum_RawChange,Up3OfBradySum),
           Up3OnBradySum = Up3OnBradySum_T0,
           Up3OnBradySum = if_else(TimepointNr=='2',Up3OnBradySum_T0+Up3OnBradySum_RawChange,Up3OnBradySum),
           Up3BradySum = Up3BradySum.imp_T0,
           Up3BradySum = if_else(TimepointNr=='2',Up3BradySum.imp_T0+Up3BradySum.imp_RawChange,Up3BradySum),
           z_MoCA__total = z_MoCA__total.imp_T0,
           z_MoCA__total = if_else(TimepointNr=='2',z_MoCA__total.imp_T0+z_MoCA__total.imp_RawChange,z_MoCA__total),
           LEDD = LEDD.imp_T0,
           LEDD = if_else(TimepointNr=='2',LEDD.imp_T0+LEDD.imp_RawChange,LEDD))

# Determine whether most affected side is the dominant side. This is important
# because patients respond with their most affected side. Affectedness and
# dominance might lead to additive effects on task performance
dfFMRI <- dfFMRI %>%
        mutate(RespHandIsDominant = if_else(RespondingHand == PrefHand
                                            | RespondingHand == 'NoPref', 1, 0),
               RespHandIsDominant = factor(RespHandIsDominant))
dfPrac <- dfPrac %>%
        mutate(RespHandIsDominant = if_else(RespondingHand == PrefHand
                                            | RespondingHand == 'NoPref', 1, 0),
               RespHandIsDominant = factor(RespHandIsDominant))

# Include only baseline and two-year follow-up data
dfFMRI <- dfFMRI %>%
        filter(TimepointNr==0 | TimepointNr==2) %>%
        mutate(across(where(is.factor), droplevels))
dfPrac <- dfPrac %>%
        filter(TimepointNr==0 | TimepointNr==2) %>%
        mutate(across(where(is.factor), droplevels))

# Inlcude only participants who performed the action selection task for 
# the demographics
dfClin <- dfClin %>% filter(MriNeuroPsychTask=='Motor')

```

# *Demographics

## *Patients versus controls

```{r demographics-pdandhc, echo=FALSE, warning=TRUE, message=TRUE}

# Count number of participants
dfClin %>%
        select(ParticipantType,TimepointNr) %>%
        table()

# Compare demographics between patients and controls
cat('>>> Demographics: controls v patients \n')
dependent <- c('ParticipantType')
exploratory <- c('Age', 'Gender', 'NpsEducYears', 'PrefHand', 'MoCASum',
                 'BDI2Sum','STAITraitSum', 'STAIStateSum', 'QUIPicdSum')
split <- c('TimepointNr')
dfClin %>% 
        filter(TimepointNr==0 | TimepointNr==2) %>%
    summary_factorlist_stratified(dependent, 
                                  exploratory,
                                  split = split,
                                  add_dependent_label=TRUE,
                                  p = TRUE, 
                                  p_cont_para = 't.test',
                                  p_cat = 'chisq',
                                  na_include = TRUE,
                                  colname_sep = ':Time') %>%
    kable(., 'pipe', digits = 5, padding=0)

# Responding hand
dfFMRI %>% 
        group_by(pseudonym,TimepointNr) %>% 
        summarise(ParticipantType=first(ParticipantType),RespondingHand=first(RespondingHand)) %>% 
        ungroup() %>% 
        select(-pseudonym) %>% 
        group_by(TimepointNr) %>% 
        table()

# Responding hand is dominant
dfFMRI %>% 
        group_by(pseudonym,TimepointNr) %>% 
        summarise(ParticipantType=first(ParticipantType),RespHandIsDominant=first(RespHandIsDominant)) %>% 
        ungroup() %>% 
        select(-pseudonym) %>% 
        group_by(TimepointNr) %>% 
        table()

# Compare responding hand between patients and controls
dfFMRI %>%
        filter(TimepointNr==0 | TimepointNr==2) %>%
        group_by(pseudonym,ParticipantType,TimepointNr) %>%
        summarise(RespondingHand=first(RespondingHand)) %>%
        mutate(ParticipantType=factor(ParticipantType)) %>%
        ungroup() %>%
        summary_factorlist_stratified('ParticipantType', 
                                  'RespondingHand',
                                  split = 'TimepointNr',
                                  add_dependent_label=TRUE,
                                  p = TRUE, 
                                  p_cont_para = 't.test',
                                  p_cat = 'chisq',
                                  na_include = TRUE,
                                  colname_sep = ':Time') %>%
    kable(., 'pipe', digits = 5, padding=0) %>% 
    print()

# Compare responding hand is dominant between patients and controls
dfFMRI %>%
        filter(TimepointNr==0 | TimepointNr==2) %>%
        group_by(pseudonym,ParticipantType,TimepointNr) %>%
        summarise(RespHandIsDominant=first(RespHandIsDominant)) %>%
        mutate(ParticipantType=factor(ParticipantType)) %>%
        ungroup() %>%
        summary_factorlist_stratified('ParticipantType', 
                                  'RespHandIsDominant',
                                  split = 'TimepointNr',
                                  add_dependent_label=TRUE,
                                  p = TRUE, 
                                  p_cont_para = 't.test',
                                  p_cat = 'chisq',
                                  na_include = TRUE,
                                  colname_sep = ':Time') %>%
    kable(., 'pipe', digits = 5, padding=0) %>% 
    print()

# Compare time to follow-up between patients and controls
exploratory_time <- c('YearsToFollowUp')
dfClin %>%
    filter(TimepointNr==2) %>%
    summary_factorlist(dependent, 
                       exploratory_time,
                       add_dependent_label=TRUE,
                       p = TRUE, 
                       p_cont_para = 't.test',
                       p_cat = 'chisq',
                       na_include = TRUE) %>%
    kable(., 'pipe', digits = 5, padding=0) %>% 
    print()

```

## *Patient-specific

```{r demographics-pdonly, echo=FALSE, warning=TRUE, message=TRUE}

# Descriptives for patients
cat('>>> Demographics: patients only \n')
exploratory_pd <- c('YearsSinceDiag','ParkinMedUser','LEDD',
                    'Up3OfHoeYah','MostAffSide','Up3OfTotal',
                    'Up3OfBradySum', 'Up3OfRigiditySum',
                    'Up3OnTotal',
                    'Up3OnBradySum', 'Up3OnRigiditySum',
                    'z_MoCA__total')
dfClin %>%
        filter(ParticipantType=='PD_POM') %>%
        summary_factorlist('TimepointNr', 
                           exploratory_pd,
                           add_dependent_label=TRUE,
                           p = FALSE,
                           na_include = TRUE) %>%
        kable(., 'pipe', digits = 5, padding=0) %>% 
        print()

# Demographics for medication converters
cat('Demographics: patients who start medication \n')
MedConverters <- dfClin %>% 
        filter(ParticipantType=='PD_POM',
               TimepointNr!=1) %>% 
        select(pseudonym, TimepointNr, ParkinMedUser) %>%
        pivot_wider(id_cols = 'pseudonym',
                    names_from = TimepointNr,
                    names_prefix = 'T',
                    values_from = ParkinMedUser) %>%
        na.omit() %>%
        mutate(Converter = if_else(T0=='No' & T2=='Yes',1,0)) %>%
        filter(Converter==1)

dfClin %>%
        filter(pseudonym %in% MedConverters$pseudonym) %>%
        summary_factorlist('TimepointNr', 
                           exploratory_pd,
                           add_dependent_label=TRUE,
                           p = FALSE,
                           na_include = TRUE) %>%
        kable(., 'pipe', digits = 5, padding=0) %>% 
        print()        

```

# *Count files

```{r count-files, , echo=FALSE, warning=TRUE, message=TRUE}

t <- dfClin %>%
        filter(TimepointNr != 1) %>%
        select(pseudonym, TimepointNr, ParticipantType)
print('>>> Ground truth:')
t %>% 
        group_by(ParticipantType,TimepointNr) %>%
        count() %>%
        print()

# 
t <- t %>%
        mutate(Timepoint = case_when(
                TimepointNr == 0 & ParticipantType == 'PD_POM' ~ 'ses-POMVisit1',
                TimepointNr == 2 & ParticipantType == 'PD_POM' ~ 'ses-POMVisit3',
                TimepointNr == 0 & ParticipantType == 'HC_PIT' ~ 'ses-PITVisit1',
                TimepointNr == 2 & ParticipantType == 'HC_PIT' ~ 'ses-PITVisit2'
        ),
        dir.clin = file.path('P:/3022026.01/pep/ClinVars_10-08-2023', pseudonym, Timepoint),
        fn.beh = file.path('P:/3022026.01/pep/bids',pseudonym,Timepoint,'beh', paste0(pseudonym,'_',Timepoint,'_task-motor_acq-MB6_run-1_events.tsv')),
        fn.mb6 = file.path('P:/3022026.01/pep/bids',pseudonym,Timepoint,'func', paste0(pseudonym,'_',Timepoint,'_task-motor_acq-MB6_run-1_echo-1_bold.nii.gz')),
        fn.t1w = file.path('P:/3022026.01/pep/bids',pseudonym,Timepoint,'anat', paste0(pseudonym,'_',Timepoint,'_acq-MPRAGE_run-1_T1w.nii.gz')),
        fn.dwi = file.path('P:/3022026.01/pep/bids',pseudonym,Timepoint,'dwi', paste0(pseudonym,'_',Timepoint,'_acq-MB3_dir-AP_run-1_dwi.nii.gz')),
        clin.exist = dir.exists(dir.clin),
        beh.exist = file.exists(fn.beh),
        mb6.exist = file.exists(fn.mb6),
        t1w.exist = file.exists(fn.t1w),
        dwi.exist = file.exists(fn.dwi))

print('>>> Clinical variable directory:')
t %>% group_by(ParticipantType,TimepointNr) %>% count(clin.exist) %>% print()
print('>>> Behavioral events file:')
t %>% group_by(ParticipantType,TimepointNr) %>% count(beh.exist) %>% print()
print('>>> MB6 scan:')
t %>% group_by(ParticipantType,TimepointNr) %>% count(mb6.exist) %>% print()
print('>>> T1w scan:')
t %>% group_by(ParticipantType,TimepointNr) %>% count(t1w.exist) %>% print()
print('>>> DWI scan:')
t %>% group_by(ParticipantType,TimepointNr) %>% count(dwi.exist) %>% print()

```

# *Disease progression

## *Clin prog: Time x Medication

```{r on-vs-off-prog, echo=FALSE, warning=TRUE, message=TRUE}

##### Correlate bradykinesia on and off delta
tmp <- dfClin %>%
        filter(ParticipantType=='PD_POM',
               TimepointNr != 1) %>%
        select(pseudonym,TimepointNr,Age,Gender,YearsSinceDiag,NpsEducYears,Up3OfBradySum,Up3OnBradySum) %>%
        pivot_wider(id_cols = c('pseudonym','Age','Gender','YearsSinceDiag','NpsEducYears'),
                    names_from = TimepointNr,
                    names_prefix = 'T',
                    values_from = c('Up3OfBradySum','Up3OnBradySum')) %>%
        mutate(Gender=factor(Gender),
               Age=Age-mean(Age,na.rm=T),
               YearsSinceDiag=YearsSinceDiag-mean(YearsSinceDiag,na.rm=T),
               NpsEducYears=NpsEducYears-mean(NpsEducYears,na.rm=T),
               Up3OfBradySum_Td = Up3OfBradySum_T2 - Up3OfBradySum_T0,
               Up3OnBradySum_Td = Up3OnBradySum_T2 - Up3OnBradySum_T0) %>%
        na.omit()

# ggpairs(tmp, columns = 2:7,
#         lower = list(continuous = 'smooth'))
write_csv(tmp, paste0(glmnet_dir,'glmnet_clin.csv'))

m <- lm(Up3OfBradySum_Td ~ Up3OnBradySum_Td + Up3OfBradySum_T0 + Up3OnBradySum_T0 + Age + Gender + YearsSinceDiag + NpsEducYears, data = tmp)
tab_model(m,digits=5,show.se = T)
s <- summary(m,digits=5)
a <- Anova(m,type=3)
print(a)
kable(a, 'pipe', digits=Inf) %>% print()
eff <- eta_squared(m)
print(eff)
g <- ggemmeans(m, terms='Up3OnBradySum_Td') %>% 
        plot(add.data=T) + theme_bw() + ggtitle('') +
        xlab('Change in bradykinesia (ON)') + ylab('Change in bradykinesia (OFF)') +
        annotate(geom = 'text', 
                 x=-20, y=30, 
                 label=paste0('N=',nrow(tmp),', ', 
                              'β=', signif(s$coefficients[2,1],digits=5),', ',
                              'SE=', signif(s$coefficients[2,2],digits=5),', ', 
                              'P=', signif(s$coefficients[2,4],digits=5),', '),
                 hjust=0) + 
        ylim(-20,30)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/ClinicalProgression/ClinProg_OnOffCorr.png',
                  g, dpi=300, device='png', base_width = 5)

# Check for outliers in bradykinesia progression
sum_stats <- tmp %>%
        summarise(N=n(),
                  Up3OnBradySum_Td_avg = mean(Up3OnBradySum_Td),
                  Up3OnBradySum_Td_sd = sd(Up3OnBradySum_Td),
                  Up3OnBradySum_Td_upr = Up3OnBradySum_Td_avg + 3*Up3OnBradySum_Td_sd,
                  Up3OnBradySum_Td_lwr = Up3OnBradySum_Td_avg - 3*Up3OnBradySum_Td_sd,
                  Up3OfBradySum_Td_avg = mean(Up3OfBradySum_Td),
                  Up3OfBradySum_Td_sd = sd(Up3OfBradySum_Td),
                  Up3OfBradySum_Td_upr = Up3OfBradySum_Td_avg + 3*Up3OfBradySum_Td_sd,
                  Up3OfBradySum_Td_lwr = Up3OfBradySum_Td_avg - 3*Up3OfBradySum_Td_sd)
outliers <- tmp %>%
        mutate(
                outlier_ON = case_when(
                        Up3OnBradySum_Td > sum_stats$Up3OnBradySum_Td_upr ~ 1,
                        Up3OnBradySum_Td < sum_stats$Up3OnBradySum_Td_lwr ~ 1,
                        .default = 0
                ),
                outlier_OFF = case_when(
                        Up3OfBradySum_Td > sum_stats$Up3OfBradySum_Td_upr ~ 1,
                        Up3OfBradySum_Td < sum_stats$Up3OfBradySum_Td_lwr ~ 1,
                        .default = 0
                )
        )
print('>>> Progression outliers:')
outliers %>%
        filter(outlier_ON == 1 | outlier_OFF == 1) %>% print()

##### Breadykinesia: Time x Medication effect
tmp <- dfClin %>%
        filter(TimepointNr != 1) %>%
        select(pseudonym,TimepointNr,Up3OfBradySum,Up3OnBradySum,Age,Gender,YearsSinceDiag,NpsEducYears) %>%
        pivot_longer(cols = c('Up3OfBradySum','Up3OnBradySum'),
                     names_to = 'Medication',
                     values_to = 'Up3BradySum') %>%
        mutate(Medication=factor(Medication,
                                 levels=c('Up3OfBradySum','Up3OnBradySum'),
                                 labels=c('Off','On')),
               TimepointNr=factor(TimepointNr),
               Age=Age-mean(Age,na.rm=T),
               YearSinceDiag=YearsSinceDiag-mean(YearsSinceDiag,na.rm=T),
               NpsEducYears=NpsEducYears-mean(NpsEducYears,na.rm=T)) %>%
        na.omit()

n <- tmp %>% select(TimepointNr,Medication) %>% table()
print(n)
m <- lmer(Up3BradySum ~ TimepointNr*Medication + Age + Gender + YearsSinceDiag + NpsEducYears + (1+TimepointNr|pseudonym), data = tmp)
tab_model(m, digits=5,show.se = T)
a <- Anova(m, type=3)
print(a)
kable(a, 'pipe', digits=Inf) %>% print()
eff <- eta_squared(m)
kable(eff,'pipe',digits=Inf) %>% print()
em <- emmeans(m, ~Medication*TimepointNr)
contrast(em, interaction = c('revpairwise','revpairwise')) %>% kable(., 'pipe', digits=Inf) %>% print()
contrast(em, interaction = c('identity','revpairwise')) %>% kable(., 'pipe', digits=Inf) %>% print()
contrast(em, interaction = c('revpairwise','identity')) %>% kable(., 'pipe', digits=Inf) %>% print()
emmeans(m, revpairwise~TimepointNr)$contrast %>% kable(., 'pipe', digits=Inf) %>% print()
emmeans(m, revpairwise~Medication)$contrast %>% kable(., 'pipe', digits=Inf) %>% print()

g <- tmp %>%
        ggplot(aes(x=Medication, y = Up3BradySum, color=TimepointNr)) +
        geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
        geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
        stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
                     mapping = aes(group=TimepointNr), position = position_dodge(width=0.8)) +
        theme_bw() + 
        scale_x_discrete(labels=c('OFF','ON')) +
        scale_color_viridis_d(option='mako', begin = .1, end = .6,
                              direction = -1) + 
        ggtitle('') +
        ylab('MDS-UPDRS III score (bradykinesia)') + xlab('Medication') + 
        guides(color = guide_legend(title='Year')) +
        ylim(0, 47) + 
        annotate(geom='text',label=paste0('N=',n[1,1]),x=0.8,y=0) + 
        annotate(geom='text',label=paste0('N=',n[2,1]),x=1.2,y=0) + 
        annotate(geom='text',label=paste0('N=',n[1,2]),x=1.8,y=0) + 
        annotate(geom='text',label=paste0('N=',n[2,2]),x=2.2,y=0)
y <- 46
d = 1.5
segmap <- data.frame(
        x =    c(1.00, 1.00,  2.00,  0.80,  1.80,  0.80,  1.80),
        xend = c(2.00, 1.00,  2.00,  1.20,  2.20,  1.20,  2.20),
        y =    c(y,    y,     y,     y-d*1, y-d*1, y-d*3, y-d*3),
        yend = c(y,    y-d*1, y-d*1, y-d*1, y-d*1, y-d*3, y-d*3)
)
sigmap <- data.frame(
        x =     c(1.50,  1.00,  2.00),
        y =     c(y,     y-d*3, y-d*3),
        label = c('***', '***', '***')
)
g <- g + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     size = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)

print(g)

save_plot('M:/Visualization/R_LongitudinalComparisonFmri/ClinicalProgression/ClinProg_lmer_Brady.png',
          g, dpi=300, device='png', base_width = 5)

##### Effect of time on z-scored MoCA
tmp <- dfClin %>%
        filter(TimepointNr != 1) %>%
        select(pseudonym,TimepointNr,Up3OfBradySum,Up3OnBradySum,LEDD,z_MoCA__total,Age,Gender,YearsSinceDiag,NpsEducYears) %>%
        pivot_longer(cols = c('Up3OfBradySum','Up3OnBradySum'),
                     names_to = 'Medication',
                     values_to = 'Up3BradySum') %>%
        mutate(Medication=factor(Medication,
                                 levels=c('Up3OfBradySum','Up3OnBradySum'),
                                 labels=c('Off','On')),
               TimepointNr=factor(TimepointNr),
               Age=Age-mean(Age,na.rm=T),
               YearSinceDiag=YearsSinceDiag-mean(YearsSinceDiag,na.rm=T),
               NpsEducYears=NpsEducYears-mean(NpsEducYears,na.rm=T)) %>%
        na.omit()

m.MOCA <- tmp %>%
        filter(Medication=='On') %>%
        lmer(z_MoCA__total ~ TimepointNr + Age + Gender + YearsSinceDiag + NpsEducYears + (1|pseudonym), data = .)
tab_model(m.MOCA, digits = 5, show.se = T)
a <- Anova(m.MOCA, type=3)
print(a)
kable(a, 'pipe', digits=Inf) %>% print()
eta_squared(m.MOCA, alternative = 'two.sided') %>% kable(., 'pipe', digits = Inf) %>% print()
em <- emmeans(m.MOCA, ~ TimepointNr, type='response')
pairs(em) %>% kable(., 'pipe',digits=Inf)

# Correlation between MoCA increase and clinical progression
tmp.dMOCA <- tmp %>%
        filter(Medication=='On') %>%
        pivot_wider(id_cols = c('pseudonym','Age','Gender','YearsSinceDiag','NpsEducYears'),
                    names_from = TimepointNr,
                    names_prefix = 'T',
                    values_from = c('z_MoCA__total','Up3BradySum')) %>%
        mutate(Gender=factor(Gender),
               Age=Age-mean(Age,na.rm=T),
               YearsSinceDiag=YearsSinceDiag-mean(YearsSinceDiag,na.rm=T),
               NpsEducYears=NpsEducYears-mean(NpsEducYears,na.rm=T),
               Up3BradySum_Td = Up3BradySum_T2 - Up3BradySum_T0,
               z_MoCA__total_Td = z_MoCA__total_T2 - z_MoCA__total_T0,
               Up3BradySum_Td = Up3BradySum_Td - mean(Up3BradySum_Td,na.rm=T),
               Up3BradySum_T0 = Up3BradySum_T0 - mean(Up3BradySum_T0,na.rm=T),
               z_MoCA__total_T0 = z_MoCA__total_T0 - mean(z_MoCA__total_T0,na.rm=T)) %>%
        na.omit()

m.dMOCA <- lm(z_MoCA__total_Td ~ Up3BradySum_Td + z_MoCA__total_T0 + Up3BradySum_T0 + Age + Gender + YearsSinceDiag, data = tmp.dMOCA)
a <- Anova(m.dMOCA,type=3)
print(a)
kable(a, 'pipe', digits = Inf) %>% print()
eff <- eta_squared(m.dMOCA)
kable(eff, 'pipe', digits = Inf) %>% print()
s <- summary(m.dMOCA)
tab_model(m.dMOCA, digits = 5, show.se = T)
print(s)
ggemmeans(m.dMOCA, terms='Up3BradySum_Td') %>% plot(add.data=T)

##### Effect of time on raw MoCA
tmp <- dfClin %>%
        filter(TimepointNr != 1) %>%
        select(pseudonym,TimepointNr,Up3OfBradySum,Up3OnBradySum,LEDD,MoCASum,Age,Gender,YearsSinceDiag,NpsEducYears) %>%
        pivot_longer(cols = c('Up3OfBradySum','Up3OnBradySum'),
                     names_to = 'Medication',
                     values_to = 'Up3BradySum') %>%
        mutate(Medication=factor(Medication,
                                 levels=c('Up3OfBradySum','Up3OnBradySum'),
                                 labels=c('Off','On')),
               TimepointNr=factor(TimepointNr),
               Age=Age-mean(Age,na.rm=T),
               YearSinceDiag=YearsSinceDiag-mean(YearsSinceDiag,na.rm=T),
               NpsEducYears=NpsEducYears-mean(NpsEducYears,na.rm=T)) %>%
        na.omit()

m.MOCA <- tmp %>%
        filter(Medication=='On') %>%
        lmer(MoCASum ~ TimepointNr + Age + Gender + YearsSinceDiag + NpsEducYears + (1|pseudonym), data = .)
tab_model(m.MOCA, digits = 5, show.se = T)
a <- Anova(m.MOCA, type=3)
print(a)
kable(a, 'pipe', digits=Inf) %>% print()
eta_squared(m.MOCA, alternative = 'two.sided') %>% kable(., 'pipe', digits = Inf) %>% print()
em <- emmeans(m.MOCA, ~ TimepointNr, type='response')
pairs(em) %>% kable(., 'pipe',digits=Inf)

##### Effect of time on medication
tmp <- dfClin %>%
        filter(TimepointNr != 1) %>%
        select(pseudonym,TimepointNr,Up3OfBradySum,Up3OnBradySum,LEDD,Age,Gender,YearsSinceDiag,NpsEducYears) %>%
        pivot_longer(cols = c('Up3OfBradySum','Up3OnBradySum'),
                     names_to = 'Medication',
                     values_to = 'Up3BradySum') %>%
        mutate(Medication=factor(Medication,
                                 levels=c('Up3OfBradySum','Up3OnBradySum'),
                                 labels=c('Off','On')),
               TimepointNr=factor(TimepointNr),
               Age=Age-mean(Age,na.rm=T),
               YearSinceDiag=YearsSinceDiag-mean(YearsSinceDiag,na.rm=T),
               NpsEducYears=NpsEducYears-mean(NpsEducYears,na.rm=T)) %>%
        na.omit()

m.LEDD <- tmp %>%
        filter(Medication=='On') %>%
        lmer(LEDD ~ TimepointNr + Age + Gender + YearsSinceDiag + NpsEducYears + (1|pseudonym), data = .)
tab_model(m.LEDD, digits = 5, show.se = T)
a <- Anova(m.LEDD, type=3)
print(a)
kable(a, 'pipe', digits = Inf) %>% print()
eff <- eta_squared(m.LEDD)
kable(eff, 'pipe', digits = Inf) %>% print()
em <- emmeans(m.LEDD, ~ TimepointNr, type='response')
pairs(em) %>% kable(., 'pipe',digits=Inf)

# Correlation between medication increase and clinical progression
tmp.dLEDD <- tmp %>%
        filter(Medication=='On') %>%
        pivot_wider(id_cols = c('pseudonym','Age','Gender','YearsSinceDiag','NpsEducYears'),
                    names_from = TimepointNr,
                    names_prefix = 'T',
                    values_from = c('LEDD','Up3BradySum')) %>%
        mutate(Gender=factor(Gender),
               Age=Age-mean(Age,na.rm=T),
               YearsSinceDiag=YearsSinceDiag-mean(YearsSinceDiag,na.rm=T),
               NpsEducYears=NpsEducYears-mean(NpsEducYears,na.rm=T),
               Up3BradySum_Td = Up3BradySum_T2 - Up3BradySum_T0,
               LEDD_Td = LEDD_T2 - LEDD_T0,
               Up3BradySum_Td = Up3BradySum_Td - mean(Up3BradySum_Td,na.rm=T),
               Up3BradySum_T0 = Up3BradySum_T0 - mean(Up3BradySum_T0,na.rm=T),
               LEDD_T0 = LEDD_T0 - mean(LEDD_T0,na.rm=T)) %>%
        na.omit()

m.dLEDD <- lm(LEDD_Td ~ Up3BradySum_Td + LEDD_T0 + Up3BradySum_T0 + Age + Gender + YearsSinceDiag, data = tmp.dLEDD)
Anova(m.dLEDD,type=3) %>% print()
eff <- eta_squared(m.dLEDD)
print(eff)
s <- summary(m.dLEDD)
tab_model(m.dLEDD, digits = 5, show.se = T)
print(s)
ggemmeans(m.dLEDD, terms='Up3BradySum_Td') %>% plot(add.data=T)
```

# *Response times

## *During fMRI

### *Group comparison

```{r rts-fMRI, echo=FALSE, warning=TRUE, message=TRUE}

f <- 'log(response_time) ~ 1 + ParticipantType*TimepointNr*trial_type + block + Age + Gender + NpsEducYears + RespHandIsDominant + (1+TimepointNr+trial_type+block|pseudonym)'

groups <- c('PD_POM','HC_PIT')
tmp <- dfFMRI %>%
    select(pseudonym, TimepointNr, YearsToFollowUp, Age,
           ParticipantType, trial_type, Gender,NpsEducYears,RespHandIsDominant,
           block, response_time) %>%
    na.omit() %>%
        mutate(ParticipantType = factor(ParticipantType,levels=c('HC_PIT','PD_POM'),
                                        labels=c('Ctrl','PD')),
               TimepointNr = factor(TimepointNr,levels=c(0,2),
                                    labels=c('BA','2Y')),
               trial_type = case_when(trial_type == 'NChoice1' ~ 'Single',
                                      trial_type == 'NChoice2' ~ 'Multiple',
                                      trial_type == 'NChoice3' ~ 'Multiple'),
               trial_type = factor(trial_type,levels=c('Single','Multiple'),
                                   labels=c('Single','Multiple')),
               Age=scale(Age,scale=F),
               NpsEducYears = scale(NpsEducYears,scale=F))
tmp.collapsed <- tmp %>% 
    group_by(pseudonym, ParticipantType, TimepointNr, trial_type) %>%
    summarise(across(c(response_time,Age,YearsToFollowUp,NpsEducYears), ~ mean(.x, na.rm=TRUE))) %>%
    ungroup() %>%
    na.omit()
tmp.n <- tmp.collapsed %>%
    group_by(pseudonym, ParticipantType,TimepointNr) %>%
    summarise(across(everything(), ~first(.x))) %>%
    ungroup() %>%
    select(ParticipantType,TimepointNr)
n <- tmp.n %>% select(ParticipantType,TimepointNr) %>% table()
print(n)

fit <- lmer(f, data=tmp, control = lmerControl(optimizer = 'bobyqa', optCtrl=list(maxfun=2e5))) 
tab_model(fit, digits=5, show.se=TRUE)
s <- summary(fit)
a <- Anova(fit,type=3)
print(a)
kable(a, 'pipe', digits = Inf) %>% print()
eta_squared(fit, alternative = 'two.sided') %>% kable(., 'pipe', digits = Inf)

em <- emmeans(fit, ~ trial_type*TimepointNr|ParticipantType, type='response')
contrast(em, interaction = c('revpairwise','revpairwise','revpairwise'), adjust='mvt', type='response', by=NULL) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction = c('revpairwise','revpairwise','revpairwise'), adjust='mvt', type='response') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

em <- emmeans(fit, ~ TimepointNr|ParticipantType, type='response')
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response', by=NULL) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

em <- emmeans(fit, ~ trial_type|ParticipantType, type='response')
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response', by=NULL) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

em <- emmeans(fit, ~ trial_type|TimepointNr, type='response')
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response', by=NULL) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

em <- emmeans(fit, ~ TimepointNr|trial_type, type='response')
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response', by=NULL) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

em <- emmeans(fit, ~ ParticipantType, type='response')
contrast(em, interaction = c('revpairwise')) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
em <- emmeans(fit, ~ TimepointNr, type='response')
contrast(em, interaction = c('revpairwise')) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
em <- emmeans(fit, ~ trial_type, type='response')
contrast(em, interaction = c('revpairwise')) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

g_ba <- tmp.collapsed %>%
        filter(TimepointNr=='BA') %>%
        ggplot(aes(x=ParticipantType, y = response_time, color=trial_type)) +
        geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
        geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
        stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
                     mapping = aes(group=trial_type), position = position_dodge(width=0.8)) +
        theme_bw() + 
        scale_x_discrete(labels=c('Control','PD-ON')) +
        scale_color_viridis_d(option='mako', begin = .1, end = .6,
                              direction = -1) + 
        ggtitle('') +
        ylab('Response time (s)') + xlab('Baseline') + 
        guides(color = guide_legend(title='Choice')) +
        ylim(0.45, 1.5) + 
        annotate(geom='text',label=paste0('N=',n[1,1]),x=1,y=0.45) +
        annotate(geom='text',label=paste0('N=',n[2,1]),x=2,y=0.45)
g_fu <- tmp.collapsed %>%
        filter(TimepointNr=='2Y') %>%
        ggplot(aes(x=ParticipantType, y = response_time, color=trial_type)) +
        geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
        geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
        stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
                     mapping = aes(group=trial_type), position = position_dodge(width=0.8)) +
        theme_bw() + 
        scale_x_discrete(labels=c('Control','PD-ON')) +
        scale_color_viridis_d(option='mako', begin = .1, end = .6,
                              direction = -1) + 
        ggtitle('') +
        ylab('Response time (s)') + xlab('Follow-up') + 
        guides(color = guide_legend(title='Choice')) +
        ylim(0.45, 1.5) + 
        annotate(geom='text',label=paste0('N=',n[1,2]),x=1,y=0.45) +
        annotate(geom='text',label=paste0('N=',n[2,2]),x=2,y=0.45)
y <- 1.4
d = 0.1
segmap <- data.frame(
        x =    c(1.00, 0.80,  1.80),
        xend = c(2.00, 1.20,  2.20),
        y =    c(y, y-d,    y-d),
        yend = c(y, y-d,    y-d)
)
sigmap <- data.frame(
        x =     c(1.50, 1.00,  2.00),
        y =     c(y, y-d*1,     y-d*1),
        label = c('***', '***', '***')
)
g_ba <- g_ba + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     size = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)
g_fu <- g_fu + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     size = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)
g <- ggarrange(g_ba,g_fu)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/ResponseTimes/RT_lmer_T0.png',
          g_ba, dpi=300, device='png', base_width = 4)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/ResponseTimes/RT_lmer_T2.png',
          g_fu, dpi=300, device='png', base_width = 4)

```

### *Association with bradykinesia

```{r rts-vs-clin, echo=FALSE, warning=TRUE, message=TRUE}

##### Data preparation #####

dfFMRI.c <- dfFMRI %>%
        filter(ParticipantType=='PD_POM') %>%
        group_by(pseudonym, ParticipantType, TimepointNr, trial_type, Age, Gender,YearsSinceDiag,NpsEducYears,RespHandIsDominant,
                 Up3BradySum_T0, Up3BradySum_T2, Up3BradySum_RawChange,
                 Up3BradySum.imp_T0, Up3BradySum.imp_T2, Up3BradySum.imp_RawChange,
                 z_MoCA__total_RawChange, z_MoCA__total_T0, z_MoCA__total_T2,
                 z_MoCA__total.imp_RawChange, z_MoCA__total.imp_T0, z_MoCA__total.imp_T2,
                 LEDD_RawChange, LEDD_T0, LEDD_T2,
                 LEDD.imp_RawChange, LEDD.imp_T0, LEDD.imp_T2) %>%
        summarise(response_time = mean(response_time)) %>%
        ungroup() %>%
        pivot_wider(id_cols = c('pseudonym','TimepointNr','ParticipantType','Age',
                                'Gender','YearsSinceDiag','NpsEducYears','RespHandIsDominant',
                                'Up3BradySum_T0','Up3BradySum_T2','Up3BradySum_RawChange',
                                'Up3BradySum.imp_T0','Up3BradySum.imp_T2','Up3BradySum.imp_RawChange',
                                'z_MoCA__total_RawChange', 'z_MoCA__total_T0', 'z_MoCA__total_T2',
                                'z_MoCA__total.imp_RawChange', 'z_MoCA__total.imp_T0', 'z_MoCA__total.imp_T2',
                                'LEDD_RawChange', 'LEDD_T0', 'LEDD_T2',
                                'LEDD.imp_RawChange', 'LEDD.imp_T0', 'LEDD.imp_T2'),
                    names_from = trial_type,
                    names_prefix = 'TT_',
                    values_from = response_time) %>%
        mutate(AS.3gt1 = TT_NChoice3-TT_NChoice1,
               AS.2gt1 = TT_NChoice2-TT_NChoice1,
               AS.23gt1 = round(((TT_NChoice2+TT_NChoice3)/2)-TT_NChoice1,digits = 5),
               AS.Mean = round((TT_NChoice3 + TT_NChoice2 + TT_NChoice1)/3,digits = 5)) %>%
        pivot_wider(id_cols = c('pseudonym','ParticipantType','Age','Gender','YearsSinceDiag','NpsEducYears','RespHandIsDominant',
                                'Up3BradySum_T0','Up3BradySum_T2','Up3BradySum_RawChange',
                                'Up3BradySum.imp_T0','Up3BradySum.imp_T2','Up3BradySum.imp_RawChange',
                                'z_MoCA__total_RawChange', 'z_MoCA__total_T0', 'z_MoCA__total_T2',
                                'z_MoCA__total.imp_RawChange', 'z_MoCA__total.imp_T0', 'z_MoCA__total.imp_T2',
                                'LEDD_RawChange', 'LEDD_T0', 'LEDD_T2',
                                'LEDD.imp_RawChange', 'LEDD.imp_T0', 'LEDD.imp_T2'),
                    names_from = TimepointNr,
                    names_sep = '.T',
                    values_from = c('TT_NChoice1','TT_NChoice2','TT_NChoice3','AS.3gt1','AS.2gt1','AS.23gt1','AS.Mean')) %>%
        mutate(AS.3gt1.Delta = AS.3gt1.T2 - AS.3gt1.T0,
               AS.2gt1.Delta = AS.2gt1.T2 - AS.2gt1.T0,
               AS.23gt1.Delta = AS.23gt1.T2 - AS.23gt1.T0,
               AS.Mean.Delta = AS.Mean.T2 - AS.Mean.T0,
               Age=scale(Age,scale=F),
               NpsEducYears = scale(NpsEducYears,scale=F),
               YearsSinceDiag=scale(YearsSinceDiag,scale=F))#,
               # Up3BradySum.imp_RawChange = scale(Up3BradySum.imp_RawChange,scale=F,center=F),
               # Up3BradySum.imp_T0 = scale(Up3BradySum.imp_T0,scale=F),
               # Up3BradySum.imp_T2 = scale(Up3BradySum.imp_T2,scale=F),
               # z_MoCA__total.imp_RawChange = scale(z_MoCA__total.imp_RawChange,scale=F),
               # z_MoCA__total.imp_T0 = scale(z_MoCA__total.imp_T0,scale=F),
               # z_MoCA__total.imp_T2 = scale(z_MoCA__total.imp_T2,scale=F),
               # LEDD.imp_RawChange = scale(LEDD.imp_RawChange,scale=F),
               # LEDD.imp_T0 = scale(LEDD.imp_T0,scale=F),
               # LEDD.imp_T2 = scale(LEDD.imp_T2,scale=F))

##### Visualization #####

# gp <- dfFMRI.c %>% select(Up3BradySum_RawChange, AS.Mean.Delta, AS.2gt1.Delta, AS.3gt1.Delta) %>% 
#         na.omit() %>%
#         ggpairs(., lower = list(continuous = 'smooth'))
# print(gp)

##### Analysis #####
isna <- apply(dfFMRI.c, 2, is.na) %>% colSums()

cat(paste('\n ', 'Missing values\n', sep=''))
missing_perc <- round(isna/nrow(dfFMRI.c), digits = 5)*100
print(isna)

m01 <- lm(AS.Mean.Delta ~ Up3BradySum.imp_RawChange + Up3BradySum.imp_T0 + z_MoCA__total.imp_RawChange + z_MoCA__total.imp_T0 + LEDD.imp_RawChange + LEDD.imp_T0 + AS.Mean.T0 + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data = dfFMRI.c)
tab_model(m01, digits=7, show.se=TRUE)
s <- summary(m01,digits=5)
print(s)
a <- Anova(m01,type=3)
print(a)
kable(a, 'pipe', digits = Inf)
g <- ggemmeans(m01, terms = 'Up3BradySum.imp_RawChange') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Between-session change') +
        xlab('Change in bradykinesia (ON)') + ylab('Change in mean response time (s)') + 
        geom_hline(yintercept = 0, color = 'darkgrey', lty = 2) + geom_vline(xintercept = 0, color = 'darkgrey', lty = 2) +
        annotate(geom='text',
                 label=paste0('N=',length(predict(m01)),', ',
                              'β=',round(s$coefficients[2,1],digits=5),', ',
                              'SE=',round(s$coefficients[2,2],digits=5),', ',
                              'P=',round(s$coefficients[2,4],digits=5)),
                 x=-20,y=-0.58, hjust=0)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/ResponseTimes/RT_bradyprog_meanRTchange.png',
                  g, dpi=300, device='png', base_width = 4)

# Multiple > Single
m02 <- lm(AS.23gt1.Delta ~ Up3BradySum.imp_RawChange + Up3BradySum.imp_T0 + z_MoCA__total.imp_RawChange + z_MoCA__total.imp_T0 + LEDD.imp_RawChange + LEDD.imp_T0 + AS.23gt1.T0 + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data = dfFMRI.c)
tab_model(m02, digits=5, show.se=TRUE)
Anova(m02,type=3) %>% print()
g <- ggemmeans(m02, terms = 'Up3BradySum.imp_RawChange') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Clinical progression ~ Change in action selection (23-1)')
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/ResponseTimes/RT_bradyprog-3gt1change.png',
                  g, dpi=300, device='png', base_width = 5)

# Mean, baseline
m03 <- lm(log(AS.Mean.T0) ~ Up3BradySum.imp_T0 + z_MoCA__total.imp_T0 + LEDD.imp_T0 + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data = dfFMRI.c)
tab_model(m03, digits=7, show.se=TRUE)
s <- summary(m03)
print(s)
a <- Anova(m03,type=3)
print(a)
kable(a, 'pipe', digits = Inf)
g <- ggemmeans(m03, terms = 'Up3BradySum.imp_T0') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Baseline') +
        xlab('Bradykinesia severity (ON)') + ylab('Mean response time (s)') + 
        annotate(geom='text',
                 label=paste0('N=',length(predict(m03)),', ',
                              'β=',round(s$coefficients[2,1],digits=5),', ',
                              'SE=',round(s$coefficients[2,2],digits=5),', ',
                              'P=',round(s$coefficients[2,4],digits=5)),
                 x=0,y=0.3, hjust=0) + ylim(0.3,1.6) + xlim(0,40)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/ResponseTimes/RT_BA-mean.png',
                  g, dpi=300, device='png', base_width = 4)
#Mean, follow-up
m04 <- lm(log(AS.Mean.T2) ~ Up3BradySum.imp_T2 + z_MoCA__total.imp_T2 + LEDD.imp_T2 + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data = dfFMRI.c)
s <- summary(m04)
tab_model(m04, digits=7, show.se=TRUE)
a <- Anova(m04,type=3)
print(a)
kable(a, 'pipe', digits = Inf)
g <- ggemmeans(m04, terms = 'Up3BradySum.imp_T2') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Two-year follow-up') +
        xlab('Bradykinesia severity (ON)') + ylab('Mean response time (s)') + 
        annotate(geom='text',
                 label=paste0('N=',length(predict(m04)),', ',
                              'β=',round(s$coefficients[2,1],digits=5),', ',
                              'SE=',round(s$coefficients[2,2],digits=5),', ',
                              'P=',round(s$coefficients[2,4],digits=5)),
                 x=0,y=0.3, hjust=0) + ylim(0.3,1.6) + xlim(0,40)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/ResponseTimes/RT_FU-mean.png',
                  g, dpi=300, device='png', base_width = 4)

# Multiple > Single, baseline
m05 <- lm(AS.23gt1.T0 ~ Up3BradySum.imp_T0 + z_MoCA__total.imp_T0 + LEDD.imp_T0 + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data = dfFMRI.c)
tab_model(m05, digits=5, show.se=TRUE)
Anova(m05,type=3) %>% print()
g <- ggemmeans(m05, terms = 'Up3BradySum.imp_T0') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Baseline severity ~ 23 > 1')
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/ResponseTimes/RT_BA-32gt1.png',
                  g, dpi=300, device='png', base_width = 5)
# Multiple > Single, follow-up
m06 <- lm(AS.23gt1.T2 ~ Up3BradySum.imp_T2 + z_MoCA__total.imp_T2 + LEDD.imp_T2 + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data = dfFMRI.c)
tab_model(m06, digits=5, show.se=TRUE)
Anova(m06,type=3) %>% print()
g <- ggemmeans(m06, terms = 'Up3BradySum.imp_T2') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Follow-up severity ~ 23 > 1')
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/ResponseTimes/RT_FU-32gt1.png',
                  g, dpi=300, device='png', base_width = 5)
```

## Practice

Not evaluated.

```{r rts-Practice, echo=FALSE, warning=TRUE, message=TRUE, eval=FALSE}

f <- 'log(response_time) ~ 1 + ParticipantType*YearsToFollowUp*trial_type + scale(Age) + Gender + (1+YearsToFollowUp|pseudonym)'

groups <- c('PD_POM','HC_PIT')
tmp <- dfPrac %>%
    select(pseudonym, TimepointNr, YearsToFollowUp, Age, Age_Dmean, ParticipantType, trial_type, Gender, response_time) %>%
    na.omit()
tmp.collapsed <- dfPrac %>% 
    group_by(pseudonym, ParticipantType, TimepointNr, trial_type) %>%
    summarise(across(c(response_time,Age,YearsToFollowUp), ~ mean(.x, na.rm=TRUE))) %>%
    ungroup() %>%
    na.omit()
# compare_rts(f=f, tmp=tmp, tmp.collapsed=tmp.collapsed)

```

# *Accuracy

## During fMRI

### *Group comparison

```{r err-fMRI, echo=FALSE, warning=TRUE, message=TRUE}

f <- 'error_rate ~ 1 + ParticipantType*TimepointNr*trial_type + block + Age + Gender + NpsEducYears + RespHandIsDominant + (1+TimepointNr+trial_type+block|pseudonym)'

groups <- c('PD_POM','HC_PIT')
tmp <- dfFMRI %>%
    select(pseudonym, TimepointNr, YearsToFollowUp, Age,
           ParticipantType, trial_type, Gender,NpsEducYears,RespHandIsDominant,
           block, error_rate, n_trials) %>%
    na.omit() %>%
        mutate(ParticipantType = factor(ParticipantType,levels=c('HC_PIT','PD_POM'),
                                        labels=c('Ctrl','PD')),
               TimepointNr = factor(TimepointNr,levels=c(0,2),
                                    labels=c('BA','2Y')),
               trial_type = case_when(trial_type == 'NChoice1' ~ 'Single',
                                      trial_type == 'NChoice2' ~ 'Multiple',
                                      trial_type == 'NChoice3' ~ 'Multiple'),
               trial_type = factor(trial_type,levels=c('Single','Multiple'),
                                   labels=c('Single','Multiple')),
               Age=scale(Age,scale=F),
               NpsEducYears=scale(NpsEducYears,scale=F))
tmp.collapsed <- tmp %>% 
        group_by(pseudonym, ParticipantType, TimepointNr, trial_type) %>%
        summarise(across(c(error_rate,Age,YearsToFollowUp,NpsEducYears), ~ mean(.x, na.rm=TRUE))) %>%
        mutate(error_rate = error_rate*100) %>%
    ungroup() %>%
    na.omit()
tmp.n <- tmp.collapsed %>%
    group_by(pseudonym, ParticipantType,TimepointNr) %>%
    summarise(across(everything(), ~first(.x))) %>%
    ungroup() %>%
    select(ParticipantType,TimepointNr)
n <- tmp.n %>% select(ParticipantType,TimepointNr) %>% table()
print(n)

fit <- glmer(f, data=tmp, family = binomial, weights = n_trials,
             control = glmerControl(optimizer = 'bobyqa', optCtrl=list(maxfun=2e5)))
tab_model(fit, digits=5, show.se=TRUE)
s <- summary(fit)
anova <- Anova(fit,type=3)
print(anova, digits = 5)

em <- emmeans(fit, ~ trial_type*TimepointNr|ParticipantType, type='response')
contrast(em, interaction = c('revpairwise','revpairwise','revpairwise'), adjust='mvt', type='response', by=NULL) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction = c('revpairwise','revpairwise','revpairwise'), adjust='mvt', type='response') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

em <- emmeans(fit, ~ TimepointNr|ParticipantType, type='response')
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response', by=NULL) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

em <- emmeans(fit, ~ trial_type|ParticipantType, type='response')
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response', by=NULL) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

em <- emmeans(fit, ~ trial_type|TimepointNr, type='response')
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response', by=NULL) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction = c('revpairwise','revpairwise'), adjust='mvt', type='response') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

em <- emmeans(fit, ~ ParticipantType, type='response')
contrast(em, interaction = c('revpairwise')) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
em <- emmeans(fit, ~ TimepointNr, type='response')
contrast(em, interaction = c('revpairwise')) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
em <- emmeans(fit, ~ trial_type, type='response')
contrast(em, interaction = c('revpairwise')) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

# Plotting
g_ba <- tmp.collapsed %>%
        filter(TimepointNr=='BA') %>%
        ggplot(aes(x=ParticipantType, y = error_rate, color=trial_type)) +
        geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
        geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
        stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
                     mapping = aes(group=trial_type), position = position_dodge(width=0.8)) +
        theme_bw() + 
        scale_x_discrete(labels=c('Control','PD-ON')) +
        scale_color_viridis_d(option='mako', begin = .1, end = .6,
                              direction = -1) + 
        ggtitle('Error rates during action selection') +
        ylab('Error rate (%)') + xlab('Baseline') + 
        guides(color = guide_legend(title='Choice')) +
        ylim(-5, 75) + 
        annotate(geom='text',label=paste0('N=',n[1,1]),x=1,y=-5) +
        annotate(geom='text',label=paste0('N=',n[2,1]),x=2,y=-5)
g_fu <- tmp.collapsed %>%
        filter(TimepointNr=='2Y') %>%
        ggplot(aes(x=ParticipantType, y = error_rate, color=trial_type)) +
        geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
        geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
        stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
                     mapping = aes(group=trial_type), position = position_dodge(width=0.8)) +
        theme_bw() + 
        scale_x_discrete(labels=c('Control','PD-ON')) +
        scale_color_viridis_d(option='mako', begin = .1, end = .6,
                              direction = -1) + 
        ggtitle('') +
        ylab('Error rate (%)') + xlab('Follow-up') + 
        guides(color = guide_legend(title='Choice')) +
        ylim(-5, 75) + 
        annotate(geom='text',label=paste0('N=',n[1,2]),x=1,y=-5) +
        annotate(geom='text',label=paste0('N=',n[2,2]),x=2,y=-5)
y <- 70
d = 5
segmap <- data.frame(
        x =    c(1.00, 0.80,  1.80),
        xend = c(2.00, 1.20,  2.20),
        y =    c(y, y-d,    y-d),
        yend = c(y, y-d,    y-d)
)
sigmap <- data.frame(
        x =     c(1.50, 1.00,  2.00),
        y =     c(y+2, y-d*1,     y-d*1),
        label = c('+', '*', '**')
)
g_ba <- g_ba + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     size = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)
g_fu <- g_fu + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     size = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)
g <- ggarrange(g_ba,g_fu)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Accuracy/ERR_lmer_T0.png',
          g_ba, dpi=300, device='png', base_width = 3.5)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Accuracy/ERR_lmer_T2.png',
          g_fu, dpi=300, device='png', base_width = 3.5)

```

### *Association with bradykinesia

```{r err-vs-clin, echo=FALSE, warning=TRUE, message=TRUE}

##### Data preparation #####

dfFMRI.c <- dfFMRI %>%
        filter(ParticipantType=='PD_POM') %>%
        group_by(pseudonym, ParticipantType, TimepointNr, trial_type, Age, Gender,
                 YearsSinceDiag,NpsEducYears,RespHandIsDominant, 
                 Up3BradySum_T0, Up3BradySum_T2, Up3BradySum_RawChange,
                 Up3BradySum.imp_T0, Up3BradySum.imp_T2, Up3BradySum.imp_RawChange,
                 z_MoCA__total_RawChange, z_MoCA__total_T0,z_MoCA__total_T2,
                 z_MoCA__total.imp_RawChange, z_MoCA__total.imp_T0,z_MoCA__total.imp_T2,
                 LEDD_RawChange, LEDD_T0,LEDD_T2,
                 LEDD.imp_RawChange, LEDD.imp_T0,LEDD.imp_T2) %>%
        summarise(error_rate = mean(error_rate)) %>%
        ungroup() %>%
        pivot_wider(id_cols = c('pseudonym','TimepointNr','ParticipantType','Age',
                                'Gender','YearsSinceDiag','NpsEducYears','RespHandIsDominant',
                                'Up3BradySum_T0','Up3BradySum_T2','Up3BradySum_RawChange',
                                'Up3BradySum.imp_T0','Up3BradySum.imp_T2','Up3BradySum.imp_RawChange',
                                'z_MoCA__total_RawChange', 'z_MoCA__total_T0', 'z_MoCA__total_T2',
                                'z_MoCA__total.imp_RawChange', 'z_MoCA__total.imp_T0', 'z_MoCA__total.imp_T2',
                                'LEDD_RawChange', 'LEDD_T0', 'LEDD_T2',
                                'LEDD.imp_RawChange', 'LEDD.imp_T0', 'LEDD.imp_T2'),
                    names_from = trial_type,
                    names_prefix = 'TT_',
                    values_from = error_rate) %>%
        mutate(AS.3gt1 = TT_NChoice3-TT_NChoice1,
               AS.2gt1 = TT_NChoice2-TT_NChoice1,
               AS.23gt1 = round(((TT_NChoice2+TT_NChoice3)/2)-TT_NChoice1,digits = 5),
               AS.Mean = round((TT_NChoice3 + TT_NChoice2 + TT_NChoice1)/3,digits = 5)) %>%
        pivot_wider(id_cols = c('pseudonym','ParticipantType','Age','Gender','YearsSinceDiag',
                                'NpsEducYears','RespHandIsDominant',
                                'Up3BradySum_T0','Up3BradySum_T2','Up3BradySum_RawChange',
                                'Up3BradySum.imp_T0','Up3BradySum.imp_T2','Up3BradySum.imp_RawChange',
                                'z_MoCA__total_RawChange', 'z_MoCA__total_T0', 'z_MoCA__total_T2',
                                'z_MoCA__total.imp_RawChange', 'z_MoCA__total.imp_T0', 'z_MoCA__total.imp_T2',
                                'LEDD_RawChange', 'LEDD_T0', 'LEDD_T2',
                                'LEDD.imp_RawChange', 'LEDD.imp_T0', 'LEDD.imp_T2'),
                    names_from = TimepointNr,
                    names_sep = '.T',
                    values_from = c('TT_NChoice1','TT_NChoice2','TT_NChoice3','AS.3gt1','AS.2gt1','AS.23gt1','AS.Mean')) %>%
        mutate(AS.3gt1.Delta = AS.3gt1.T2 - AS.3gt1.T0,
               AS.2gt1.Delta = AS.2gt1.T2 - AS.2gt1.T0,
               AS.23gt1.Delta = AS.23gt1.T2 - AS.23gt1.T0,
               AS.Mean.Delta = AS.Mean.T2 - AS.Mean.T0,
               Age=scale(Age,scale=F),
               NpsEducYears = scale(NpsEducYears,scale=F),
               YearsSinceDiag=scale(YearsSinceDiag,scale=F),
               Up3BradySum.imp_RawChange = scale(Up3BradySum.imp_RawChange,scale=F),
               Up3BradySum.imp_T0 = scale(Up3BradySum.imp_T0,scale=F),
               Up3BradySum.imp_T2 = scale(Up3BradySum.imp_T2,scale=F),
               z_MoCA__total.imp_RawChange = scale(z_MoCA__total.imp_RawChange,scale=F),
               z_MoCA__total.imp_T0 = scale(z_MoCA__total.imp_T0,scale=F),
               z_MoCA__total.imp_T2 = scale(z_MoCA__total.imp_T2,scale=F),
               LEDD.imp_RawChange = scale(LEDD.imp_RawChange,scale=F),
               LEDD.imp_T0 = scale(LEDD.imp_T0,scale=F),
               LEDD.imp_T2 = scale(LEDD.imp_T2,scale=F))

##### Visualization #####

# gp <- dfFMRI.c %>% select(Up3BradySum_RawChange, AS.Mean.Delta, AS.2gt1.Delta, AS.3gt1.Delta) %>% 
#         na.omit() %>%
#         ggpairs(., lower = list(continuous = 'smooth'))
# print(gp)

##### Analysis #####
isna <- apply(dfFMRI.c, 2, is.na) %>% colSums()

cat(paste('\n ', 'Missing values\n', sep=''))
missing_perc <- round(isna/nrow(dfFMRI.c), digits = 5)*100
print(isna)

# Mean, delta
m01 <- lm(AS.Mean.Delta ~ Up3BradySum.imp_RawChange + Up3BradySum.imp_T0 + z_MoCA__total.imp_RawChange + z_MoCA__total.imp_T0 + LEDD.imp_RawChange + LEDD.imp_T0 + AS.Mean.T0 + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data = dfFMRI.c)
tab_model(m01, digits=5, show.se=TRUE)
Anova(m01,type=3) %>% print()
g <- ggemmeans(m01, terms = 'Up3BradySum.imp_RawChange') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('') +
        xlab('Change in bradykinesia') + ylab('Change in mean error rate')
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Accuracy/ER_clinprog-meanchange.png',
                  g, dpi=300, device='png', base_width = 5)

# Selection, delta
m02 <- lm(AS.23gt1.Delta ~ Up3BradySum.imp_RawChange + Up3BradySum.imp_T0 + z_MoCA__total.imp_RawChange + z_MoCA__total.imp_T0 + LEDD.imp_RawChange + LEDD.imp_T0  + AS.23gt1.T0 + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data = dfFMRI.c)
tab_model(m02, digits=5, show.se=TRUE)
Anova(m02,type=3) %>% print()
g <- ggemmeans(m02, terms = 'Up3BradySum.imp_RawChange') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Clinical progression ~ Change in action selection (23-1)')
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Accuracy/ER_clinprog-23gt1.png',
                  g, dpi=300, device='png', base_width = 5)

# Mean, baseline
m03 <- lm(AS.Mean.T0 ~ Up3BradySum.imp_T0 + z_MoCA__total.imp_T0 + LEDD.imp_T0 + Age + Gender + YearsSinceDiag, data = dfFMRI.c)
tab_model(m03, digits=5, show.se=TRUE)
Anova(m03,type=3) %>% print()
g <- ggemmeans(m03, terms = 'Up3BradySum.imp_T0') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Baseline severity ~ Mean error rate')
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Accuracy/ER_BA-mean.png',
                  g, dpi=300, device='png', base_width = 5)

# Mean, follow-up
m04 <- lm(AS.Mean.T2 ~ Up3BradySum.imp_T2 + z_MoCA__total.imp_T2 + LEDD.imp_T2 + Age + Gender + YearsSinceDiag, data = dfFMRI.c)
tab_model(m04, digits=5, show.se=TRUE)
Anova(m04,type=3) %>% print()
g <- ggemmeans(m04, terms = 'Up3BradySum.imp_T2') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Follow-up severity ~ Mean error rate')
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Accuracy/ER_FU-mean.png',
                  g, dpi=300, device='png', base_width = 5)

# Selection, baseline
m05 <- lm(AS.23gt1.T0 ~ Up3BradySum.imp_T0 + z_MoCA__total.imp_T0 + LEDD.imp_T0 + Age + Gender + YearsSinceDiag, data = dfFMRI.c)
tab_model(m05, digits=5, show.se=TRUE)
Anova(m05,type=3) %>% print()
g <- ggemmeans(m05, terms = 'Up3BradySum.imp_T0') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Baseline severity ~ 23 > 1')
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Accuracy/ER_BA-23gt1.png',
                  g, dpi=300, device='png', base_width = 5)

# Selection, follow-up
m06 <- lm(AS.23gt1.T2 ~ Up3BradySum.imp_T2 + z_MoCA__total.imp_T2 + LEDD.imp_T2 + Age + Gender + YearsSinceDiag, data = dfFMRI.c)
tab_model(m06, digits=5, show.se=TRUE)
Anova(m06,type=3) %>% print()
g <- ggemmeans(m06, terms = 'Up3BradySum.imp_T2') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Follow-up severity ~ 23 > 1')
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Accuracy/ER_FU-23gt1.png',
                  g, dpi=300, device='png', base_width = 5)
```

## Practice

Not evaluated.

```{r err-Prac, echo=FALSE, warning=TRUE, message=TRUE, eval=FALSE}

f <- 'error_rate ~ 1 + ParticipantType*YearsToFollowUp*trial_type + scale(Age.ba) + Gender + (1+YearsToFollowUp|pseudonym)'

groups <- c('PD_POM','HC_PIT')
tmp <- dfPrac %>%
    select(pseudonym, TimepointNr, YearsToFollowUp, Age, Age_Dmean, Age.ba, ParticipantType, trial_type, Gender, error_rate, n_trials) %>%
    na.omit()
# Collapse across blocks for visualization purposes
tmp.collapsed <- dfPrac %>% 
    group_by(pseudonym, ParticipantType, TimepointNr, trial_type) %>%
    summarise(across(c(error_rate,Age,YearsToFollowUp), ~ mean(.x, na.rm=TRUE))) %>%
    ungroup() %>%
    mutate(error_rate = error_rate*100) %>%
    na.omit()
compare_correct(f=f, tmp=tmp, tmp.collapsed = tmp.collapsed)

```

# *Post hoc analyses

## Speed-accuracy trade-off 

```{r SAT, echo=FALSE, warning=TRUE, message=TRUE, eval=FALSE}

dfSAT <- dfTask %>%
        group_by(pseudonym, ParticipantType, TimepointNr) %>%
        summarise(RT=median(response_time),
                  ERR=mean(error_rate)*100) %>%
        ungroup()

# ggplot(test, aes(x=RT,y=ERR,color=TimepointNr)) + 
#         geom_point() +
#         geom_smooth(method='loess', se = FALSE) + 
#         facet_grid(~ParticipantType)

# Show relationship between error rates and response times
# Is there a speed accuracy trade-off?
# https://rpubs.com/ablonder/198840
rt_range <- 2
n_bins <- 30
break_seq <- seq(0, rt_range, rt_range/n_bins)

timeslice_range <- dfSAT %>%
        na.omit() %>%
        mutate(RT_bin = cut(RT, breaks = break_seq)) %>%
        group_by(RT_bin, ParticipantType, TimepointNr) %>%
        mutate(RT_bin_avg = mean(RT))

count_range <- timeslice_range %>%
        group_by(RT_bin, ParticipantType, TimepointNr) %>%
        summarise(subjcount = n_distinct(pseudonym), totalcount = n())

timeslice_range <- timeslice_range %>%
        group_by(RT_bin_avg, ParticipantType, TimepointNr, pseudonym) %>% 
        summarise(ss_acc = mean(ERR)) %>% 
        group_by(RT_bin_avg, ParticipantType, TimepointNr) %>%
        summarise(mean = mean(ss_acc),
                  n = n())

g <- ggplot(timeslice_range, aes(x=RT_bin_avg, y=mean, weight=n, 
                                 color = TimepointNr, group = TimepointNr)) + 
        geom_point(aes(size = n)) +
        geom_smooth(method = "lm", formula = y ~ poly(x,2), se = FALSE) +
        scale_color_grey() +
        geom_hline(yintercept = 0, lty = "dashed") +
        ggtitle('Error rates as a function of RTs') +
        xlab("Average RT (s)") +
        ylab("Error rate (%)") +
        facet_grid(~ParticipantType)

print(g)

# Baseline
dfSAT_BA <- dfSAT %>%
        filter(TimepointNr==0)
ct_BA <- cor.test(dfSAT_BA$RT, dfSAT_BA$ERR, alternative = c('two.sided'), 
               method = c('spearman'), exact = FALSE)
print(ct_BA)
g <- ggplot(dfSAT_BA, aes(x=RT,y=ERR)) + 
        geom_point() + 
        geom_smooth(method='lm')
print(g)

# Follow-up
dfSAT_FU <- dfSAT %>%
        filter(TimepointNr==2)
ct_FU <- cor.test(dfSAT_FU$RT, dfSAT_FU$ERR, alternative = c('two.sided'), 
               method = c('spearman'), exact = FALSE)
print(ct_FU)
g <- ggplot(dfSAT_FU, aes(x=RT,y=ERR)) + 
        geom_point() + 
        geom_smooth(method='lm')
print(g)

# Comparison of correlation coefs, testing whether relationships between
# RTs and error rates differ by time point, which could indicate a larger 
# speed-accuracy trade-off at one visit compared to the other
# NOTE: Under construction
cocor <- cocor.indep.groups(r1.jk = ct_BA$estimate,
                   r2.hm = ct_FU$estimate,
                   n1 = nrow(dfSAT_BA),
                   n2 = nrow(dfSAT_FU),
                   alternative = 'two.sided')

print(cocor)
```

## *Missed responses

```{r miss, echo=FALSE, warning=TRUE, message=TRUE}

tmp <- dfFMRI %>%
        group_by(pseudonym,TimepointNr) %>%
        summarise(ParticipantType=first(ParticipantType),
                  total_misses_by_time=first(total_misses_by_time),
                  Up3BradySum=first(Up3BradySum),
                  z_MoCA__total=first(z_MoCA__total),
                  LEDD=first(LEDD),
                  Age=first(Age),
                  Gender=first(Gender),
                  YearsSinceDiag=first(YearsSinceDiag),
                  NpsEducYears=first(NpsEducYears),
                  RespHandIsDominant=first(RespHandIsDominant)) %>%
        ungroup() %>%
        mutate(ParticipantType=factor(ParticipantType),
               TimepointNr=factor(TimepointNr),
               Age=Age-mean(Age,na.rm=T),
               Gender=factor(Gender),
               YearsSinceDiag=YearsSinceDiag-mean(YearsSinceDiag,na.rm=T),
               NpsEducYears=NpsEducYears-mean(NpsEducYears,na.rm=T),
               RespHandIsDominant=factor(RespHandIsDominant))

# Group comparison
f <- 'total_misses_by_time ~ ParticipantType*TimepointNr + Age + Gender + NpsEducYears + RespHandIsDominant + (1|pseudonym)'
m <- glmer(f, data = tmp, family = 'poisson', control=glmerControl(optimizer = 'bobyqa'))
tab_model(m, title='Number of missed responses')
Anova(m,type=3) %>% print()
#effectsize(m)
em <- emmeans(m, ~ TimepointNr|ParticipantType, type='response')
contrast(em, interaction=c('revpairwise','revpairwise'), adjust = 'mvt', by=NULL) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction=c('revpairwise','revpairwise'), adjust = 'mvt', by = 'ParticipantType') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
contrast(em, interaction=c('revpairwise','revpairwise'), adjust = 'mvt', by = 'TimepointNr') %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
modplot <- predictorEffects(m)
plot(modplot)

# Clinical correlations
# Delta
m.lm <- tmp %>%
 filter(ParticipantType=='PD_POM') %>%
 pivot_wider(id_cols = c('pseudonym','Age','Gender','YearsSinceDiag',
                         'NpsEducYears','RespHandIsDominant'),
             names_from = TimepointNr,
             names_prefix = 'TT_',
             values_from = c(total_misses_by_time,Up3BradySum,z_MoCA__total,LEDD)) %>%
 mutate(total_misses_by_time_Delta = total_misses_by_time_TT_2 - total_misses_by_time_TT_0,
        Up3BradySum_Delta = Up3BradySum_TT_2 - Up3BradySum_TT_0,
        Up3BradySum_Delta = Up3BradySum_Delta-mean(Up3BradySum_Delta,na.rm=T),
        Up3BradySum_TT_0 = Up3BradySum_TT_0-mean(Up3BradySum_TT_0,na.rm=T),
        z_MoCA__total_Delta = z_MoCA__total_TT_2 - z_MoCA__total_TT_0,
        z_MoCA__total_Delta = z_MoCA__total_Delta-mean(z_MoCA__total_Delta,na.rm=T),
        z_MoCA__total_TT_0 = z_MoCA__total_TT_0-mean(z_MoCA__total_TT_0,na.rm=T),
        LEDD_Delta = LEDD_TT_2 - LEDD_TT_0,
        LEDD_Delta = LEDD_Delta-mean(LEDD_Delta,na.rm=T),
        LEDD_TT_0 = LEDD_TT_0-mean(LEDD_TT_0,na.rm=T),
        total_misses_by_time_TT_0=total_misses_by_time_TT_0-mean(total_misses_by_time_TT_0,na.rm=T)) %>%
 na.omit() %>%
 lm(total_misses_by_time_Delta ~ Up3BradySum_Delta + z_MoCA__total_Delta + LEDD_Delta + total_misses_by_time_TT_0 + Up3BradySum_TT_0 + z_MoCA__total_TT_0 + LEDD_TT_0 + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data=.)
tab_model(m.lm, digits = 5, show.se = TRUE)
Anova(m.lm,type = 3) %>% print()

# Baseline
m.glm <- tmp %>%
 filter(ParticipantType=='PD_POM',
        TimepointNr==0) %>%
        mutate(Up3BradySum = Up3BradySum-mean(Up3BradySum,na.rm=T),
               z_MoCA__total = z_MoCA__total-mean(z_MoCA__total,na.rm=T),
               LEDD = LEDD-mean(LEDD,na.rm=T)) %>%
 na.omit() %>%
 glm(total_misses_by_time ~ Up3BradySum + z_MoCA__total + LEDD + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data=., family = 'poisson')
tab_model(m.glm, digits = 5, show.se = TRUE)
Anova(m.glm,type = 3) %>% print()
g <- ggemmeans(m.glm, terms = 'Up3BradySum') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Baseline severity ~ Number of misses')
print(g)

# Follow-up
m.glm <- tmp %>%
 filter(ParticipantType=='PD_POM',
        TimepointNr==2) %>%
        mutate(Up3BradySum = Up3BradySum-mean(Up3BradySum,na.rm=T),
               z_MoCA__total = z_MoCA__total-mean(z_MoCA__total,na.rm=T),
               LEDD = LEDD-mean(LEDD,na.rm=T)) %>%
 na.omit() %>%
 glm(total_misses_by_time ~ Up3BradySum + z_MoCA__total + LEDD + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data=., family = 'poisson')
tab_model(m.glm, digits = 5, show.se = TRUE)
Anova(m.glm,type = 3) %>% print()
g <- ggemmeans(m.glm, terms = 'Up3BradySum') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Baseline severity ~ Number of misses')
print(g)

# Plotting
g <- tmp %>%
        mutate(ParticipantType=factor(ParticipantType,levels=c('HC_PIT','PD_POM'),labels=c('Control','Patient'))) %>%
        ggplot(aes(x=TimepointNr, y = total_misses_by_time, color = ParticipantType)) +
        geom_jitter(alpha=0.2, width=0.1) +
        geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
        stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5) +
        facet_grid(~ParticipantType) + 
        theme_bw() + 
        scale_x_discrete(labels=c('Baseline','Two years')) +
        scale_color_viridis_d(option='mako', begin = .1, end = .6,
                              direction = -1) +
        guides(color = guide_legend(title='Group')) +
        ggtitle('Group comparison of miss rates') +
        ylab('Number of misses') + xlab('Time point') +
        ylim(-1,60)
# y <- 55
# d = 3
# segmap <- data.frame(
#         x =    c(1.00, 1.00,  2.00,  0.80,  1.80),
#         xend = c(2.00, 1.00,  2.00,  1.20,  2.20),
#         y =    c(y,    y,     y,     y-d*1, y-d*1),
#         yend = c(y,    y-d*1, y-d*1, y-d*1, y-d*1)
# )
# sigmap <- data.frame(
#         x =     c(1.50),
#         y =     c(y),
#         label = c('***')
# )
# g <- g + 
#         geom_segment(data=segmap,
#                      mapping = aes(x=x,xend=xend,y=y,yend=yend),
#                      inherit.aes = FALSE,
#                      size = lsize) +
#         geom_text(data=sigmap, 
#                   mapping = aes(x=x,y=y,label=label),
#                   inherit.aes = FALSE,
#                   size = sigsize)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Misc/Miss_lmer.png',
          g, dpi=300, device='png', base_width = 5)

```

## *Button selection variability

```{r CoV, echo=FALSE, warning=TRUE, message=TRUE}

tmp <- dfFMRI %>%
        group_by(pseudonym,TimepointNr) %>%
        summarise(ParticipantType=first(ParticipantType),
                  Button.Press.CoV=first(Button.Press.CoV),
                  Up3BradySum=first(Up3BradySum),
                  Up3BradySum=first(Up3BradySum),
                  z_MoCA__total=first(z_MoCA__total),
                  LEDD=first(LEDD),
                  Age=first(Age),
                  Gender=first(Gender),
                  YearsSinceDiag=first(YearsSinceDiag),
                  NpsEducYears=first(NpsEducYears),
                  RespHandIsDominant=first(RespHandIsDominant)) %>%
        ungroup() %>%
        mutate(ParticipantType=factor(ParticipantType),
               TimepointNr=factor(TimepointNr),
               Age=Age-mean(Age,na.rm=T),
               Gender=factor(Gender),
               YearsSinceDiag=YearsSinceDiag-mean(YearsSinceDiag,na.rm=T),
               NpsEducYears=NpsEducYears-mean(NpsEducYears,na.rm=T),
               RespHandIsDominant=factor(RespHandIsDominant)) %>%
        filter(Button.Press.CoV>0.001)

# Group comparison
m.lmer <- tmp %>%
        lmer(log(Button.Press.CoV) ~ ParticipantType*TimepointNr + Age + Gender + NpsEducYears + RespHandIsDominant + (1|pseudonym), data=.)
tab_model(m.lmer, title='Coefficient of variation')
Anova(m.lmer, type=3) %>% print(digits=5)
eta_squared(m.lmer) %>% print(digits=5)
ems <- emmeans(m.lmer, revpairwise ~ ParticipantType, type = 'response')
summary(ems) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
ems <- emmeans(m.lmer, revpairwise ~ TimepointNr, type = 'response')
summary(ems) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

# Clinical correlations
# Delta
m.lm <- tmp %>%
 filter(ParticipantType=='PD_POM') %>%
 pivot_wider(id_cols = c('pseudonym','Age','Gender','YearsSinceDiag',
                         'NpsEducYears','RespHandIsDominant'),
             names_from = TimepointNr,
             names_prefix = 'TT_',
             values_from = c(Button.Press.CoV,Up3BradySum,z_MoCA__total,LEDD)) %>%
 mutate(Button.Press.CoV_Delta = Button.Press.CoV_TT_2 - Button.Press.CoV_TT_0,
        Up3BradySum_Delta = Up3BradySum_TT_2 - Up3BradySum_TT_0,
        Up3BradySum_Delta = Up3BradySum_Delta-mean(Up3BradySum_Delta,na.rm=T),
        Up3BradySum_TT_0 = Up3BradySum_TT_0-mean(Up3BradySum_TT_0,na.rm=T),
        z_MoCA__total_Delta = z_MoCA__total_TT_2 - z_MoCA__total_TT_0,
        z_MoCA__total_Delta = z_MoCA__total_Delta-mean(z_MoCA__total_Delta,na.rm=T),
        z_MoCA__total_TT_0 = z_MoCA__total_TT_0-mean(z_MoCA__total_TT_0,na.rm=T),
        LEDD_Delta = LEDD_TT_2 - LEDD_TT_0,
        LEDD_Delta = LEDD_Delta-mean(LEDD_Delta,na.rm=T),
        LEDD_TT_0 = LEDD_TT_0-mean(LEDD_TT_0,na.rm=T),
        Button.Press.CoV_TT_0=Button.Press.CoV_TT_0-mean(Button.Press.CoV_TT_0,na.rm=T)) %>%
 na.omit() %>%
 lm(Button.Press.CoV_Delta ~ Up3BradySum_Delta + z_MoCA__total_Delta + LEDD_Delta + Button.Press.CoV_TT_0 + Up3BradySum_TT_0 + z_MoCA__total_TT_0 + LEDD_TT_0 + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data=.)
tab_model(m.lm, digits = 5, show.se = TRUE)
Anova(m.lm,type = 3) %>% print()
ggemmeans(m.lm, terms = 'Up3BradySum_Delta') %>% plot()

# Baseline
m.lm <- tmp %>%
 filter(ParticipantType=='PD_POM',
        TimepointNr==0) %>%
        mutate(Up3BradySum = Up3BradySum-mean(Up3BradySum,na.rm=T),
               z_MoCA__total = z_MoCA__total-mean(z_MoCA__total,na.rm=T),
               LEDD = LEDD-mean(LEDD,na.rm=T)) %>%
 na.omit() %>%
 lm(Button.Press.CoV ~ Up3BradySum + z_MoCA__total + LEDD + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data=.)
tab_model(m.lm, digits = 5, show.se = TRUE)
Anova(m.lm,type = 3) %>% print()
g <- ggemmeans(m.lm, terms = 'Up3BradySum') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Baseline severity ~ BP_COV')
print(g)

# Follow-up
m.lm <- tmp %>%
 filter(ParticipantType=='PD_POM',
        TimepointNr==2) %>%
        mutate(Up3BradySum = Up3BradySum-mean(Up3BradySum,na.rm=T),
               z_MoCA__total = z_MoCA__total-mean(z_MoCA__total,na.rm=T),
               LEDD = LEDD-mean(LEDD,na.rm=T)) %>%
 na.omit() %>%
 lm(Button.Press.CoV ~ Up3BradySum + z_MoCA__total + LEDD + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data=.)
tab_model(m.lm, digits = 5, show.se = TRUE)
Anova(m.lm,type = 3) %>% print()
g <- ggemmeans(m.lm, terms = 'Up3BradySum') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Follow-up severity ~ BP_COV')
print(g)

# Plotting
g <- tmp %>%
        mutate(ParticipantType=factor(ParticipantType,levels=c('HC_PIT','PD_POM'),labels=c('Control','Patient'))) %>%
        ggplot(aes(x=TimepointNr, y = Button.Press.CoV, color = ParticipantType)) +
        geom_jitter(alpha=0.2, width=0.1) +
        geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
        stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5) +
        facet_grid(~ParticipantType) + 
        theme_bw() + 
        scale_x_discrete(labels=c('Baseline','Two years')) +
        scale_color_viridis_d(option='mako', begin = .1, end = .6,
                              direction = -1) +
        guides(color = guide_legend(title='Group')) +
        ggtitle('Group comparison of button press variability') +
        ylab('Coefficient of variation') + xlab('Time point') +
        ylim(0,1.5)
# y <- 1
# d = 0.4
# segmap <- data.frame(
#         x =    c(2.5),
#         xend = c(2.5),
#         y =    c(y),
#         yend = c(y-1*d),
#         ParticipantType=c('Patient')
# )
# sigmap <- data.frame(
#         x =     c(1.50),
#         y =     c(y),
#         label = c('***'),
#         ParticipantType=c('Patient')
# )
# g <- g +
#         geom_segment(data=segmap,
#                      mapping = aes(x=x,xend=xend,y=y,yend=yend),
#                      inherit.aes = FALSE,
#                      size = lsize) +
#         geom_text(data=sigmap,
#                   mapping = aes(x=x,y=y,label=label),
#                   inherit.aes = FALSE,
#                   size = sigsize)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Misc/CoV_lmer.png',
          g, dpi=300, device='png', base_width = 5)

```

## *Switch ratio

```{r switch-repeat, echo=FALSE, warning=TRUE, message=TRUE}

tmp <- dfFMRI %>%
        group_by(pseudonym,TimepointNr) %>%
        summarise(ParticipantType=first(ParticipantType),
                  Button.Press.SwitchRatio=first(Button.Press.SwitchRatio),
                  Up3BradySum=first(Up3BradySum),
                  Up3BradySum=first(Up3BradySum),
                  z_MoCA__total=first(z_MoCA__total),
                  LEDD=first(LEDD),
                  Age=first(Age),
                  Gender=first(Gender),
                  YearsSinceDiag=first(YearsSinceDiag),
                  NpsEducYears=first(NpsEducYears),
                  RespHandIsDominant=first(RespHandIsDominant)) %>%
        mutate(Button.Press.SwitchRatio=Button.Press.SwitchRatio*100) %>%
        ungroup() %>%
        mutate(ParticipantType=factor(ParticipantType),
               TimepointNr=factor(TimepointNr),
               Age=Age-mean(Age,na.rm=T),
               Gender=factor(Gender),
               YearsSinceDiag=YearsSinceDiag-mean(YearsSinceDiag,na.rm=T),
               NpsEducYears=NpsEducYears-mean(NpsEducYears,na.rm=T),
               RespHandIsDominant=factor(RespHandIsDominant))

# Group comparison
m.lmer <- tmp %>%
        lmer(log(Button.Press.SwitchRatio) ~ ParticipantType*TimepointNr + Age + Gender + NpsEducYears + RespHandIsDominant + (1|pseudonym), data=.)
tab_model(m.lmer, title='Switch-repeat ratio')
Anova(m.lmer, type=3) %>% print(digits=5)
eta_squared(m.lmer) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()
ems <- emmeans(m.lmer, revpairwise ~ ParticipantType, type = 'response')
summary(ems) %>% kable(., 'pipe', digits = Inf, padding=0) %>% print()

# Clinical correlations
# Delta
m.lm <- tmp %>%
 filter(ParticipantType=='PD_POM') %>%
 pivot_wider(id_cols = c('pseudonym','Age','Gender'),
             names_from = TimepointNr,
             names_prefix = 'TT_',
             values_from = c(Button.Press.SwitchRatio,Up3BradySum,z_MoCA__total,LEDD)) %>%
 mutate(Button.Press.SwitchRatio_Delta = Button.Press.SwitchRatio_TT_2 - Button.Press.SwitchRatio_TT_0,
        Up3BradySum_Delta = Up3BradySum_TT_2 - Up3BradySum_TT_0,
        Up3BradySum_Delta = Up3BradySum_Delta-mean(Up3BradySum_Delta,na.rm=T),
        Up3BradySum_TT_0 = Up3BradySum_TT_0-mean(Up3BradySum_TT_0,na.rm=T),
        z_MoCA__total_Delta = z_MoCA__total_TT_2 - z_MoCA__total_TT_0,
        z_MoCA__total_Delta = z_MoCA__total_Delta-mean(z_MoCA__total_Delta,na.rm=T),
        z_MoCA__total_TT_0 = z_MoCA__total_TT_0-mean(z_MoCA__total_TT_0,na.rm=T),
        LEDD_Delta = LEDD_TT_2 - LEDD_TT_0,
        LEDD_Delta = LEDD_Delta-mean(LEDD_Delta,na.rm=T),
        LEDD_TT_0 = LEDD_TT_0-mean(LEDD_TT_0,na.rm=T),
        Button.Press.SwitchRatio_TT_0=Button.Press.SwitchRatio_TT_0-mean(Button.Press.SwitchRatio_TT_0,na.rm=T)) %>%
 na.omit() %>%
 lm(Button.Press.SwitchRatio_Delta ~ Up3BradySum_Delta + z_MoCA__total_Delta + LEDD_Delta + Button.Press.SwitchRatio_TT_0 + Up3BradySum_TT_0 + z_MoCA__total_TT_0 + LEDD_TT_0 + Age + Gender, data=.)
tab_model(m.lm, digits = 5, show.se = TRUE)
Anova(m.lm,type = 3) %>% print()
ggemmeans(m.lm, terms='Up3BradySum_Delta') %>% plot()

# Baseline
m.lm <- tmp %>%
 filter(ParticipantType=='PD_POM',
        TimepointNr==0) %>%
        mutate(Up3BradySum = Up3BradySum-mean(Up3BradySum,na.rm=T),
               z_MoCA__total = z_MoCA__total-mean(z_MoCA__total,na.rm=T),
               LEDD = LEDD-mean(LEDD,na.rm=T)) %>%
 na.omit() %>%
 lm(Button.Press.SwitchRatio ~ Up3BradySum + z_MoCA__total + LEDD + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data=.)
tab_model(m.lm, digits = 5, show.se = TRUE)
Anova(m.lm,type = 3) %>% print()
g <- ggemmeans(m.lm, terms = 'Up3BradySum') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Baseline severity ~ BP_COV')
print(g)

# Follow-up
m.lm <- tmp %>%
 filter(ParticipantType=='PD_POM',
        TimepointNr==2) %>%
        mutate(Up3BradySum = Up3BradySum-mean(Up3BradySum,na.rm=T),
               z_MoCA__total = z_MoCA__total-mean(z_MoCA__total,na.rm=T),
               LEDD = LEDD-mean(LEDD,na.rm=T)) %>%
 na.omit() %>%
 lm(Button.Press.SwitchRatio ~ Up3BradySum + z_MoCA__total + LEDD + Age + Gender + YearsSinceDiag + NpsEducYears + RespHandIsDominant, data=.)
tab_model(m.lm, digits = 5, show.se = TRUE)
Anova(m.lm,type = 3) %>% print()
g <- ggemmeans(m.lm, terms = 'Up3BradySum') %>% plot(add.data=TRUE,color='darkred') + theme_bw() + ggtitle('Follow-up severity ~ BP_COV')
print(g)

# Plotting
g <- tmp %>%
        mutate(ParticipantType=factor(ParticipantType,levels=c('HC_PIT','PD_POM'),labels=c('Control','Patient'))) %>%
        ggplot(aes(x=TimepointNr, y = Button.Press.SwitchRatio, color = ParticipantType)) +
        geom_jitter(alpha=0.2, width=0.1) +
        geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
        stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5) +
        facet_grid(~ParticipantType) + 
        theme_bw() + 
        scale_x_discrete(labels=c('Baseline','Two years')) +
        scale_color_viridis_d(option='mako', begin = .1, end = .6,
                              direction = -1) +
        guides(color = guide_legend(title='Group')) +
        ggtitle('Group comparison of switching rates') +
        ylab('Switching rate (%)') + xlab('Time point') +
        ylim(0,100)
# y <- 1
# d = 0.4
# segmap <- data.frame(
#         x =    c(2.5),
#         xend = c(2.5),
#         y =    c(y),
#         yend = c(y-1*d),
#         ParticipantType=c('Patient')
# )
# sigmap <- data.frame(
#         x =     c(1.50),
#         y =     c(y),
#         label = c('***'),
#         ParticipantType=c('Patient')
# )
# g <- g +
#         geom_segment(data=segmap,
#                      mapping = aes(x=x,xend=xend,y=y,yend=yend),
#                      inherit.aes = FALSE,
#                      size = lsize) +
#         geom_text(data=sigmap,
#                   mapping = aes(x=x,y=y,label=label),
#                   inherit.aes = FALSE,
#                   size = sigsize)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/Misc/Switch_lmer.png',
          g, dpi=300, device='png', base_width = 5)

```

# *Brain activity

## *Voxelwise: Group comparison

### *AFNI

GROUP x TIME x CONDITION, analyzed using 3dLMEr

```{r pd-vs-hc, echo=FALSE, warning=TRUE, message=TRUE}

# Read data
dat <- read_csv('P:/3024006.02/Analyses/motor_task/Group/Longitudinal/AFNI/WholeBrain/3dLME_disease/stats/con_combined_Group2_x_TimepointNr2_x_Type3_z_Group_by_Type_dataTable_22-Jul-2024.txt') %>%
        left_join(., read_csv('P:/3024006.02/Analyses/motor_task/Group/Longitudinal/AFNI/WholeBrain/3dLME_disease/stats/con_combined_Group2_x_TimepointNr2_x_Type3_Z_Group_dataTable_22-Jul-2024.txt')) %>%
        mutate(trial_type = case_when(trial_type == '1c' ~ 'Single',
                                      trial_type == '23c' ~ 'Multiple'),
               trial_type = factor(trial_type, 
                                   levels = c('Single','Multiple'),
                                   labels = c('Single','Multiple')),
               TimepointNr = case_when(TimepointNr == 'T0' ~ 0,
                                       TimepointNr == 'T1' ~ 2),
               TimepointNr = factor(TimepointNr)) %>%
        group_by(Subj,Group,TimepointNr,trial_type) %>%
        summarise(con_combined__z_Group_by_Type_cid1 = mean(con_combined__z_Group_by_Type_cid1),
                  con_combined__z_Group_by_Type_cid2 = mean(con_combined__z_Group_by_Type_cid2),
                  con_combined__z_Group_by_Type_cid3 = mean(con_combined__z_Group_by_Type_cid3),
                  con_combined__z_Group_by_Type_cid4 = mean(con_combined__z_Group_by_Type_cid4),
                  con_combined__z_Group_by_Type_cid5 = mean(con_combined__z_Group_by_Type_cid5),
                  con_combined__z_Group_by_Type_cid6 = mean(con_combined__z_Group_by_Type_cid6),
                  con_combined__z_Group_by_Type_cid7 = mean(con_combined__z_Group_by_Type_cid7),
                  con_combined__z_Group_by_Type_cid8 = mean(con_combined__z_Group_by_Type_cid8),
                  con_combined__Z_Group_cid1 = mean(con_combined__Z_Group_cid1)) %>%
        ungroup()

n <- dat %>%
        count(Group,TimepointNr,trial_type)
print(n)

# # Group x Time
# print('>>> GROUP x TIME interaction')
# g <- dat %>%
#         ggplot(aes(x=Group,y=con_combined__z_Group_by_Type_cid1,color=TimepointNr)) + 
#         geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
#         geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
#         stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
#                      mapping = aes(group=TimepointNr), position = position_dodge(width=0.8)) +
#         theme_bw() + 
#         scale_x_discrete(labels=c('Ctrl','PD-ON')) +
#         scale_color_viridis_d(option='mako', begin = .1, end = .6,
#                               direction = -1) + 
#         ggtitle('') +
#         ylab('Precuneus (beta)') + xlab('') + 
#         guides(color = guide_legend(title='Year')) +
#         annotate(geom='text',label=paste0('N=',n[1,4]),x=0.8,y=-14) + 
#         annotate(geom='text',label=paste0('N=',n[3,4]),x=1.2,y=-14) +
#         annotate(geom='text',label=paste0('N=',n[5,4]),x=1.8,y=-14) + 
#         annotate(geom='text',label=paste0('N=',n[7,4]),x=2.2,y=-14) +
#         ylim(-14, 12)
# y <- 11
# d = 0.75
# segmap <- data.frame(
#         x =    c(1.00, 1.00,  2.00,  0.80,  1.80,  0.80,  1.20,  1.80,  2.20,  1.80),
#         xend = c(2.00, 1.00,  2.00,  1.20,  2.20,  0.80,  1.20,  1.80,  2.20,  2.20),
#         y =    c(y,    y,     y,     y-d*1, y-d*1, y-d*1, y-d*1, y-d*1, y-d*1, y-d*3),
#         yend = c(y,    y-d*1, y-d*1, y-d*1, y-d*1, y-d*2, y-d*2, y-d*2, y-d*2, y-d*3)
# )
# sigmap <- data.frame(
#         x =     c(1.50, 2.00),
#         y =     c(y,    y-d*3),
#         label = c('***', '***')
# )
# g <- g + 
#         geom_segment(data=segmap,
#                      mapping = aes(x=x,xend=xend,y=y,yend=yend),
#                      inherit.aes = FALSE,
#                      size = lsize) +
#         geom_text(data=sigmap, 
#                   mapping = aes(x=x,y=y,label=label),
#                   inherit.aes = FALSE,
#                   size = sigsize)
# print(g)
# save_plot('M:/Visualization/R_LongitudinalComparisonFmri/AFNI/3dLMEr_GroupComp_GROUPxTYPE_cb1.png',
#           g, dpi=300, device='png', base_width = 5)

# Group
print('>>> Main effect of GROUP')
g <- dat %>%
        ggplot(aes(x=Group,y=con_combined__Z_Group_cid1,color=TimepointNr)) + 
        geom_jitter(alpha=0.2, position = position_jitterdodge(jitter.width = 0.2)) +
        geom_boxplot(outlier.shape = NA, width=0.5, position = position_dodge(width=0.8)) + 
        stat_summary(fun.data = 'mean_cl_boot', geom = 'point', color='darkred', size=1.5, 
                     mapping = aes(group=TimepointNr), position = position_dodge(width=0.8)) +
        theme_bw() + 
        scale_x_discrete(labels=c('Ctrl','PD-ON')) +
        scale_color_viridis_d(option='mako', begin = .1, end = .6,
                              direction = -1) + 
        ggtitle('Decreased basal ganglia activity') +
        ylab('Putamen activity (beta)') + xlab('') + 
        guides(color = guide_legend(title='Year')) +
        annotate(geom='text',label=paste0('N=',n[1,4]),x=0.8,y=-2.9) + 
        annotate(geom='text',label=paste0('N=',n[3,4]),x=1.2,y=-2.9) +
        annotate(geom='text',label=paste0('N=',n[5,4]),x=1.8,y=-2.9) + 
        annotate(geom='text',label=paste0('N=',n[7,4]),x=2.2,y=-2.9) +
        ylim(-3, 6.5)
y <- 6
d = 0.5
segmap <- data.frame(
        x =    c(1.00, 1.00,  2.00,  0.80,  1.80),
        xend = c(2.00, 1.00,  2.00,  1.20,  2.20),
        y =    c(y,    y,     y,     y-d*1, y-d*1),
        yend = c(y,    y-d*1, y-d*1, y-d*1, y-d*1)
)
sigmap <- data.frame(
        x =     c(1.50),
        y =     c(y),
        label = c('***')
)
g <- g + 
        geom_segment(data=segmap,
                     mapping = aes(x=x,xend=xend,y=y,yend=yend),
                     inherit.aes = FALSE,
                     linewidth = lsize) +
        geom_text(data=sigmap, 
                  mapping = aes(x=x,y=y,label=label),
                  inherit.aes = FALSE,
                  size = sigsize)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/3dLMEr_Group_putamen.png',
          g, dpi=300, device='png', base_width = 5)

# Write for glmnet analysis
dat %>%
        filter(Group == 'PD_POM') %>%
        select(Subj, TimepointNr, trial_type, con_combined__Z_Group_cid1) %>%
        group_by(Subj, TimepointNr) %>%
        summarise(con_combined__Z_Group_cid1 = mean(con_combined__Z_Group_cid1)) %>%
        ungroup() %>%
        pivot_wider(id_cols = 'Subj',
                    names_from = 'TimepointNr',
                    names_prefix = 'T',
                    values_from = 'con_combined__Z_Group_cid1') %>%
        mutate(Td = T2 - T0) %>%
        rename(pseudonym=Subj, putamen_activity_T0 = T0, putamen_activity_T2 = T2,
               putamen_activity_Td = Td) %>%
        write_csv(., paste0(glmnet_dir,'glmnet_func_putamen.csv'))

```

## *Voxelwise: Brain-clinical associations

Correlation between change in bradykinesia and change in brain activity, analyzed using 3dttest++ / randomise adjusting for baseline severity and activity.

### *FSL

```{r brain-clin-fsl, echo=FALSE, warning=TRUE, message=TRUE, eval=T}

dat_delta <- read_csv('P:/3024006.02/Analyses/motor_task/Group/Longitudinal/FSL/stats/con_0007/vals/rand_delta_clincorr_all_vxlEV_tfce_corrp_tstat2_stats_stats_avg_agg.txt',
                      col_names = c('Subj','cid1_delta','cid2_delta')) %>%
        mutate(Subj = str_replace(Subj, '.nii',''),
               Subj = str_sub(Subj, start=80))

dat_ba <- read_csv('P:/3024006.02/Analyses/motor_task/Group/Longitudinal/FSL/stats/con_0007/vals/rand_delta_clincorr_all_vxlEV_tfce_corrp_tstat2_stats_BASELINE_stats_avg_agg.txt',
                   col_names = c('Subj','cid1_ba','cid2_ba')) %>%
        select(-Subj)

# dat_covs <- read_delim('P:/3024006.02/Analyses/motor_task/Group/Longitudinal/FSL/data/con_0007/covs__delta_clincorr_all_vxlEV.txt',
#                      col_names = c('Up3OnBradySum_delta','Up3OnBradySum_T0','MoCA_delta','MoCA_T0','LEDD_delta','LEDD_T0','Age','Gender','NpsEducYears','DisDur','RespHandIsDominant','mean','vxlEV'))

dat <- bind_cols(dat_delta, dat_ba) %>%
        mutate(Subj = str_sub(Subj, start=1L, end=16),
               Subj = str_glue('sub-POMU{Subj}')) %>%
        rename(pseudonym=Subj)

dat_covs <- dfFMRI %>%
        filter(ParticipantType=='PD_POM',
               TimepointNr==0) %>%
        group_by(pseudonym) %>%
        summarise(pseudonym=first(pseudonym),Up3BradySum.imp_RawChange=first(Up3BradySum.imp_RawChange),Up3BradySum.imp_T0=first(Up3BradySum.imp_T0),
                  z_MoCA__total.imp_RawChange=first(z_MoCA__total.imp_RawChange),z_MoCA__total.imp_T0=first(z_MoCA__total.imp_T0),
               LEDD.imp_RawChange=first(LEDD.imp_RawChange),LEDD.imp_T0=first(LEDD.imp_T0),Age=first(Age),Gender=first(Gender),NpsEducYears=first(NpsEducYears),
               YearsSinceDiag=first(YearsSinceDiag),RespHandIsDominant=first(RespHandIsDominant)) %>%
        mutate(Gender=factor(Gender),RespHandIsDominant=factor(RespHandIsDominant))

dat <- dat %>%
        left_join(dat_covs, by = 'pseudonym')

m <- dat %>%
        lm(cid2_delta ~ Up3BradySum.imp_RawChange + Up3BradySum.imp_T0 + cid2_ba + Age + Gender + NpsEducYears + YearsSinceDiag + RespHandIsDominant, data = .)
s <- summary(m)

tab_model(m,title='Brain-Clinical Associations',digits=5,show.se = T)
ols_vif_tol(m) %>% print()
Anova(m,type=3) %>% print()
LmPlots(m)
pred <- ggpredict(m, terms='Up3BradySum.imp_RawChange')
g <- plot(pred, add.data = TRUE,color='darkred') + theme_bw() + ggtitle('Premotor cortex (FEF)') +
        ylab('Change in activity (Multiple > Single, z)') + xlab('Change in bradykinesia (ON)') +
        annotate('text',y=-3,x=-20,
                 label=paste0('N=',length(predict(m)), ', ',
                              'β=',round(s$coefficients[2,1], digits=5), ', ',
                              'SE=',round(s$coefficients[2,2], digits=5), ', ',
                              'P<0.001 (FWEc-corrected)'), 
                 hjust=0) +
        ylim(-3,5.3) + 
        geom_hline(yintercept = 0, color = 'darkgrey', lty = 2) + geom_vline(xintercept = 0, color = 'darkgrey', lty = 2)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/FSL/randomise_severity_23gt1.png',
                  g, dpi=300, device='png', base_width = 5)

# Write for glmnet analysis
dat %>%
        select(pseudonym, cid2_ba, cid2_delta) %>%
        mutate(cid2_fu = cid2_ba + cid2_delta) %>%
        rename(premotor_activity_T0 = cid2_ba, premotor_activity_T2 = cid2_fu, premotor_activity_Td = cid2_delta) %>%
        relocate(pseudonym, premotor_activity_T0, premotor_activity_T2, premotor_activity_Td) %>%
        write_csv(., paste0(glmnet_dir,'glmnet_func_premotor.csv'))
        
```

## *Reproducability of correlations between selection-related activity and bradykinesia severity

Spherical ROIs from PMd and SPL (BRAIN, 2024)

```{r sphere_rois, echo=FALSE, warning=TRUE, message=TRUE, eval=T}

# Load data
roi_file <- c('P:/3024006.02/Analyses/motor_task/Group/Longitudinal/ExtractedRoiData/con_combined_ClustROI-both.csv')
if(grepl('con_combined_SphericalROI_2xPrefrontal_2xParietal', roi_file)){
        df.sphere <- read_csv(roi_file) %>%
                select(-c(InputFile,Subtype,YearsToFollowUp,Age_timevar)) %>%
                mutate(PMd_6ma = (vals1 + vals2) / 2,
                       PMd_6a = (vals3 + vals4) / 2,
                       SPL_7ma = (vals5 + vals6) / 2,
                       IPL_PFm = (vals7 + vals8) / 2,
                       PMd = (PMd_6ma + PMd_6a) / 2,
                       Parietal = (SPL_7ma+IPL_PFm) / 2,
                       Avg = (PMd_6ma + PMd_6a + SPL_7ma + IPL_PFm) / 4,
                       MAF_Avg = (vals1 + vals3 + vals5 + vals7) / 4) %>%
                rename(pseudonym = Subj, ParticipantType=Group, 
                       L_PMd_6ma = vals1, R_PMd_6ma = vals2, L_PMd_6a = vals3, R_PMd_6a = vals4,
                       L_SPL_7ma = vals5, R_SPL_7ma = vals6, L_IPL_PFm = vals7, R_IPL_PFm = vals8) %>%
                pivot_longer(cols = c(Avg, MAF_Avg, PMd_6ma, PMd_6a, SPL_7ma, IPL_PFm, PMd, Parietal,
                                      L_PMd_6ma, R_PMd_6ma, L_PMd_6a, R_PMd_6a,
                                      L_SPL_7ma, R_SPL_7ma, L_IPL_PFm, R_IPL_PFm),
                             names_to = 'ROI',
                             values_to = 'Beta') %>%
                pivot_wider(id_cols = c('pseudonym','ParticipantType','TimepointNr','Age','Sex','NpsEducYears','RespHandIsDominant','ROI'),
                            names_from = trial_type,
                            names_prefix = 'Cond_',
                            values_from = 'Beta') %>%
                mutate(SelectionActivity = Cond_23c - Cond_1c,
                       MotorActivity = (Cond_1c + Cond_23c) / 2,
                       TimepointNr = if_else(TimepointNr == 'T0',0,2),
                       TimepointNr = factor(TimepointNr),
                       Sex = factor(Sex),
                       RespHandIsDominant = factor(RespHandIsDominant)) %>%
                left_join(.,
                          dfClin %>% 
                                  filter(TimepointNr != 1) %>%
                                  select(pseudonym,TimepointNr,ParticipantType,YearsSinceDiag,Up3OfBradySum,Up3OnBradySum,
                                         z_MoCA__total, LEDD),
                          by = c('pseudonym','TimepointNr','ParticipantType'))
}else if(grepl('con_0007_SphericalROI_2xPrefrontal_2xParietal', roi_file)){
        df.sphere <- read_csv(roi_file) %>%
                select(-c(InputFile,Subtype,YearsToFollowUp,Age_timevar)) %>%
                mutate(PMd_6ma = (vals1 + vals2) / 2,
                       PMd_6a = (vals3 + vals4) / 2,
                       SPL_7ma = (vals5 + vals6) / 2,
                       IPL_PFm = (vals7 + vals8) / 2,
                       PMd = (PMd_6ma + PMd_6a) / 2,
                       Parietal = (SPL_7ma+IPL_PFm) / 2,
                       Avg = (PMd_6ma + PMd_6a + SPL_7ma + IPL_PFm) / 4,
                       MAF_Avg = (vals1 + vals3 + vals5 + vals7) / 4) %>%
                rename(pseudonym = Subj, ParticipantType=Group, 
                       L_PMd_6ma = vals1, R_PMd_6ma = vals2, L_PMd_6a = vals3, R_PMd_6a = vals4,
                       L_SPL_7ma = vals5, R_SPL_7ma = vals6, L_IPL_PFm = vals7, R_IPL_PFm = vals8) %>%
                pivot_longer(cols = c(Avg, MAF_Avg, PMd_6ma, PMd_6a, SPL_7ma, IPL_PFm, PMd, Parietal,
                                      L_PMd_6ma, R_PMd_6ma, L_PMd_6a, R_PMd_6a,
                                      L_SPL_7ma, R_SPL_7ma, L_IPL_PFm, R_IPL_PFm),
                             names_to = 'ROI',
                             values_to = 'Beta') %>%
                mutate(SelectionActivity = Beta,
                       TimepointNr = if_else(TimepointNr == 'T0',0,2),
                       TimepointNr = factor(TimepointNr),
                       Sex = factor(Sex),
                       RespHandIsDominant = factor(RespHandIsDominant)) %>%
                left_join(.,
                          dfClin %>% 
                                  filter(TimepointNr != 1) %>%
                                  select(pseudonym,TimepointNr,ParticipantType,YearsSinceDiag,Up3OfBradySum,Up3OnBradySum,
                                         z_MoCA__total, LEDD),
                          by = c('pseudonym','TimepointNr','ParticipantType'))
}else if(grepl('con_combined_ClustROI', roi_file)){
        df.sphere <- read_csv(roi_file) %>%
                select(-c(InputFile,Subtype,YearsToFollowUp,Age_timevar)) %>%
                rename(pseudonym = Subj, ParticipantType=Group) %>%
                pivot_longer(cols = c(vals),
                             names_to = 'ROI',
                             values_to = 'Beta') %>%
                pivot_wider(id_cols = c('pseudonym','ParticipantType','TimepointNr','Age','Sex','NpsEducYears','RespHandIsDominant','ROI'),
                            names_from = trial_type,
                            names_prefix = 'Cond_',
                            values_from = 'Beta') %>%
                mutate(SelectionActivity = Cond_23c - Cond_1c,
                       MotorActivity = (Cond_1c + Cond_23c) / 2,
                       ROI = 'ClustROI',
                       TimepointNr = if_else(TimepointNr == 'T0',0,2),
                       TimepointNr = factor(TimepointNr),
                       Sex = factor(Sex),
                       RespHandIsDominant = factor(RespHandIsDominant)) %>%
                # left_join(.,
                #           dfClin %>%
                #                   filter(TimepointNr != 1) %>%
                #                   select(pseudonym,TimepointNr,ParticipantType,YearsSinceDiag,Up3OfBradySum,Up3OnBradySum,
                #                          z_MoCA__total, LEDD),
                #           by = c('pseudonym','TimepointNr','ParticipantType'))
                left_join(.,
                          dfFMRI %>%
                                  filter(ParticipantType=='PD_POM') %>%
                                  group_by(pseudonym, ParticipantType) %>%
                                  summarise(Up3OfBradySum_T0=first(Up3OfBradySum_T0),
                                            Up3OfBradySum_T2=first(Up3OfBradySum_T2),
                                            Up3BradySum.imp_T0=first(Up3BradySum.imp_T0), 
                                            Up3BradySum.imp_T2=first(Up3BradySum.imp_T2),
                                            z_MoCA__total.imp_T0=first(z_MoCA__total.imp_T0),
                                            z_MoCA__total.imp_T2=first(z_MoCA__total.imp_T2),
                                            LEDD.imp_T0=first(LEDD.imp_T0),
                                            LEDD.imp_T2=first(LEDD.imp_T2)) %>%
                                  ungroup() %>%
                                  pivot_longer(cols = Up3OfBradySum_T0:LEDD.imp_T2,
                                               names_pattern = '(.*)_T(.*)',
                                               names_to = c('.value','TimepointNr')))
}else if(grepl('con_0007_ClustROI', roi_file)){
        df.sphere <- read_csv(roi_file) %>%
                select(-c(InputFile,YearsToFollowUp,Age_timevar)) %>%
                rename(pseudonym = Subj, ParticipantType=Group) %>%
                mutate(SelectionActivity=vals,
                       ROI = 'ClustROI',
                       TimepointNr = if_else(TimepointNr == 'T0',0,2),
                       TimepointNr = factor(TimepointNr),
                       Sex = factor(Sex),
                       RespHandIsDominant = factor(RespHandIsDominant)) %>%
                left_join(.,
                          dfClin %>% 
                                  filter(TimepointNr != 1) %>%
                                  select(pseudonym,TimepointNr,ParticipantType,YearsSinceDiag,Up3OfBradySum,Up3OnBradySum,
                                         z_MoCA__total, LEDD),
                          by = c('pseudonym','TimepointNr','ParticipantType'))
        
}

# Remove two subjects with 0 bradykinesia
df.sphere <- df.sphere %>%
        filter(!(ParticipantType == 'PD_POM' & Up3BradySum.imp<1))

sumtab <- df.sphere %>%
        group_by(ParticipantType,TimepointNr,ROI) %>%
        summarise(N = n(), AVG = mean(SelectionActivity,na.rm=T), SD = sd(SelectionActivity,na.rm=T))
print(sumtab)

test_and_plot_fun <- function(dat){
        
        dat1 <- dat
        
        # m <- lm(SelectionActivity ~ Up3OfBradySum + z_MoCA__total + LEDD + Age + Sex + NpsEducYears + RespHandIsDominant, data = dat1)
        m <- lm(SelectionActivity ~ Up3BradySum.imp + z_MoCA__total.imp + LEDD.imp + Age + Sex + NpsEducYears + RespHandIsDominant, data = dat1)
        s <- summary(m)
        a <- Anova(m,type=3)
        print(a)
        kable(a, 'pipe', digits = Inf) %>% print()
        tab_model(m, digits=7, show.se = T) %>% print()
        eta <- eta_squared(m)
        kable(eta, 'pipe', digits = Inf) %>% print()
        
        # Visualize (http://www.sthda.com/english/wiki/wiki.php?id_contents=7904)
        transparent_theme <- theme(
                axis.title.x = element_blank(),
                axis.title.y = element_blank(),
                axis.text.x = element_blank(), 
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                panel.grid = element_blank(),
                axis.line = element_blank(),
                panel.background = element_rect(fill = "transparent",colour = NA),
                plot.background = element_rect(fill = "transparent",colour = NA))
        # g1 <- dat1 %>%
        #         filter(ParticipantType=='PD_POM') %>%
        #         ggplot(aes(x=Up3OfBradySum,y=SelectionActivity)) + 
        #         geom_point(alpha = 0.5) + 
        #         geom_smooth(method='lm', color='darkred') + 
        #         theme_bw() + 
        #         xlim(0,42) + ylim(-2,7) + xlab('Bradykinesia severity (OFF)') + ylab('Beta (Multiple > Single)')
        g1 <- ggemmeans(m, terms = 'Up3BradySum.imp') %>% 
                plot(add.data=TRUE,color='darkred') + 
                theme_bw() + 
                ggtitle('') +
                xlim(0,40) + ylim(-2,7) + xlab('Bradykinesia severity (ON)') + ylab('Beta (Multiple > Single)')
        g2 <- dat1 %>%
                filter(ParticipantType=='HC_PIT') %>%
                ggplot(aes(x=factor(1),y=SelectionActivity)) + 
                geom_jitter(width=0.15,size=1, alpha=0.75, shape=17) + 
                geom_boxplot(width=0.5, alpha=0.75, fill='lightgrey', outlier.shape = NA) +
                ylim(-2,7) +
                transparent_theme
        g2_grob <- ggplotGrob(g2)
        xmin <- min(-5); xmax <- max(41)
        ymin <- min(-2); ymax <- max(5)
        g1 + 
                annotation_custom(grob = g2_grob,
                                  xmin = xmin+1, xmax = xmin+10, 
                                  ymin = ymin, ymax = ymax) + 
                annotate(geom = 'text', 
                         x=0, y=-1.9, hjust=0, 
                         label = paste0('N=',length(predict(m)),', ',
                                        'β=',round(s$coefficients[2,1],digits=5),', ',
                                        'SE=',round(s$coefficients[2,2],digits=5),', ',
                                        'P=',round(s$coefficients[2,4],digits=5)),
                         size=3.5)
        
}
g1 <- df.sphere %>% 
        filter(TimepointNr==0,
               ROI=='ClustROI') %>%
        test_and_plot_fun() + ggtitle('Baseline')
g2 <- df.sphere %>% 
        filter(TimepointNr==2,
               ROI=='ClustROI') %>%
        test_and_plot_fun() + ggtitle('Follow-up')
g <- ggarrange(g1,g2)
print(g)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/AFNI/ClustROI_Selection_vs_Bradykinesia_T0.png',
                  g1, dpi=300, device='png', base_width = 4)
save_plot('M:/Visualization/R_LongitudinalComparisonFmri/AFNI/ClustROI_Selection_vs_Bradykinesia_T2.png',
                  g2, dpi=300, device='png', base_width = 4)

```
