---
title: "R Notebook"
output: 
 html_notebook:
  toc: yes
  toc_float: true
  number_sections: true
---

```{r libs, include=F}
library(MASS)
library(simglm)
library(tidyverse)
library(GGally)
library(ggExtra)
library(ggeffects)
library(lme4)
library(rmcorr)
set.seed(160)
```

# Simulation exercise

## Simulate subject-specific intercepts and slopes

```{r}
# Determine parameters for simulation of intercepts and slopes
# These parameters 
N <- 350               # Sample size
perc_drop <- 0.05      # Percentage drop-out
mns <- c(16.50, 3.50, 6.70, 0.34)  # 1=Intercept, 2=Slope
sds <- c(7,10,  7.47, 2.04, 1.31)
cv_mat <- rbind(c( 1.00, -0.28, 0.05, 0.07),
                c(-0.28,  1.00, 0.02, -0.26),
                c(0.05,   0.02, 1.00, 0.13),
                c(0.07,   -0.26, 0.13, 1.00))  # Covariance between intercepts and slopes
# Note: cv_mat[4,2] and [2,4] controls 
subj_coefs <- data.frame(mvrnorm(n = N, mu = mns, Sigma = cv_mat))  # Simulate coefficients
colnames(subj_coefs) <- c('y_intercept', 'y_slope', 'x_intercept', 'x_slope')

# Check the covariance structure between variables
head(subj_coefs) %>% print()
ggpairs(subj_coefs, progress = F, lower = list(continuous = 'smooth'))



# REMOVABLE!! Random test of multicollinearity consequences
mns <- c(5, 5, 5, 5)  # 1=Intercept, 2=Slope
sds <- c(2, 2, 2, 2)
cv_mat <- rbind(c( 1.00, 0.28, 0.28, 0.28),
                c(0.28,  1.00, 0.02, 0.25),
                c(0.28,   0.02, 1.00, 0.85),
                c(0.28,   0.25, 0.85, 1.00)) 
subj_coefs <- data.frame(mvrnorm(n = N, mu = mns, Sigma = cv_mat))  # Simulate coefficients
colnames(subj_coefs) <- c('v1', 'v2', 'v3', 'v4')
# Check the covariance structure between variables
head(subj_coefs) %>% print()
ggpairs(subj_coefs, progress = F, lower = list(continuous = 'smooth'))

m1 <- lm(v1 ~ v2, data = subj_coefs)
m2 <- lm(v1 ~ v3, data = subj_coefs)
m3 <- lm(v1 ~ v4, data = subj_coefs)
m4 <- lm(v1 ~ v2 + v3, data = subj_coefs)
m5 <- lm(v1 ~ v2 + v4, data = subj_coefs)
m6 <- lm(v1 ~ v2 + v3 + v4, data = subj_coefs)
m7 <- lm(v1 ~ v3 + v4, data = subj_coefs)
```

## Generate data in wide format

```{r}
# Generate wide-format data set
sim_data_w <- bind_cols(subj_id = paste('sub-', as.character(seq(1,N)), sep = ''),
                      y_T0 = subj_coefs[,1],
                      y_T1 = subj_coefs[,1] + 1*subj_coefs[,2] + rnorm(1, 0, 1),
                      y_T2 = subj_coefs[,1] + 2*subj_coefs[,2] + rnorm(1, 0, 1),
                      y_T3 = subj_coefs[,1] + 3*subj_coefs[,2] + rnorm(1, 0, 1),
                      x_T0 = subj_coefs[,3],
                      x_T1 = subj_coefs[,3] + 1*subj_coefs[,4] + rnorm(1, 0, 1),
                      x_T2 = subj_coefs[,3] + 2*subj_coefs[,4] + rnorm(1, 0, 1),
                      x_T3 = subj_coefs[,3] + 3*subj_coefs[,4] + rnorm(1, 0, 1))
# Inject missing completely at random (MCAR)
sim_data_w$y_T1[rbinom(nrow(sim_data_w), 1, 1*perc_drop)>0] <- NA
sim_data_w$y_T2[rbinom(nrow(sim_data_w), 1, 2*perc_drop)>0] <- NA
sim_data_w$y_T3[rbinom(nrow(sim_data_w), 1, 3*perc_drop)>0] <- NA
sim_data_w$x_T1[rbinom(nrow(sim_data_w), 1, 1*perc_drop)>0] <- NA
sim_data_w$x_T2[rbinom(nrow(sim_data_w), 1, 2*perc_drop)>0] <- NA
sim_data_w$x_T3[rbinom(nrow(sim_data_w), 1, 3*perc_drop)>0] <- NA

# Calculate deltas
sim_data_w <- sim_data_w %>%
        mutate(y_Td = y_T1-y_T0,
               x_Td = x_T1-x_T0)

head(sim_data_w)
ggpairs(sim_data_w[,10:11], progress = F, lower = list(continuous = 'smooth'))

# IN PROGRESS: Alternative data simulation through 'simglm'
#https://simglm.brandonlebeau.org/articles/simulation_arguments.html#correlated-fixed-and-random-effects
# Simulate time-varying variable #1
sim_arguments1 <- list(
        formula = y ~ 1 + time + (1 + time | id),
        reg_weights = c(18, 3.5),
        fixed = list(time = list(var_type = 'time')),
        randomeffect = list(int_id = list(variance = 6.1, var_level = 2),
                            time_id = list(variance = 3.2, var_level = 2)),
        correlate = list(random = data.frame(x='int_id',
                                             y='time_id',
                                             corr=0.3)),
        missing_data = list(miss_prop = 0.1, new_outcome = 'y_missing',
                            type = 'random'),
        sample_size = list(level1 = 5, level2 = 350)
)
nested_data1 <- sim_arguments1 %>%
        simulate_fixed(data = NULL, .) %>%
        simulate_randomeffect(sim_arguments1) %>%
        simulate_error(sim_arguments1) %>%
        correlate_variables(sim_arguments1) %>%
        generate_response(sim_arguments1) %>%
        generate_missing(sim_arguments1)
head(nested_data1)
s <- model_fit(nested_data1, sim_arguments1) %>%
        summary()
s$varcor; s$coefficients

sim_arguments2 <- list(
        formula = y ~ 1 + time + (1 + time | id),
        reg_weights = c(5, 0.78),
        fixed = list(time = list(var_type = 'time')),
        randomeffect = list(int_id = list(variance = 2.8, var_level = 2),
                            time_id = list(variance = 0.8, var_level = 2)),
        correlate = list(random = data.frame(x='int_id',
                                             y='time_id',
                                             corr=0.14)),
        missing_data = list(miss_prop = 0.1, new_outcome = 'y_missing',
                            type = 'random'),
        sample_size = list(level1 = 5, level2 = 350)
)
nested_data2 <- sim_arguments2 %>%
        simulate_fixed(data = NULL, .) %>%
        simulate_randomeffect(sim_arguments2) %>%
        simulate_error(sim_arguments2) %>%
        correlate_variables(sim_arguments2) %>%
        generate_response(sim_arguments2) %>%
        generate_missing(sim_arguments2)
head(nested_data2)
s <- model_fit(nested_data2, sim_arguments2) %>%
        summary()
s$varcor; s$coefficients

nested_data3 <- bind_cols(id=nested_data1$id,
                          time=nested_data1$time,
                          x=nested_data1$y_missing,
                          y=nested_data2$y_missing)

```

## Pivot to long format

```{r}
# Pivot to long-format data set
sim_data_l1 <- sim_data_w %>%
        select(subj_id, y_T0, y_T1, x_T0, x_T1) %>%
 pivot_longer(cols = c(y_T0,y_T1,x_T0,x_T1),
              names_to = c('var','time'),
              names_pattern = '(.*)_(.*)',
              values_to = 'val') %>%
        mutate(time=as.numeric(factor(time))-1)
head(sim_data_l1) %>% print()

sim_data_l2 <- sim_data_l1 %>%
        pivot_wider(id_cols = c('subj_id','time'),
                    names_from = 'var',
                    values_from = 'val') %>%
 mutate(time=as.numeric(factor(time))-1)
head(sim_data_l2) %>% print()

```

# One group
# Two visits
# Two continuous scores
 # Both increasing
 # Both decreasing
 # One increasing, one decreasing
# Random missing values
# Inter-subject relationship
 # Positive
 # Negative
# Intra-subject relationship
 # Positive
 # Negative
# Covariates
 # Time-invariant categorical
 # Time-invariant continuous
 # Time-varying categorical
 # Time-varying continuous

# Analyses
 # Raw change
 # Single-corrected raw change
 # Double-corrected raw change
 # Growth curve modelling
  # Interaction with time
  # Disaggregation
  # Including baseline in model as covar (Fitzmaurize et al. 2015, p. 217)
 # rmcorr
 # Spearman